---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output: html_document
---

# Preface
This R Script is to be used to load, explore, and synthesize the data for assessing the (soil) ecological multifunctionality in the Jena experiment and it's response to different levels of plant diversity. The following ecological functions were are incuded: Aboveground biomass production, C&N content in the soil, Realized Species Richness, nitrogen leaching (into the groundwater), pollination, above- and belowground herbivory (for belowground measured as energy fluxes in the food web), soil organic carbon (SOC), leaf dry matter content (LDMC), soil diversity, microbial biomass + respiration. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(warn = TRUE, 
                      message = FALSE, 
                      error = FALSE, 
                      echo = TRUE, 
                      cache = FALSE, 
                      fig.width = 8, 
                      fig.height = 6, 
                      dpi = 300,
                      out.width = "100%",
                      fig.align = "center",
                      comment = NA,
                      tidy = TRUE,
                      collapse = TRUE)
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data, include=FALSE}
# load data on plot design (needed for 'block code')
design <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Plot_Design/90_20_data.xlsx")

#load data on realized species richness (ID = 487)
SR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")
  #change data type for realized SR & sowndiv
  SR$realizedSR <- as.numeric(SR$realizedSR)
  SR$sowndiv <- as.numeric(SR$sowndiv)

#load data on aboveground biomass production (ID = 624) 
biomassDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/624_9_data.xlsx")
  biomassDat$plotcode <- as.factor(biomassDat$plotcode) # change plotcode to a factor

#load data on root biomass production (ID = 714) 
roots <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Root Biomass_03-24/714_4_data.xlsx") %>% 
  rename_at('plot', ~'plotcode') # rename plot to plotcode
  roots$plotcode <- as.factor(roots$plotcode) # change plotcode to a factor
  roots$year <- as.factor(roots$year) # change year to a factor  
  roots$root.bm <- as.numeric(roots$root.bm) # change root biomass values to numeric

#load data on soil CN (2020: ID = 424; 2011: ID = 254; 2002 - 2008: ID = 278): 
carbon20 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/424_7_data.xlsx")
carbon11 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/254_3_data.xlsx")
carbon02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/278_2_data.xlsx")
  # change data types
  carbon02$N_Soil <- as.numeric(carbon02$N_Soil)
  carbon02$Corg_Soil <- as.numeric(carbon02$Corg_Soil)
  carbon02$Cinorg_Soil <- as.numeric(carbon02$Cinorg_Soil)
  carbon02$BulkDensity_Soil <- as.numeric(carbon02$BulkDensity_Soil)
  carbon11$BulkDensity_Soil <- as.numeric(carbon11$BulkDensity_Soil)

  # merge the 2002, 2011 and 2020 datasets
  carbonAll <- bind_rows(carbon02, carbon11, carbon20)
  carbonAll$plot <- as.factor(carbonAll$plot) # change plot to a factor

# load data on microbial biomass and respiration from 2002 to 2020 (ID = 78) 2022 (ID = 398) and 2021 (ID = 240)
microbial1 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration_03-20/78_5_data.xlsx")
mirobial2 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration_21/240_6_data.xlsx") %>%
  filter(grepl("D3", plot)) %>%  # filter for treatment D3 (main experiment)
  mutate(plot = gsub("D3", "", plot))
  
  # change data types
  microbial1$basal_respiration <- as.numeric(microbial1$basal_respiration)
  microbial1$soil_microbial_biomass_C <- as.numeric(microbial1$soil_microbial_biomass_C)
  microbial1$respiratory_quotient <- as.numeric(microbial1$respiratory_quotient)
  microbial1$soil_water_content <- as.numeric(microbial1$soil_water_content)
  

  # merge the datasets
  microbial <- bind_rows(mirobial2, microbial1) %>%
    rename_at('plot', ~'plotcode') %>% # rename plot to plotcode
    mutate(plotcode = as.factor(plotcode)) # change plotcode to a factor



#load aboveground herbivory data (only treatment D3 -> main experiment)
herbivoryAbove22 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Above/480_4_data.xlsx") %>%  filter(treatment == "D3")
herbivoryAbove24 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Above/708_2_data.xlsx")

# columns to merge
keepCols <- c("plot", "percHerb", "consBM")

herbivoryAbove <- bind_rows(
  select(herbivoryAbove22, all_of(keepCols)),
                           select(herbivoryAbove24, all_of(keepCols))) %>%
  rename_at('plot', ~'plotcode') %>% # rename plot to plotcode
  mutate(plotcode = as.factor(plotcode)) %>%  # change plotcode to a factor
  mutate(percHerb = as.numeric(percHerb)) # change percHerb to numeric


#load data on soil diversity
  # Nematodes 2017 (ID = 323)
  nematodes <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Nematodes_dBEF_17/323_3_data.xlsx")
  # Bacteria
  # AMF (ID = 560)
  # Oomycota (ID = 538)
  # Cercozoa (ID = 537)
  # Arthropodes


#load data on nitrogen leaching (this should be in kg/ha per year)
leaching <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/Overall_Leaching_Plot.xlsx")

#load data on pollinators
pollinators <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/423_3_data.xlsx") %>% 
  mutate(species_richness = as.numeric(species_richness)) # change species richness to numeric

#load data on flower cover
flowers <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

#load LDMC data
# ldmc <- read_csv("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/LDMC/Leaf_Root_traits_dbef_2021.csv")

#load climate data
# climate <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Climate_Data/484_7_data.xlsx")

#load data on enzyme activity?

#soil infiltration rates

```

## Data wrangling
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated. 
```{r Biomass data}
# calculate biomass for each year per plot and average across subsamples
biomassAnnual <- biomassDat %>% 
  filter(year > 2003) %>% 
  group_by(plotcode, year, subsample) %>%
  summarise(sum_biomass = sum(biomass, na.rm = TRUE), .groups = "drop") %>% 
  group_by(plotcode, year) %>%
  summarise(mean_biomass = mean(sum_biomass, na.rm = TRUE), .groups = "drop")

biomassAnnual$year <- as.factor(biomassAnnual$year) # change year to a factor

# sum up biomass for each plot and calculate the mean biomass across years
biomassMeans <- biomassAnnual %>% 
  group_by(plotcode) %>% 
  summarise(mean_biomass = mean(mean_biomass, na.rm = TRUE), .groups = "drop") %>% 
  mutate(biomass_standardized = mean_biomass / max(mean_biomass)) # standardize the biomass values

# calculate mean biomass for root dataset
rootsMean <- roots %>% 
  group_by(plotcode) %>% 
  mutate(mean_root_biomass = mean(root.bm, na.rm = TRUE)) %>% 
  distinct(plotcode, mean_root_biomass) %>%
  ungroup() %>% 
  na.omit() %>%
  mutate(root_biomass_standardized = mean_root_biomass / max(mean_root_biomass)) # standardize the root biomass values

```


```{r Microbial data}
microbialMean <- microbial %>%
  filter(basal_respiration >= 0) %>%  # filter out negative values
  group_by(plotcode) %>% 
  mutate(mean_micbiomass = mean(soil_microbial_biomass_C, na.rm = T), mean_micrespiration = mean(basal_respiration, na.rm = T)) %>% 
  distinct(plotcode, mean_micbiomass, mean_micrespiration) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(micbiomass_standardized = mean_micbiomass / max(mean_micbiomass), # standardize the microbial biomass values
         micrespiration_standardized = mean_micrespiration / max(mean_micrespiration)) # standardize the microbial respiration values

```

```{r soil carbon}
# create a new column with mean C across layers per plot for all years
carbonMean <- carbonAll %>% group_by(plot) %>% 
                      mutate(SOC_overall_mean = mean(Corg_Soil, na.rm = TRUE)) %>% 
                      rename_at('plot', ~'plotcode') %>% 
                      distinct(plotcode, SOC_overall_mean) %>%
                      ungroup() %>%
                      mutate(SOC_standardized = SOC_overall_mean/max(SOC_overall_mean)) # standardize the values
  
```

```{r Species Richness data}
realdiv <- SR %>%                    #calculate the mean realized SR for the years 2003-2019
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)

SR <- distinct(SR, plotcode, sowndiv)

```


```{r leaching}
leaching <- leaching %>% 
            rename_at('plot', ~'plotcode') # rename plot to plotcode

# standardize the leaching values
leaching$leaching_standardized <- leaching$TNleaching_av / max(leaching$TNleaching_av)

# higher values are worse, therefore the values have to be inverted
leaching$leaching_standardized <- 1 - leaching$leaching_standardized

```


```{r Herbivory}
herbivoryMean <- herbivoryAbove %>% 
            group_by(plotcode) %>%
            mutate(meanHerb = mean(percHerb, na.rm = TRUE)) %>%
            distinct(plotcode, meanHerb) %>%
            ungroup() %>%
            na.omit() %>%
            mutate(herbivory_standardized = meanHerb / max(meanHerb)) # standardize the herbivory values

# higher values are worse, therefore the values have to be inverted
herbivoryMean$herbivory_standardized <- 1 - herbivoryMean$herbivory_standardized

```

```{r Pollinators}
# calculate the mean pollinator diversity per plot
pollinatorsMean <- pollinators %>% 
  group_by(plotcode) %>% 
  mutate(mean_diversity = mean(species_richness, na.rm = TRUE)) %>%
  distinct(plotcode, mean_diversity) %>%
  ungroup() %>%
  mutate(mean_div_standardized = mean_diversity / max(mean_diversity)) # standardize the pollinator richness values

```  

```{r Flower Cover}
# calculate the mean flower cover per plot
flowersMean <- flowers %>% 
  group_by(plotcode) %>% 
  mutate(mean_flower_cover = mean(flowercover.plot, na.rm = TRUE)) %>%
  distinct(plotcode, mean_flower_cover) %>%
  ungroup() %>%
  mutate(mean_flower_cover_standardized = mean_flower_cover / max(mean_flower_cover)) # standardize the flower cover values

```


## Synthesize in one Dataset
```{r Data Synthesis}
# merge all datasets into one dataset

ecologicalMF <- leaching %>% 
  left_join(microbialMean, by = join_by(plotcode)) %>% 
  left_join(rootsMean, by = join_by(plotcode)) %>% 
  left_join(biomassMeans, by = join_by(plotcode)) %>% 
  left_join(carbonMean, by = join_by(plotcode)) %>% 
  left_join(realdiv, by = join_by(plotcode)) %>%
  left_join(select(design, plotcode, block)) %>%
  left_join(SR, by = join_by(plotcode)) %>% 
  left_join(herbivoryMean, by = join_by(plotcode)) %>%
  left_join(pollinatorsMean, by = join_by(plotcode)) %>%
  left_join(flowersMean, by = join_by(plotcode))

# calculate the ecological multifunctionality as the mean of the standardized values  
ecologicalMF$multifunctionality <- rowMeans(
  ecologicalMF[, c(
    "biomass_standardized",
    "micbiomass_standardized",
    "root_biomass_standardized",
    "SOC_standardized",
    "leaching_standardized",
    "micrespiration_standardized",
    "mean_div_standardized",
    "mean_flower_cover_standardized",
    "herbivory_standardized"
  )],
  na.rm = TRUE
)

# save the ecological multifunctionality dataset
# writexl::write_xlsx(ecologicalMF, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality.xlsx")


```

