---
title: "Economic Multifunctionality Analysis"
author: "Marius Munschek"
date: "2025-05-28"
output: html_document
---

#Preface
This R Script is written by Marius Munschek to calculate the economic multifunctionality for ecosystem services in the Jena Experiment. The performance of the following ecosystem sevices per plot is averaged over the years 2011 - 2020, then translated into monetary values which are aggregated to obtain economic multifunctionality values per plot:
- Biomass production
- Nitrogen leaching
- Carbon sequestration

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, fig.width=10, fig.height=6, dpi=300, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", fig.path="figures/", out.width="100%", out.height="100%")

source("C:/Users/mm68tulo/Documents/R/functions/myfunctions.R")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
```

## Set up a theme
```{r Theme}
theme_set(theme_custom())
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
#load synthesis dataset
synthDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF.xlsx") %>% 
  filter(sowndiv < 60) %>%  #remove rows with NA in sowndiv and sowndiv greater than 16
  mutate(block = as.factor(block))
```

## Data wrangling
Change data types and filter for relevant observations.
```{r Synthesis data}
synthDat$plotcode <- as.factor(synthDat$plotcode) #change the plotcode to a factor

# add up all the low costs for lower total value
synthDat$costs_total_lo <- synthDat$SOC_costs_lo + synthDat$leach_costs_lo + synthDat$profit_mean

# add up all the high costs for higher total value
synthDat$costs_total_hi <- synthDat$SOC_costs_hi + synthDat$leach_costs_hi + synthDat$profit_mean

noMono <- synthDat %>% 
  filter(sowndiv > 1) #filter out monocultures

```


## Plot the densities of the data
```{r Densities}
# check distribution of sowndiv
p1 <- ggplot2::ggplot(synthDat,
                aes(x = log2(sowndiv))
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "Density"
  )

p1

# check distribution of realSR
p2 <- ggplot2::ggplot(noMono,
                aes(x = meanSR)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Realized Species Richness",
    y = "Density"
  )

p2

p3 <- ggplot2::ggplot(synthDat,
                aes(x = mean_biomass)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Biomass (g)",
    y = "Density"
  )

p4 <- ggplot2::ggplot(synthDat,
                aes(x = profit_mean)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Operating Profit (€/ha)",
    y = "Density"
  )

p5 <- ggplot2::ggplot(synthDat,
                aes(x = SOC_costs_mid)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "SOC Costs (€/ha)",
    y = "Density"
  )

p6 <- ggplot2::ggplot(synthDat,
                aes(x = leach_costs_mid)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Leaching Costs (€/ha/a)",
    y = "Density"
  )

p7 <- ggplot2::ggplot(synthDat,
                aes(x = meanSR, y = insurance)
                ) +
  geom_point(
    fill = "#28a87d",
    alpha = .5
  ) +
  geom_smooth(method = "lm") +
  labs(
    x = "Mean Realized Plant Diversity",
    y = "Insurance value"
  )
p7

# show all in one grid
p1 + p2 + p3 + p4 + p5 + p6 + p7 +
  plot_layout(ncol = 3) +
  plot_annotation(title = "Densities of Sown Plant Diversity, Realized Species Richness, Biomass, Profit, SOC Costs, Leaching Costs and Total Costs") &
  theme(plot.title = element_text(size=20,hjust=0.5))

ggsave("Densities.png",
       plot = last_plot(),
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```


## Statistical Models
### sowndiv without Monocultures
```{r Sown Diversity WITHOUT Monocultures}
# Profit against species richness
fitProfitSown <- brm(
  profit_mean ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono) 

    # check model assumptions 
    pp_check(fitProfitSown, type="scatter_avg")
    pp_check(fitProfitSown, ndraws=50)
    check_model(fitProfitSown, check="linearity")
    check_model(fitProfitSown, check="homogeneity")
    plot(fitProfitSown)
    
    # check output of model
    summary(fitProfitSown)

# SOC against species richness
fitSOCsown <- brm(
  SOC_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono)

    # check model assumptions
    pp_check(fitSOCsown, type="scatter_avg") 
    pp_check(fitSOCsown, ndraws=50)
    check_model(fitSOCsown, check="linearity") 
    check_model(fitSOCsown, check="homogeneity")
    plot(fitSOCsown)
    
    # check output of model
    summary(fitSOCsown)

# Leaching costs against species richness
fitLeachSown <- brm(
  leach_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono)

    # check model assumptions
    pp_check(fitLeachSown, type="scatter_avg") 
    pp_check(fitLeachSown, ndraws=50)
    check_model(fitLeachSown, check="linearity") 
    check_model(fitLeachSown, check="homogeneity")
    plot(fitLeachSown)
    
    # check output of model
    summary(fitLeachSown)

# Insurance value against sowndiv
    fitInsuranceSown <- brm(
      insurance ~ log2(sowndiv) + (1|block),
      control = list(adapt_delta = 0.95),
      iter = 3000,
      noMono
    )
    
    # check model assumptions
    pp_check(fitInsuranceSown, type="scatter_avg") 
    pp_check(fitInsuranceSown, ndraws=50)
    check_model(fitInsuranceSown, check="linearity") 
    check_model(fitInsuranceSown, check="homogeneity")
    plot(fitInsuranceSown)
    
    # check output of model
    summary(fitInsuranceSown)
    
# Total Costs against sowndiv
fitTotalSown <- brm(
  costs_total_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  noMono)

    # check model assumptions
    pp_check(fitTotalSown, type="scatter_avg") 
    pp_check(fitTotalSown, ndraws=50)
    check_model(fitTotalSown, check="linearity") 
    check_model(fitTotalSown, check="homogeneity")
    plot(fitTotalSown)
    
    # check output of model
    summary(fitTotalSown)

```

### Realized Diversity without Monocultures
```{r Realized Diversity without Monocultures}
# Profit against species richness
fitProfitReal <- brm(
  profit_mean ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono) 

    # check model assumptions 
    pp_check(fitProfitReal, type="scatter_avg")
    pp_check(fitProfitReal, ndraws=50)
    check_model(fitProfitReal, check="linearity")
    check_model(fitProfitReal, check="homogeneity")
    plot(fitProfitReal)
    
    # check output of model
    summary(fitProfitReal)

# SOC against species richness
fitSOCreal <- brm(
  SOC_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono)

    # check model assumptions
    pp_check(fitSOCreal, type="scatter_avg") 
    pp_check(fitSOCreal, ndraws=50)
    check_model(fitSOCreal, check="linearity") 
    check_model(fitSOCreal, check="homogeneity")
    plot(fitSOCreal)
    
    # check output of model
    summary(fitSOCreal)

# Leaching costs against species richness
    fitLeachReal <- brm(
      leach_costs_mid ~ log2(meanSR) + (1|block),
      control = list(adapt_delta = 0.95),
      iter = 3000,
      noMono)

    # check model assumptions
    pp_check(fitLeachReal, type="scatter_avg") 
    pp_check(fitLeachReal, ndraws=50)
    check_model(fitLeachReal, check="linearity") 
    check_model(fitLeachReal, check="homogeneity")
    plot(fitLeachReal)
    
    # check output of model
    summary(fitLeachReal)
    
# Insurance value against realized species richness
    fitInsuranceReal <- brm(
      insurance ~ log2(meanSR) + (1|block),
      control = list(adapt_delta = 0.95),
      iter = 3000,
      noMono
    )
    
    # check model assumptions
    pp_check(fitInsuranceReal, type="scatter_avg") 
    pp_check(fitInsuranceReal, ndraws=50)
    check_model(fitInsuranceReal, check="linearity") 
    check_model(fitInsuranceReal, check="homogeneity")
    plot(fitInsuranceReal)
    
    # check output of model
    summary(fitInsuranceReal)    

# Total Costs against realized species richness
fitTotalReal <- brm(
  costs_total_mid ~ log2(meanSR) + (1|block),
  noMono)

    # check model assumptions
    pp_check(fitTotalReal, type="scatter_avg") 
    pp_check(fitTotalReal, ndraws=50)
    check_model(fitTotalReal, check="linearity") 
    check_model(fitTotalReal, check="homogeneity")
    plot(fitTotalReal)
    
    # check output of model
    summary(fitTotalReal)

    
```

## Plotting
### Plot realized species diversity effects
```{r Plotting}
#Profit data without monocultures and realdiv
c_real_Profit <- conditional_effects(fitProfitReal)
plot_real_Profit_data <- c_real_Profit[[1]]
  
p_real_Profit <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = profit_mean), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_Profit_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_Profit_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Mean operating profit (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(50, 100, 150, 200, 250))
  

p_real_Profit

ggsave("Profit_real_noMono.png",
       plot = p_real_Profit,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


#SOC costs without monocultures and mean realized diversity
c_real_soc <- conditional_effects(fitSOCreal)
plot_real_soc_data <- c_real_soc[[1]]
  
p_real_soc <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = SOC_costs_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_soc_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_soc_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Social benefits of SOC sequestration (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_real_soc

ggsave("SOC_real_noMono.png",
       plot = p_real_soc,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Leaching costs without monocultures and mean realized diversity
c_real_leach <- conditional_effects(fitLeachReal)
plot_real_leach_data <- c_real_leach[[1]]
  
p_real_leach <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = leach_costs_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_leach_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_leach_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Social costs of nitrogen leaching (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12))

p_real_leach

ggsave("Leach_real_noMono.png",
       plot = p_real_leach,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Insurance value without monocultures and mean realized diversity
c_real_insurance <- conditional_effects(fitInsuranceReal)
plot_real_insurance_data <- c_real_insurance[[1]]
  
p_real_insurance <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = insurance), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_insurance_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_insurance_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Insurance value (€)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12))

p_real_insurance

ggsave("Insurance_real_noMono.png",
       plot = p_real_insurance,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

# combine the plots for realized SR in one grid
p_services_real <- p_real_Profit + p_real_soc + p_real_leach + p_real_insurance +
  plot_annotation(title = "Profit, SOC benefit, Leaching Costs, and Insurance value plotted against realized plant diversity") &
  theme(plot.title = element_text(size=13,hjust=0.5))

ggsave("Single_Services_real.png",
       plot = last_plot(),
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#economic MF without monocultures and mean realized diversity
c_real_total <- conditional_effects(fitTotalReal)
plot_real_total_data <- c_real_total[[1]]

c_real_total_lo <- conditional_effects(fitTotalLo)
plot_real_total_data_lo <- c_real_total_lo[[1]]

c_real_total_hi <- conditional_effects(fitTotalHi)
plot_real_total_data_hi <- c_real_total_hi[[1]]
  
p_real_total <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = costs_total_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_total_data, aes(x = meanSR, y = estimate__), color = "blue", size = 1, alpha = .6) +
  geom_line(data = plot_real_total_data_lo, aes(x = meanSR, y = estimate__), color = "green", linewidth = .8, alpha = .6) +
  geom_line(data = plot_real_total_data_hi, aes(x = meanSR, y = estimate__), color = "yellow", linewidth = .8, alpha = .6) +
  
  # add line at y = 0
  geom_hline(yintercept = 0, color = "grey90", linewidth = 0.8, linetype = 1, alpha = .5) +

  # Credible interval ribbon

#    geom_ribbon(data = plot_real_total_data_lo, aes(x = meanSR, ymin = lower__, ymax = upper__), 
#              fill = "green", alpha = 0.2) +
#      geom_ribbon(data = plot_real_total_data_hi, aes(x = meanSR, ymin = lower__, ymax = upper__), 
#              fill = "yellow", alpha = 0.2) +
    geom_ribbon(data = plot_real_total_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.4) +
  labs(x = "Mean realized plant diversity", y = "Economic multifunctionality (€/ha/a)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_real_total

ggsave("total_real_noMono.png",
       plot = p_real_total,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

### Plot sown diversity effect
```{r}
#Profit data without monocultures and sowndiv
c_sown_Profit <- conditional_effects(fitProfitSown)
plot_sown_Profit_data <- c_sown_Profit[[1]]
  
p_sown_Profit <- ggplot() +
  geom_point(data = noMono, aes(x = sowndiv, y = profit_mean), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_sown_Profit_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_sown_Profit_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "Mean Profit (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 8, 16))

p_sown_Profit

ggsave("Profit_sown_noMono.png",
       plot = p_sown_Profit,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


#SOC costs without monocultures and mean sownized diversity
c_sown_soc <- conditional_effects(fitSOCsown)
plot_sown_soc_data <- c_sown_soc[[1]]
  
p_sown_soc <- ggplot() +
  geom_point(data = noMono, aes(x = sowndiv, y = SOC_costs_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_sown_soc_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_sown_soc_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "Costs/benefits due to SOC loss/increase (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 8, 16)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_sown_soc

ggsave("SOC_sown_noMono.png",
       plot = p_sown_soc,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Leaching costs without monocultures and mean sownized diversity
c_sown_leach <- conditional_effects(fitLeachSown)
plot_sown_leach_data <- c_sown_leach[[1]]
  
p_sown_leach <- ggplot() +
  geom_point(data = noMono, aes(x = sowndiv, y = leach_costs_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_sown_leach_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_sown_leach_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "Social costs of nitrogen leaching") +
  scale_x_continuous(breaks = c(2, 4, 8, 16))

p_sown_leach

ggsave("Leach_sown_noMono.png",
       plot = p_sown_leach,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


#economic MF without monocultures and mean sownized diversity
c_sown_total <- conditional_effects(fitTotalSown)
plot_sown_total_data <- c_sown_total[[1]]
  
p_sown_total <- ggplot() +
  geom_point(data = noMono, aes(x = sowndiv, y = costs_total_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_sown_total_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +
  
  # add line at y = 0
  geom_hline(yintercept = 0, color = "grey90", linewidth = 0.8, linetype = 1, alpha = .5) +

  # Credible interval ribbon
  geom_ribbon(data = plot_sown_total_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "Economic multifunctionality (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 8, 16)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_sown_total

ggsave("total_sown_noMono.png",
       plot = p_sown_total,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

## Supplementary Analysis
### Linear models with monocultures
#### &Sown diversity
```{r Sown Diversity WITH Monocultures, include=FALSE}
# Profit against sown plant diversity
fitProfit <- brm(
  profit_mean ~ log2(sowndiv),
  synthDat) 

    # check model assumptions 
    pp_check(fitProfit, type="scatter_avg")
    pp_check(fitProfit, ndraws=50)
    check_model(fitProfit, check="linearity")
    check_model(fitProfit, check="homogeneity")
    plot(fitProfit)
    
    # check output of model
    summary(fitProfit)

# SOC against sown plant diversity
fitSOC <- brm(
  SOC_costs_mid ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitSOC, type="scatter_avg") 
    pp_check(fitSOC, ndraws=50)
    check_model(fitSOC, check="linearity") 
    check_model(fitSOC, check="homogeneity")
    plot(fitSOC)
    
    # check output of model
    summary(fitSOC)

# Leaching costs against sown plant diversity
fitLeach <- brm(
  leach_costs_mid ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitLeach, type="scatter_avg") 
    pp_check(fitLeach, ndraws=50)
    check_model(fitLeach, check="linearity") 
    check_model(fitLeach, check="homogeneity")
    plot(fitLeach)
    
    # check output of model
    summary(fitLeach)

# Insurance value against sown plant diversity
fitInsurance <- brm(
  insurance ~log2(sowndiv),
  synthDat
)    
    
    # check model assumptions
    pp_check(fitInsurance, type="scatter_avg") 
    pp_check(fitInsurance, ndraws=50)
    check_model(fitInsurance, check="linearity") 
    check_model(fitInsurance, check="homogeneity")
    plot(fitInsurance)
    
    #check model output
    summary(fitInsurance)
    
# Total Costs against sown plant diversity
fitTotal <- brm(
  costs_total_mid ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitTotal, type="scatter_avg") 
    pp_check(fitTotal, ndraws=50)
    check_model(fitTotal, check="linearity") 
    check_model(fitTotal, check="homogeneity")
    plot(fitTotal)
    
    # check output of model
    summary(fitTotal)
    
```

#### &Realized diversity
```{r Realized Diversity With Monocultures, include=FALSE}
# Profit against species richness - including monocultures
fitProfitRealAll <- brm(
  profit_mean ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat) 

    # check model assumptions 
    pp_check(fitProfitRealAll, type="scatter_avg")
    pp_check(fitProfitRealAll, ndraws=50)
    check_model(fitProfitRealAll, check="linearity")
    check_model(fitProfitRealAll, check="homogeneity")
    plot(fitProfitRealAll)
    
    # check output of model
    summary(fitProfitRealAll)

# SOC against species richness - including monocultures
fitSOCrealAll <- brm(
  SOC_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitSOCrealAll, type="scatter_avg") 
    pp_check(fitSOCrealAll, ndraws=50)
    check_model(fitSOCrealAll, check="linearity") 
    check_model(fitSOCrealAll, check="homogeneity")
    plot(fitSOCrealAll)
    
    # check output of model
    summary(fitSOCrealAll)

# Leaching costs against species richness - including monocultures
fitLeachRealAll <- brm(
  leach_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitLeachRealAll, type="scatter_avg") 
    pp_check(fitLeachRealAll, ndraws=50)
    check_model(fitLeachRealAll, check="linearity") 
    check_model(fitLeachRealAll, check="homogeneity")
    plot(fitLeachRealAll)
    
    # check output of model
    summary(fitLeachRealAll)

# Insurance against species richness - including monocultures
fitInsuranceRealAll <- brm(
  insurance ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitInsuranceRealAll, type="scatter_avg") 
    pp_check(fitInsuranceRealAll, ndraws=50)
    check_model(fitInsuranceRealAll, check="linearity") 
    check_model(fitInsuranceRealAll, check="homogeneity")
    plot(fitInsuranceRealAll)
    
    # check output of model
    summary(fitInsuranceRealAll)        

# Total Costs against mean realized species richness - including monocultures
fitTotalRealAll <- brm(
  costs_total_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitTotalRealAll, type="scatter_avg") 
    pp_check(fitTotalRealAll, ndraws=50)
    check_model(fitTotalRealAll, check="linearity") 
    check_model(fitTotalRealAll, check="homogeneity")
    plot(fitTotalRealAll)
    
    # check output of model
    summary(fitTotalRealAll)


```


#### Plot realized species diversity effects including monocultures
```{r Plotting}
#Profit data without monocultures and realdiv
c_real_profit_all <- conditional_effects(fitProfitRealAll)
plot_real_profit_all_data <- c_real_profit_all[[1]]
  
p_real_profit_all <- ggplot() +
  geom_point(data = synthDat, aes(x = meanSR, y = profit_mean), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_profit_all_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_profit_all_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Mean operating profit (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(50, 100, 150, 200, 250))
  

p_real_profit_all

ggsave("Profit_real_all.png",
       plot = p_real_profit_all,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


#SOC costs without monocultures and mean realized diversity
c_real_soc_all <- conditional_effects(fitSOCrealAll)
plot_real_soc_all_data <- c_real_soc_all[[1]]
  
p_real_soc_all <- ggplot() +
  geom_point(data = synthDat, aes(x = meanSR, y = SOC_costs_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_soc_all_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_soc_all_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Social benefits of SOC sequestration (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_real_soc_all

ggsave("SOC_real_all.png",
       plot = p_real_soc_all,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Leaching costs without monocultures and mean realized diversity
c_real_leach_all <- conditional_effects(fitLeachRealAll)
plot_real_leach_all_data <- c_real_leach_all[[1]]
  
p_real_leach_all <- ggplot() +
  geom_point(data = synthDat, aes(x = meanSR, y = leach_costs_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_leach_all_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_leach_all_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Social costs of nitrogen leaching (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12))

p_real_leach_all

ggsave("Leach_real_all.png",
       plot = p_real_leach_all,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Insurance value without monocultures and mean realized diversity
c_real_insurance_all <- conditional_effects(fitInsuranceRealAll)
plot_real_insurance_all_data <- c_real_insurance_all[[1]]
  
p_real_insurance_all <- ggplot() +
  geom_point(data = synthDat, aes(x = meanSR, y = insurance), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_insurance_all_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_insurance_all_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Insurance value (€)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12))

p_real_insurance_all

ggsave("Insurance_real_all.png",
       plot = p_real_insurance_all,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

# combine the plots for realized SR in one grid
p_services_real <- p_real_profit_all + p_real_soc_all + p_real_leach_all + p_real_insurance_all +
  plot_annotation(title = "Profit, SOC benefit, Leaching Costs, and Insurance value plotted against realized plant diversity") &
  theme(plot.title = element_text(size=13,hjust=0.5))

ggsave("Single_Services_real.png",
       plot = last_plot(),
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#economic MF without monocultures and mean realized diversity
c_real_total <- conditional_effects(fitTotalReal)
plot_real_total_data <- c_real_total[[1]]

c_real_total_lo <- conditional_effects(fitTotalLo)
plot_real_total_data_lo <- c_real_total_lo[[1]]

c_real_total_hi <- conditional_effects(fitTotalHi)
plot_real_total_data_hi <- c_real_total_hi[[1]]
  
p_real_total <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = costs_total_mid), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_real_total_data, aes(x = meanSR, y = estimate__), color = "blue", size = 1, alpha = .6) +
  geom_line(data = plot_real_total_data_lo, aes(x = meanSR, y = estimate__), color = "green", linewidth = .8, alpha = .6) +
  geom_line(data = plot_real_total_data_hi, aes(x = meanSR, y = estimate__), color = "yellow", linewidth = .8, alpha = .6) +
  
  # add line at y = 0
  geom_hline(yintercept = 0, color = "grey90", linewidth = 0.8, linetype = 1, alpha = .5) +

  # Credible interval ribbon

#    geom_ribbon(data = plot_real_total_data_lo, aes(x = meanSR, ymin = lower__, ymax = upper__), 
#              fill = "green", alpha = 0.2) +
#      geom_ribbon(data = plot_real_total_data_hi, aes(x = meanSR, ymin = lower__, ymax = upper__), 
#              fill = "yellow", alpha = 0.2) +
    geom_ribbon(data = plot_real_total_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.4) +
  labs(x = "Mean realized plant diversity", y = "Economic multifunctionality (€/ha/a)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_real_total

ggsave("total_real_noMono.png",
       plot = p_real_total,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```



### Sensitivity Analysis with different price scenarios
```{r Sensitivity Analysis}
# Low Total Costs against species richness
fitTotalLo <- brm(
  costs_total_lo ~ log2(meanSR) + (1|block),
  noMono)

    # check model assumptions
    pp_check(fitTotalLo, type="scatter_avg") 
    pp_check(fitTotalLo, ndraws=50)
    check_model(fitTotalLo, check="linearity") 
    check_model(fitTotalLo, check="homogeneity")
    plot(fitTotalLo)
    
    # check output of model
    summary(fitTotalLo)

# High Total Costs against species richness
fitTotalHi <- brm(
  costs_total_hi ~ log2(meanSR) + (1|block),
  noMono)

    # check model assumptions
    pp_check(fitTotalHi, type="scatter_avg") 
    pp_check(fitTotalHi, ndraws=50)
    check_model(fitTotalHi, check="linearity") 
    check_model(fitTotalHi, check="homogeneity")
    plot(fitTotal4)
    
    # check output of model
    summary(fitTotalHi)
    
```

### Plot sensitivity analysis
```{r}

```

