---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output:
  html_document: default
  pdf_document: default
---

# Preface
This R Script is intended to be used to load, explore, and synthesize the datasets for the following metrics measured in the Jena main experiment: Aboveground biomass production, C&N content in the soil, Realized Species Richness, climate variability, leaching N (nitrogen leaching), pollination, flower diversity, total flowering period per plot, above- and belowground herbivory (for belowground measured as energy fluxes in the food web), soil organic carbon (SOC), leaf dry matter content (LDMC), soil diversity, microbial biomass + respiration, 
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE, fig.width = 8, fig.height = 6, dpi = 300, cache = TRUE, cache.lazy = FALSE, cache.path = "cache/", tidy = TRUE, comment = NA, collapse = TRUE)
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
# load data on plot design (needed for 'block code')
design <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Plot_Design/90_20_data.xlsx")

#load data on realized species richness (ID = 487)
SR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx") 
  #change data type for realized SR & sowndiv
  SR$realizedSR <- as.numeric(SR$realizedSR)
  SR$sowndiv <- as.numeric(SR$sowndiv)

#load data on aboveground biomass production (ID = 624) 
biomassDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/624_9_data.xlsx")  %>%  # remove sowndiv = 60 plots 
  filter(!plotcode %in% c("B4A01", "B3A14", "B2A03", "B1A22")) %>% 
  filter(year > 2002)
  biomassDat$plotcode <- as.factor(biomassDat$plotcode)

  # load data on SGMs 
SGMS <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/SGMs.xlsx")
  SGMS$year <- as.factor(SGMS$year) # change year to a factor

#load data on soil organic carbon (SOC) (2020: ID = 424 ). 2020 data is the most recent year that includes bulk density measurements.
carbon20 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/424_7_data.xlsx")

# load 2002 SOC data as a baseline
carbon02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/278_2_data.xlsx") %>% 
  filter(year == 2002)
  
  # change data types
  carbon02$Corg_Soil <- as.numeric(carbon02$Corg_Soil)
  carbon02$BulkDensity_Soil <- as.numeric(carbon02$BulkDensity_Soil)

#load data on nitrogen leaching (this should be in kg/ha per year)
leaching <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/Yearly_Leaching_Plot.xlsx") %>% 
  rename_at('plot', ~'plotcode') %>% # rename plot to plotcode
  mutate(plotcode = as.factor(plotcode)) %>% 
  filter(!plotcode %in% c("B4A01", "B3A14", "B2A03", "B1A22"))  
```


## Data wrangling & calculations
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated.
Monetary values for individual ecosystem services will be calculated.
```{r Biomass data}
# calculate biomass for each year per plot and average across subsamples
biomassAnnual <- biomassDat %>% 
  group_by(plotcode, year, subsample) %>%
  summarise(sum_biomass = sum(biomass, na.rm = TRUE), .groups = "drop") %>% 
  group_by(plotcode, year) %>%
  summarise(biomass_mean = mean(sum_biomass, na.rm = TRUE)* 0.01 , .groups = "drop") # convert to ton per hectare (*0.01)

biomassAnnual$year <- as.factor(biomassAnnual$year) 

# merge annual biomass data with the SGMs dataset
biomassAnnual <- biomassAnnual %>% 
  left_join(SGMS, by = join_by(year)) %>% 
  select(plotcode, year, biomass_mean, price, costs_ha, costs_ton)  # select only relevant columns
  
# calculate yearly operating profit (€/ha/a)
biomassAnnual$profit_annual <- biomassAnnual$biomass_mean * (biomassAnnual$price - biomassAnnual$costs_ton)

# calculate the biomass stability as the coefficient of variance of the biomass across years (CV = SD/mean)
biomassAnnual <- biomassAnnual %>% 
  group_by(plotcode) %>% 
  mutate(CV = (sd(biomass_mean)/mean(biomass_mean))) 

# sum up biomass for each plot and calculate the mean biomass and mean profit across years. (calculating "mean of CV" doesn't change it's value on plot level, because the CV was already the same for every year per plot)
biomassMeans <- biomassAnnual %>% 
  group_by(plotcode) %>% 
  summarise(
    biomass_mean = mean(biomass_mean, na.rm = TRUE),
    profit_mean = median(profit_annual, na.rm = TRUE),
    CV = mean(CV, na.rm = TRUE),
    .groups = "drop"
  )

# calculate certainty equivalent (CE) and risk premium (RP) with profit, CV and risk aversion coefficient (taken from Rommel & Sagebiel et al., 2023)
biomassMeans$CE <- biomassMeans$profit_mean * (1+(biomassMeans$CV)^2)^(-0.6)
biomassMeans$RP <- biomassMeans$profit_mean - biomassMeans$CE

```


```{r soil carbon}
# rename plot to plotcode and change data types 
carbon02 <- carbon02 %>%
  rename_at('plot', ~'plotcode')

carbon20 <- carbon20 %>%
  rename_at('plot', ~'plotcode')

# create a new column with mean C across layers per plot for 2004 & 2006
carbon02_clean <- carbon02 %>% group_by(plotcode) %>% 
                    mutate(mean_org_02 = mean(Corg_Soil), mean_bulk = mean(BulkDensity_Soil)) %>%
                      distinct(plotcode, mean_org_02, mean_bulk)

# create a new column with mean C across layers per plot for 2017 & 2020
carbon20_clean <- carbon20 %>% group_by(plotcode) %>% 
                    mutate(mean_org_20 = mean(Corg_Soil), mean_bulk = mean(BulkDensity_Soil)) %>%
                      distinct(plotcode, mean_org_20, mean_bulk)


# create a column with mid SOC value for beginning and end
carbon02_clean$SOC_costs_mid_02 <- carbon02_clean$mean_org_02 * carbon02_clean$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 300 
carbon20_clean$SOC_costs_mid_20 <- carbon20_clean$mean_org_20 * carbon20_clean$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 300 
```

```{r Species Richness data}
# calculate the mean realized SR for the years 2003-2022
realdiv <- SR %>% 
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)

SR <- distinct(SR, plotcode, sowndiv) %>% mutate(sowndiv = as.numeric(sowndiv))
```


```{r leaching}
# calculate the mean annual leaching per plot across years
leachingMean <- leaching %>% 
            summarise(leaching_mean = mean(leach_y, na.rm = TRUE)/100, .by = plotcode)

# calculate the costs in €/ha/a for each plot
leachingMean$leach_costs_lo <- leachingMean$leaching_mean * 5 *(-1)
leachingMean$leach_costs_mid <- leachingMean$leaching_mean * 11.23 *(-1)
leachingMean$leach_costs_hi <- leachingMean$leaching_mean * 20 *(-1)
```

## Synthesize in one Dataset
```{r Data Synthesis}
synthDat <- biomassMeans %>% # merge the biomass and the realized SR data
  left_join(SR, by = join_by(plotcode)) %>%
  left_join(carbon02_clean, by = join_by(plotcode)) %>% 
  left_join(carbon20_clean, by = join_by(plotcode)) %>%
  left_join(select(design, plotcode, block)) %>% 
  left_join(leachingMean, by = join_by(plotcode))  %>%
  left_join(realdiv, by = join_by(plotcode)) 
  

synthDat$block <- as.factor(synthDat$block) # change block to a factor
synthDat$plotcode <- as.factor(synthDat$plotcode)  # change plotcode to a factor


  # define quadratic relationship between biodiversity and insurance value
  quad <- lm(CV ~ sowndiv + I(sowndiv^2), synthDat)

  # calculate dCV/dSR
  dCVdSR <- function(SR) {
  c <- coef(quad)
  c[2] + 2 * c[3] * SR
}
  
  # insurance value like in Martin's code
  synthDat <- synthDat %>% 
    mutate(dCVdSR = dCVdSR(sowndiv),
         insurance = sowndiv * dCVdSR * (-1))
 
# create a column with social costs of carbon (for different pricing scenarios)
synthDat$SOC_costs_mid <- (synthDat$SOC_costs_mid_20 - synthDat$SOC_costs_mid_02) /17
synthDat$SOC_costs_lo <- (synthDat$SOC_costs_mid_20 - synthDat$SOC_costs_mid_02) / 17 / 4 # price is only 75€ per ton which is 25% of 300
synthDat$SOC_costs_hi <- (synthDat$SOC_costs_mid_20 - synthDat$SOC_costs_mid_02) / 17 / 0.375 # price is 800€ per ton 

# add up all the costs for total value
synthDat$costs_total_mid <- synthDat$SOC_costs_mid + synthDat$leach_costs_mid + synthDat$profit_mean # + synthDat$insurance

# save the synthesized dataset
writexl::write_xlsx(synthDat, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF_final.xlsx")
writexl::write_xlsx(carbon02_clean, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/carbon02_dom.xlsx")

```