---
title: "Ecological Multifunctinality Analysis"
author: "Marius Munschek"
date: "2025-05-30"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the ecological multifunctionality for (mostly) soil functions in the Jena experiment. I will focus on the following functions:
- Biomass production above and below
- Nitrogen leaching
- Carbon sequestration 
- 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation
## Loading packages
```{r Packages, include=FALSE}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
```

## Set up a theme
```{r include=FALSE}
theme_set(theme_minimal(base_size = 14)+
            theme(plot.title = element_text(size=13,hjust=0.5),
                  axis.text = element_text(size = 7),
                  axis.text.y = element_text(size = 13),
                  axis.text.x = element_text(size = 13)))
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data, include=FALSE}
# load synthesis dataset
ecologicalMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/ecological_multifunctionality.xlsx")

```


### Multidiversity Index
Here I want to integrate the Diversity metrics
```{r Multidiversity}

```

## Statistical Models 
Also single functions or just multifunctionality?
```{r Biomass ~ SR}


```


```{r SOC ~ realized SR}

```

```{r N leaching ~ realized SR}

```

```{r HerbivoryAbove ~ Realized SR}

```


## Plotting
### Plot biomass against realized species richness
```{r Plotting}
p1 

```

