---
title: "Economic Multifunctionality Analysis"
author: "Marius Munschek"
date: "2025-05-28"
output: html_document
---

#Preface
This R Script is written by Marius Munschek to calculate the economic multifunctionality for ecosystem services in the Jena Experiment. The performance of the following ecosystem sevices per plot is averaged over the years 2011 - 2020, then translated into monetary values which are aggregated to obtain economic multifunctionality values per plot:
- Biomass production
- Nitrogen leaching
- Carbon sequestration

```{r setup, include=FALSE}
library(ragg)
library(here)

knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = TRUE,
  cache = TRUE, cache.lazy = FALSE,
  cache.path = "cache/",
  out.width = "100%", out.height = "100%",
  dev = "ragg_png"
)

# Make sure cache/figures/output exist (prevents silent failures on new machines)
dir.create(here("cache"), showWarnings = FALSE, recursive = TRUE)
dir.create(here("figures"), showWarnings = FALSE, recursive = TRUE)

# Source project functions from the repo (portable)
source(here("functions", "myfunctions.R"))
source(here("functions", "fit_or_load.R"))
stopifnot(exists("theme_custom"))

# Define a portable output directory for ggsave()
out_dir <- here("3_Graphs")
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(stringr)
library(tidyr)

```

## Load datasets - Filter for relevant observations/measurements
```{r Read Data}
#load synthesis dataset
synthDat <- read_xlsx(
  here("1_Data", "economic_MF_final.xlsx")
) %>%
  filter(sowndiv < 60) %>%
  mutate(block = as.factor(block))

synthDat$plotcode <- as.factor(synthDat$plotcode) #change the plotcode to a factor

```

## Statistical Models
### Linear models with monocultures
#### &Sown diversity
```{r Sown Diversity WITH Monocultures, include=FALSE}
# Profit against sown plant diversity
fitProfit <- fit_or_load("fitProfit", function() {
  brm(
  profit_mean ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat) 
})

    # check model assumptions 
    pp_check(fitProfit, type="scatter_avg")
    pp_check(fitProfit, ndraws=50)
    check_model(fitProfit, check="linearity")
    check_model(fitProfit, check="homogeneity")
    plot(fitProfit)
    
    # check output of model
    summary(fitProfit)

# SOC against sown plant diversity
fitSOC <- fit_or_load("fitSOC", function() {
  brm(
  SOC_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitSOC, type="scatter_avg") 
    pp_check(fitSOC, ndraws=50)
    check_model(fitSOC, check="linearity") 
    check_model(fitSOC, check="homogeneity")
    plot(fitSOC)
    
    # check output of model
    summary(fitSOC)

# Leaching costs against sown plant diversity
fitLeach <- fit_or_load("fitLeach", function() {
  brm(
  leach_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitLeach, type="scatter_avg") 
    pp_check(fitLeach, ndraws=50)
    check_model(fitLeach, check="linearity") 
    check_model(fitLeach, check="homogeneity")
    plot(fitLeach)
    
    # check output of model
    summary(fitLeach)

    
# Total Costs (medium) against sown plant diversity
fitTotal <- fit_or_load("fitTotal", function() {
  brm(
  costs_total_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitTotal, type="scatter_avg") 
    pp_check(fitTotal, ndraws=50)
    check_model(fitTotal, check="linearity") 
    check_model(fitTotal, check="homogeneity")
    plot(fitTotal)
    
    # check output of model
    summary(fitTotal)


# define model for RP ~ sowndiv
fitRP <- fit_or_load("fitRP", function() {
  brm(
    RP ~ log2(sowndiv) + (1|block),
             control = list(adapt_delta = 0.95),
             iter = 3000,
             data = synthDat)
})

    # check model assumptions
    pp_check(fitRP, type="scatter_avg") 
    pp_check(fitRP, ndraws=50)
    check_model(fitRP, check="linearity") 
    check_model(fitRP, check="homogeneity")
    plot(fitRP)
    
    # check output of model
    summary(fitRP)

# calculate effect size of insurance
post <- as_draws_df(fitRP)

insurance_draws <- -post$b_log2sowndiv

insurance_summary <- tibble(
  mean = mean(insurance_draws),
  lwr  = quantile(insurance_draws, 0.025),
  upr  = quantile(insurance_draws, 0.975)
)

insurance_summary
    
    
```

#### &Realized diversity
```{r Realized Diversity WITH Monocultures, include=FALSE}
# Profit against species richness - including monocultures
fitProfitReal <- fit_or_load("fitProfitReal", function() {
  brm(
  profit_mean ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat) 
})
    # check model assumptions 
    pp_check(fitProfitReal, type="scatter_avg")
    pp_check(fitProfitReal, ndraws=50)
    check_model(fitProfitReal, check="linearity")
    check_model(fitProfitReal, check="homogeneity")
    plot(fitProfitReal)
    
    # check output of model
    summary(fitProfitReal)

# SOC against species richness - including monocultures
fitSOCreal <- fit_or_load("fitSOCreal", function() {
  brm(
  SOC_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitSOCreal, type="scatter_avg") 
    pp_check(fitSOCreal, ndraws=50)
    check_model(fitSOCreal, check="linearity") 
    check_model(fitSOCreal, check="homogeneity")
    plot(fitSOCreal)
    
    # check output of model
    summary(fitSOCreal)

# Leaching costs against species richness - including monocultures
fitLeachReal <- fit_or_load("fitLeachReal", function() {
  brm(
  leach_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitLeachReal, type="scatter_avg") 
    pp_check(fitLeachReal, ndraws=50)
    check_model(fitLeachReal, check="linearity") 
    check_model(fitLeachReal, check="homogeneity")
    plot(fitLeachReal)
    
    # check output of model
    summary(fitLeachReal)

# Risk premium against realized diversity
fitRPreal <- fit_or_load("fitRPreal", function() {
  brm(RP ~ log2(sowndiv) + (1|block),
             control = list(adapt_delta = 0.95),
             iter = 3000,
             data = synthDat)
})

    # check model assumptions
    pp_check(fitRPreal, type="scatter_avg") 
    pp_check(fitRPreal, ndraws=50)
    check_model(fitRPreal, check="linearity") 
    check_model(fitRPreal, check="homogeneity")
    plot(fitRPreal)
    
    # check output of model
    summary(fitRPreal)

# calculate effect size of insurance
post_real <- as_draws_df(fitRPreal)

insurance_draws_real <- -post_real$b_log2sowndiv

insurance_summary_real <- tibble(
  mean = mean(insurance_draws_real),
  lwr  = quantile(insurance_draws_real, 0.025),
  upr  = quantile(insurance_draws_real, 0.975)
)

insurance_summary       

# Total Costs against mean realized species richness
fitTotalReal <- fit_or_load("fitTotalReal", function() {
  brm(
  costs_total_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitTotalReal, type="scatter_avg") 
    pp_check(fitTotalReal, ndraws=50)
    check_model(fitTotalReal, check="linearity") 
    check_model(fitTotalReal, check="homogeneity")
    plot(fitTotalReal)
    
    # check output of model
    summary(fitTotalReal)


```

## Plotting
```{r Plot sown diversity  - profit}
#Profit data with monocultures and sowndiv
c_sown_Profit <- conditional_effects(fitProfit)
plot_sown_Profit_data <- c_sown_Profit[[1]]
  
p_sown_Profit <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = profit_mean),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data = plot_sown_Profit_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval
  geom_ribbon(
    data = plot_sown_Profit_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant species richness",
    y = expression(
  "Profit (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)
  ) +
  theme_custom() +
  theme(
    aspect.ratio = 0.6   # width:height ~ 1 : 0.6 (tweak to taste)
  )

p_sown_Profit

ggsave(
  filename = "Fig_sown_profit.png",
  plot     = p_sown_Profit,
  width    = 8,     # in inches; adjust to Nature’s figure width
  height   = 5,
  units    = "in",
  device   = "png",
  dpi      = 300,
  path     = out_dir
)
```

```{r Plot sown diversity - SOC}
# SOC costs with monocultures and mean sown diversity
c_sown_soc <- conditional_effects(fitSOC)
plot_sown_soc_data <- c_sown_soc[[1]]

p_sown_soc <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = SOC_costs_mid),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data   = plot_sown_soc_data,
    aes(x  = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_sown_soc_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1500, -1000, -500, 0, 500, 1000)
  ) +
  labs(
    x = "Sown plant species richness",
y = expression(
  atop(
    "Social value of",
    "SOC (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
  )
)
  ) +
  theme_custom()

p_sown_soc

ggsave(
  filename = "SOC_sown.png",
  plot     = p_sown_soc,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 6.27,
  height   = 5,
  units    = "in"
)

```

```{r Plot sown diversity - Leaching}
#Leaching costs and mean sown diversity
c_sown_leach <- conditional_effects(fitLeach)
plot_sown_leach_data <- c_sown_leach[[1]]
  
# Leaching costs: sown diversity
p_sown_leach <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = leach_costs_mid),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data = plot_sown_leach_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval
  geom_ribbon(
    data = plot_sown_leach_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-125, -100, -75, -50, -25)
  ) +
  labs(
    x = "Sown plant species richness",
y = expression(
  atop(
    "Social costs of",
    NO[3] * "\u2013N leaching (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
  )
)
  ) +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_sown_leach

ggsave(
  filename = "Fig_Leach_sown.png",
  plot     = p_sown_leach,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```

```{r Plot sown diversity - RP}
#Insurance value and sown diversity
c_sown_rp <- conditional_effects(fitRP)
plot_sown_rp_data <- c_sown_rp[[1]]

p_sown_rp <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = sowndiv, y = RP),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4,
    position = position_jitter(width = 0.12, height = 0)
  ) +
  geom_line(
    data     = plot_sown_rp_data,
    aes(x    = 2^log2(sowndiv), y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  geom_ribbon(
    data  = plot_sown_rp_data,
    aes(x = 2^log2(sowndiv), ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant species richness",
    y = expression(
  "Risk premium (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)

  ) +
  theme_custom() +
  theme(aspect.ratio = 0.6)

p_sown_rp

ggsave(
  filename = "Fig_sown_RP.png",
  plot     = p_sown_rp,
  width    = 8,     # in inches; adjust to Nature’s figure width
  height   = 5,
  units    = "in",
  device   = "png",
  dpi      = 300,
  path     = out_dir 
)

```


```{r Plot sown diversity - Economic MF}
#economic MF and sown diversity
c_sown_total <- conditional_effects(fitTotal)
plot_sown_total_data <- c_sown_total[[1]]
  
# Economic multifunctionality ~ sown diversity
p_sown_total <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = costs_total_mid),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data = plot_sown_total_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_sown_total_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1500, -1000, -500, 0, 500, 1000)
  ) +
  labs(
    x = "Sown plant species richness",
y = expression(
  "EMF (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)
  ) +
  theme_custom()

p_sown_total

ggsave(
  filename = "fig_total_sown.png",
  plot     = p_sown_total,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 6.27,
  height   = 5,
  units    = "in"
)


```


```{r Create multi-panel plot}

fig_sown_all <- (
  p_sown_Profit +
  p_sown_soc
) / (
  p_sown_leach +
  p_sown_rp
) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    plot.tag = element_text(
      face = "bold",
      size = 12,
      hjust = -0.2,
      vjust = 1.5
    )
  )

fig_sown_all

ggsave(
  filename = "Fig_sown_multifunctionality.png",
  plot     = fig_sown_all,
  width    = 8,
  height   = 5,
  units    = "in",
  dpi      = 300,
  device   = "png",
  path     = out_dir
)


```


### Sensitivity Analysis with different price scenarios
```{r Sensitivity Analysis}
# Low Total Costs against species richness
fitTotalLo <- fit_or_load("fitTotalLo", function() {
  brm(
  costs_total_lo ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})

    # check model assumptions
    pp_check(fitTotalLo, type="scatter_avg") 
    pp_check(fitTotalLo, ndraws=50)
    check_model(fitTotalLo, check="linearity") 
    check_model(fitTotalLo, check="homogeneity")
    plot(fitTotalLo)
    
    # check output of model
    summary(fitTotalLo)

# High Total Costs against species richness
fitTotalHi <- fit_or_load("fitTotalHi", function() {
  brm(
  costs_total_hi ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)
})
    # check model assumptions
    pp_check(fitTotalHi, type="scatter_avg") 
    pp_check(fitTotalHi, ndraws=50)
    check_model(fitTotalHi, check="linearity") 
    check_model(fitTotalHi, check="homogeneity")
    plot(fitTotalHi)
    
    # check output of model
    summary(fitTotalHi)
    
```

### Plot sensitivity analysis
```{r Plot Sown Diversity - Economic MF - Alternative Pricing}
c_sown_total <- conditional_effects(fitTotal)
plot_sown_total_data <- c_sown_total[[1]]

c_sown_total_lo <- conditional_effects(fitTotalLo)
plot_sown_total_data_lo <- c_sown_total_lo[[1]]

c_sown_total_hi <- conditional_effects(fitTotalHi)
plot_sown_total_data_hi <- c_sown_total_hi[[1]]


# Combine the three conditional_effects data frames
plot_sown_total_all <- bind_rows(
  mutate(plot_sown_total_data,    scenario = "Mid price (default)"),
  mutate(plot_sown_total_data_lo, scenario = "Low price"),
  mutate(plot_sown_total_data_hi, scenario = "High price")
)

# prepare data for combined plot
library(tidyr)

raw_sown_total_all <- synthDat %>%
  dplyr::select(sowndiv, costs_total_mid, costs_total_lo, costs_total_hi) %>%
  pivot_longer(
    cols      = c(costs_total_mid, costs_total_lo, costs_total_hi),
    names_to  = "scenario_raw",
    values_to = "economic_mf"
  ) %>%
  mutate(
    scenario = dplyr::recode(
      scenario_raw,
      "costs_total_mid" = "Mid price (default)",
      "costs_total_lo"  = "Low price",
      "costs_total_hi"  = "High price"
    )
  )

# Plot without raw data points
p_sown_total_sens <- ggplot() +
  
  # horizontal reference line at y = 0
  geom_hline(
    yintercept = 0,
    colour     = "grey80",
    linewidth  = 0.6
  ) +
  
  # CI ribbon for mid/default scenario only
  geom_ribbon(
    data = filter(plot_sown_total_all, scenario == "Mid price (default)"),
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  
  # model lines for all three price scenarios
  geom_line(
    data = plot_sown_total_all,
    aes(x = sowndiv, y = estimate__, colour = scenario, linetype = scenario),
    linewidth = 0.9
  ) +
  
  # colours and line types for scenarios
  scale_colour_manual(
    name   = "Price scenario",
    values = c(
      "Mid price (default)" = "#0072B2",  # blue
      "Low price"           = "#009E73",  # green
      "High price"          = "#E69F00"   # orange
    )
  ) +
  scale_linetype_manual(
    name   = "Price scenario",
    values = c(
      "Mid price (default)" = "solid",
      "Low price"           = "dashed",
      "High price"          = "dotted"
    )
  ) +
  
  # log2 x-axis
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  
  # y-axis
  scale_y_continuous(
    breaks = c(-1500, -1000, -500, 0, 500, 1000, 1500)
  ) +
  
  # labels
  labs(
    x = "Sown plant species richness",
    y = expression(
  "Economic multifunctionality (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)

  ) +
  
  theme_custom() +
  theme(
    legend.position  = "bottom",
    legend.direction = "horizontal",
    legend.box       = "horizontal",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 11),
    legend.spacing.x = unit(4, "pt")
  )

p_sown_total_sens

ggsave(
  filename = "total_sown_different_pricing.png",
  plot     = p_sown_total_sens,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 5.9,
  height   = 4.5,
  units    = "in"
)

```

## Supplementary Analysis
#### Plot realized species diversity effects including monocultures
```{r Plot Realized SR - Profit}
# Profit data with monocultures and realized diversity
c_real_profit_all <- conditional_effects(fitProfitReal)
plot_real_profit_all_data <- c_real_profit_all[[1]]

p_real_profit_all <- ggplot() +
  # raw data
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = profit_mean),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data     = plot_real_profit_all_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data  = plot_real_profit_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  
  # --- LOG2 AXIS ---
  scale_x_continuous(
  expand = expansion(mult = c(0.02, 0.02))
  ) + 

  # labels
  labs(
    x = "Realized plant species richness",
    y = expression(atop(
        "Profit (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)
)
  ) +
  
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_real_profit_all

ggsave(
  filename = "Profit_real_all.png",
  plot     = p_real_profit_all,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```


```{r Plot Realized SR - SOC}
# SOC benefits without monocultures and realised diversity
c_real_soc_all <- conditional_effects(fitSOCreal)
plot_real_soc_all_data <- c_real_soc_all[[1]]

p_real_soc_all <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = SOC_costs_mid),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  geom_line(
    data     = plot_real_soc_all_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  geom_ribbon(
    data  = plot_real_soc_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1500,-1000, -500, 0, 500, 1000)
  ) +
  labs(
    x = "Realized plant species richness",
y = expression(
  atop(
    "Social value of",
    "SOC (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
  )
)
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(aspect.ratio = 0.6)


p_real_soc_all

ggsave(
  filename = "SOC_real_all.png",
  plot     = p_real_soc_all,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)
```


```{r Plot Realized SR - Leaching}
# Leaching costs without monocultures and realised diversity
c_real_leach_all <- conditional_effects(fitLeachReal)
plot_real_leach_all_data <- c_real_leach_all[[1]]

p_real_leach_all <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = leach_costs_mid),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  geom_line(
    data     = plot_real_leach_all_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  geom_ribbon(
    data  = plot_real_leach_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0.02, 0.02))
  ) +
   scale_y_continuous(
    breaks = c(-125,-100,-75,-50,-25, 0)
  ) +  
  labs(
    x = "Realized plant species richness",
y = expression(
  atop(
    "Social costs of",
    NO[3] * "\u2013N leaching (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
  )
)
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(aspect.ratio = 0.6)


p_real_leach_all

ggsave(
  filename = "Leach_real_all.png",
  plot     = p_real_leach_all,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)
```


```{r Plot realized diversity - RP}
#Insurance value and sown diversity
c_real_rp <- conditional_effects(fitRPreal)
plot_real_rp_data <- c_real_rp[[1]]


p_real_rp <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = sowndiv, y = RP),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4,
    position = position_jitter(width = 0.12, height = 0)
  ) +
  geom_line(
    data     = plot_sown_rp_data,
    aes(x    = 2^log2(sowndiv), y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  geom_ribbon(
    data  = plot_sown_rp_data,
    aes(x = 2^log2(sowndiv), ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant species richness",
y = expression(atop(
  "Risk premium (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)
)
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(aspect.ratio = 0.6)

p_real_rp

ggsave(
  filename = "Fig_real_RP.png",
  plot     = p_real_rp,
  width    = 8,     # in inches; adjust to Nature’s figure width
  height   = 5,
  units    = "in",
  device   = "png",
  dpi      = 300,
  path     = out_dir 
)

```


```{r Plot Realized SR - Multi-panel-PLot}
p_services_real <- (
  p_real_profit_all + p_real_soc_all
) /
(
  p_real_leach_all + p_real_rp
) +
  plot_annotation(tag_levels = "a") &
  theme(
    plot.tag = element_text(
      face = "bold",
      size = 12,
      hjust = -0.2,
      vjust = 1.5
    )
  )

p_services_real

ggsave(
  filename = "Single_Services_real.png",
  plot     = p_services_real,
  bg       = "white",
  path     = out_dir,
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```

```{r Plot Realized SR - Economic MF}
# Economic MF and realised diversity
c_real_total <- conditional_effects(fitTotalReal)
plot_real_total_data <- c_real_total[[1]]

# Economic MF (total costs) vs mean realised diversity

p_real_total <- ggplot() +
  # raw data
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = costs_total_mid),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data     = plot_real_total_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data  = plot_real_total_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  # log2 x-axis, no manual breaks/limits
  scale_x_continuous(
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Realized plant species richness",
y = expression(
  "Economic multifunctionality (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
)

  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_real_total

ggsave(
  filename = "total_real_all.png",
  plot     = p_real_total,
  device   = "png",
  path     = out_dir,
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)


```
