---
title: "Ecological Multifunctinality Analysis"
author: "Marius Munschek"
date: "2025-05-30"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the ecological multifunctionality for (mostly) soil functions in the Jena experiment. I will focus on the following functions:
- Biomass production above and below
- Nitrogen leaching
- Carbon sequestration 
- 
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", out.width="100%", out.height="100%")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(marginaleffects)
```

## Set up a theme
```{r Theme}
theme_set(
  theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(size = 13, hjust = 0.5),
      axis.text = element_text(size = 7),
      axis.text.y = element_text(size = 13),
      axis.text.x = element_text(size = 13),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
# load synthesis dataset
ecologicalMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality.xlsx")
economicMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF.xlsx")

```


### Multidiversity Index
Here I want to integrate the Diversity metrics
```{r Multidiversity}

```

## Plot the densities of the data
```{r Densities}
# check distribution of multifunctionality
ggplot2::ggplot(ecologicalMF,
                aes(x = multifunctionality)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Ecological Multifunctionality",
    y = "Density"
  )

# check the distribution of realized SR
ggplot2::ggplot(ecologicalMF,
                aes(x = meanSR)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Mean Realized Plant Diversity",
    y = "Density"
  )

```


## Statistical Models 
Also single functions or just multifunctionality?
```{r MF ~ sowndiv}
fitMFsown <- brm(
  multifunctionality ~ log2(sowndiv),
 ecologicalMF)

#check model assumptions
    pp_check(fitMFsown, type="scatter_avg") 
    pp_check(fitMFsown, ndraws=50)
    check_model(fitMFsown, check="linearity") 
    check_model(fitMFsown, check="homogeneity")
    plot(fitMFsown)

# check model output
summary(fitMFsown)

```


```{r MF ~ realized SR}

fitMFreal <- brm(
  multifunctionality ~ log2(meanSR),
  ecologicalMF)

# check model assumptions
    pp_check(fitMFreal, type="scatter_avg") 
    pp_check(fitMFreal, ndraws=50)
    check_model(fitMFreal, check="linearity") 
    check_model(fitMFreal, check="homogeneity")
    plot(fitMFreal)
    
#check model output
summary(fitMFreal)


```

```{r correlation economic and ecological MF}
# merge ecological and economic multifunctionality data
joinedMF <- ecologicalMF %>%
  left_join(economicMF, by = "plotcode") 

# Correlation between economic and ecological multifunctionality
cor.test(joinedMF$multifunctionality, joinedMF$costs_total_mid)

```

## Plotting
### Plot biomass against realized species richness
```{r Plotting}
#plot multifunctionality as a response to sowndiv
ggplot(ecologicalMF, aes(x=log2(sowndiv), y=multifunctionality)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  labs(title = "Ecological Multifunctionality vs. Sown Plant Diversity",
       x = "Log2(Sown Plant Diversity)",
       y = "Ecological Multifunctionality")

# plot MF as a function of meanSR
ggplot(ecologicalMF, aes(x=log2(meanSR), y=multifunctionality)) +
    geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  labs(title = "Multifunctionality vs. Realized Species Richness",
       x = "Log2(Realized Species Richness)",
       y = "Multifunctionality")

```

