---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output: html_document
---

# Preface: This R Script is intended to be used to load, explore, and synthesize the datasets for the following metrics measured in the Jena main experiment: Aboveground biomass production, C&N content in the soil, Realized Species Richness, climate variability, water N, pollination, flower diversity, total flowering period per plot, above- and belowground herbivory, soil orgnic carbon (SOC), leaf dry matter content (LDMC), soil aggregate stability, soil diversity, microbial biomass + respiration, 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading packages

```{r Packages}
library(dplyr)
library(readr)
library(readxl)
```


## Load and inspect data

```{r Data}
#load data on aboveground biomass production (ID = 624) 
biomass <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Aboveground/624_9_data.xlsx")
head(biomass)


#load data on realized species richness (ID = 487)
realSR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")
head(realSR)

#load data on soil CN (ID = 655)
soilCN <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/655_12_data.xlsx")

#load data on biomass CN (ID = 557)
biomassCN <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Biomass_Above/557_3_data.xlsx")

# load data on microbial biomass and respiration from 2022 (ID = 398)
microbial <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration/398_3_data.xlsx")


#load data from ZARAH???
unknown <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/2024_May_Main_O2_Results.xlsx")

#load data on water quality
water <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/errechnete Wasserflüsse+Auswaschung.xlsx")

```

## Data wrangling

```{r Wrangling}
#create new dataset that contains only measurements for the year 2022
biomass <- biomass %>% 
  filter(year == 2022) %>% 
  group_by(plotcode) %>% 
  mutate(sum_biomass = sum(biomass)/2) %>% #sum up biomass for each plot and calculate mean between subsamples 
  ungroup() %>%
  distinct(plotcode, sum_biomass)

realSR <- filter(realSR, year == 2022 & season == "Total" ) #keep only total SR in 2022

synthesis <- biomass %>% # merge the datasets 
  left_join(realSR, by = join_by(plotcode))


```



Note to myself: Should I use the corresponding realized species richness from the respective years fopr each soil function?

## Plot biomass against realized species richness

```{r Plotting}
plot(synthesis$realizedSR, synthesis$sum_biomass) #how does the relationship look like?
```

