---
title: "Ecological Multifunctinality Analysis"
author: "Marius Munschek"
date: "2025-05-30"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the ecological multifunctionality for (mostly) soil functions in the Jena Experiment. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", out.width="100%", out.height="100%")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(marginaleffects)
library(ggpubr)
```

## Set up a theme
```{r Theme}
theme_set(
  theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(size = 13, hjust = 0.5),
      axis.text = element_text(size = 7),
      axis.text.y = element_text(size = 12),
      axis.text.x = element_text(size = 12),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
# load synthesis dataset
ecologicalMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality.xlsx") %>% 
  filter(!is.na(sowndiv) & sowndiv < 60 & sowndiv > 1) # remove rows with NA in sowndiv and sowndiv greater than 16
ecologicalMF_red <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality_reduced.xlsx") %>% 
  filter(!is.na(sowndiv) & sowndiv < 60 & sowndiv > 1) # remove rows with NA in sowndiv and sowndiv greater than 16
economicMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF.xlsx") %>% 
  filter(sowndiv > 1 & sowndiv < 60) # remove rows with sowndiv greater than 16

ecologicalMF$plotcode <- as.factor(ecologicalMF$plotcode)
ecologicalMF_red$plotcode <- as.factor(ecologicalMF_red$plotcode)
economicMF$plotcode <- as.factor(economicMF$plotcode)

```

## Plot the densities of the data
```{r Densities}
# check distribution of multifunctionality
ggplot2::ggplot(ecologicalMF,
                aes(x = multifunctionality)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Ecological Multifunctionality",
    y = "Density"
  )

# check the distribution of realized SR
ggplot2::ggplot(ecologicalMF,
                aes(x = meanSR)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Mean Realized Plant Diversity",
    y = "Density"
  )

```

## Statistical Models 
Also single functions or just multifunctionality?
```{r MF ~ sowndiv}
fitMFsown <- brm(
  multifunctionality ~ log2(sowndiv),
 ecologicalMF)

#check model assumptions
    pp_check(fitMFsown, type="scatter_avg") 
    pp_check(fitMFsown, ndraws=50)
    check_model(fitMFsown, check="linearity") 
    check_model(fitMFsown, check="homogeneity")
    plot(fitMFsown)

# check model output
summary(fitMFsown)

```

```{r MF ~ realized SR}

fitMFreal <- brm(
  multifunctionality ~ log2(meanSR),
  ecologicalMF)

# check model assumptions
    pp_check(fitMFreal, type="scatter_avg") 
    pp_check(fitMFreal, ndraws=50)
    check_model(fitMFreal, check="linearity") 
    check_model(fitMFreal, check="homogeneity")
    plot(fitMFreal)
    
#check model output
summary(fitMFreal)


```

## Plotting
### Plot biomass against realized species richness
```{r Plotting}
#plot multifunctionality as a response to sowndiv
p1 <- ggplot(ecologicalMF, aes(x=log2(sowndiv), y=multifunctionality)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
    position = position_jitter(w = 0.03, h = 0)
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
   stat_poly_eq(
    aes(label = paste(..eq.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 5
  ) +
  labs(title = "Ecological Multifunctionality vs. Sown Plant Diversity",
       x = "Log2(Sown Plant Diversity)",
       y = "Ecological Multifunctionality")

p1

# save figure
ggsave("EcologicalMF_sown_noMono.png",
       plot = p1,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

# plot MF as a function of meanSR
p2 <- ggplot(ecologicalMF, aes(x=log2(meanSR), y=multifunctionality)) +
    geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
   stat_poly_eq(
    aes(label = paste(..eq.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 5
  ) +
  labs(title = "Ecological Multifunctionality vs. Realized Species Richness",
       x = "Log2(Realized Species Richness)",
       y = "Ecological Multifunctionality")

p2

# save figure
ggsave("EcologicalMF_realizedSR_noMono.png",
       plot = p2,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

```{r correlation economic and ecological MF}
# merge reduced ecological and economic multifunctionality data
joinedMF <- economicMF %>%
  select(c(plotcode, costs_total_mid)) %>% 
  left_join(ecologicalMF_red, by = "plotcode") 

# Correlation between economic and ecological multifunctionality
cor.test(joinedMF$multifunctionality, joinedMF$costs_total_mid)

# plot the correlation
corrPlot <- ggplot(joinedMF, aes(x = multifunctionality, y = costs_total_mid)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4
  ) +
  stat_cor(method = "pearson", label.x = 0.2, label.y = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Economic vs. Ecological Multifunctionality",
       x = "Ecological Multifunctionality",
       y = "Economic Multifunctionality (Costs/Benefits)")

corrPlot

# save figure
ggsave("Corr_EconomicMF_EcologicalMF.png",
       plot = corrPlot,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

### Marginal Effects and Fitted Model Plots
# plot the marginal effects of the models
delta_y <- function(x) {
  0.306 * log2((x + 1) / x)
}

# Create a data frame for a range of x values
x_vals <- seq(1, 20, by = 1)
effects <- delta_y(x_vals)
df <- data.frame(x = x_vals, delta_y = effects)

# Plot
marginal_plot <- ggplot(df, aes(x = x, y = delta_y)) +
  geom_line(color = "blue", size = 1) +
  labs(
    title = "Marginal Effect of increasing Plant Species number by 1",
    x = "Number of plant species (x)",
    y = "Change in ecological multifunctionality"
  ) +
  theme_minimal()
marginal_plot

# save plot of marginal effects
ggsave("Marginal_Effects_Ecological.png",
       plot = marginal_plot,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


# plot data with logarithmic relationship
# Create a new data frame for prediction
x_seq <- seq(min(ecologicalMF$meanSR), max(ecologicalMF$meanSR), length.out = 100)
pred_df <- data.frame(meanSR = x_seq)
pred_df$predicted_y <- predict(fitMFreal, newdata = pred_df)

# Plot: data + fitted curve
fittedPlot <- ggplot(ecologicalMF, aes(x = meanSR, y = multifunctionality)) +
  geom_point(color = "steelblue", alpha = 0.6) +  # raw data
  geom_line(data = pred_df, aes(x = meanSR, y = pred_df$predicted_y[,"Estimate"]),
            color = "darkgreen", size = 1.2) +    # model prediction
  labs(
    title = "",
    x = "Mean Realized Species Richness",
    y = "Ecological Multifunctionality",
  ) +
  theme_minimal()

fittedPlot

# save plot of fitted model over raw data
ggsave("Fitted_Relationship_Over_Raw_Data_Ecological.png",
       plot = fittedPlot,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```
# Supplements
## Plot the single functions
```{r Single functions}
# plot aggregate stability as function of meanSR
ggplot(ecologicalMF, aes(x=log2(sowndiv), y=aggregates_standardized)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
    position = position_jitter(w = 0.03, h = 0)
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  labs(title = "Aggregate Stability vs. Realized Species Richness",
       x = "Log2(Realized Species Richness)",
       y = "Aggregate Stability")


```

