---
title: "Sub-Analysis of the dominance experiment"
author: "Marius Munschek"
date: "2025-11-06"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the economic multifunctionality of the Jena dominance Experiment. this serves as a sensitivity analysis for the main experiment to prove that my findings hold true, even when only considering the 9 most productive and dominant plant species of the experiment.

```{r setup, include=FALSE}
library(ragg)

knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", out.width="100%", out.height="100%", dev = "ragg_png")

source("C:/Users/mm68tulo/Documents/R/functions/myfunctions.R")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(ggpubr)
```

## Set up a theme
```{r Theme}
theme_set(theme_custom())
```

## Load datasets and filter for relevant observations/measurements
```{r Biomass Data}
#  load biomass data from 2005 to 2017
path <- "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/dom_plantBiomass_2002to2015"
dom_bio_2002 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2002.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2003 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2003.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2004 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2004.tab"), skip = 40, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2005 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2005.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2006 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2006.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2007 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2007.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2008 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2008.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2009 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2009.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2010 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2010.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2011 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2011.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2012 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2012.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2013 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2013.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2014 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2014.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2015 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2015.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2017 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/dom_plantBiomass_2017/376_12_data.xlsx") %>% 
  select(c(Plot, bm)) %>% 
  rename(`Experimental plot` = Plot, `Sown plant biom [g/m**2]` = bm)


# merge Biomass files from 2005 to 2015
dom_bio <- bind_rows(
  mutate(dom_bio_2002, year = 2002),
  mutate(dom_bio_2003, year = 2003),
  mutate(dom_bio_2004, year = 2004),
  mutate(dom_bio_2005, year = 2005),
  mutate(dom_bio_2006, year = 2006),
  mutate(dom_bio_2007, year = 2007),
  mutate(dom_bio_2008, year = 2008),
  mutate(dom_bio_2009, year = 2009),
  mutate(dom_bio_2010, year = 2010),
  mutate(dom_bio_2011, year = 2011),
  mutate(dom_bio_2012, year = 2012),
  mutate(dom_bio_2013, year = 2013),
  mutate(dom_bio_2014, year = 2014),
  mutate(dom_bio_2015, year = 2015),
  mutate(dom_bio_2017, year = 2017)
)
  
# load data on hay prices and costs
SGMS <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/SGMs.xlsx") %>% 
  filter(year < 2016 | year == 2017)
  SGMS$year <- as.factor(SGMS$year) # change year to a factor

# get the number of sown species
design <- dom_bio_2005 %>%
  mutate(
    sowndiv = apply(select(., 9:17), 1, function(x) sum(!is.na(x)))
  ) %>%
  select(`Experimental plot`, sowndiv) %>%
  distinct() %>% 
  rename_at('Experimental plot', ~'plotcode')

# calculate biomass for each year per plot and average across sub samples
dom_bio_a <- dom_bio %>% 
  filter(`Sown plant biom [g/m**2]` >= 0) %>%
  group_by(`Experimental plot`, year) %>%
  summarise(sum_biomass = sum(`Sown plant biom [g/m**2]`, na.rm = TRUE) * 0.01, .groups = "drop")

# merge with profit data
dom_bio_a <- dom_bio_a %>% 
  left_join(SGMS, by = join_by(year)) %>% 
  rename_at('Experimental plot', ~'plotcode')

```

```{r Carbon Data}
# load data on soil organic carbon from 2017 (corg is in milligrams per gram)
dom_soc <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/dom_SOC_2017/carbon_dom.xlsx")  %>% 
  left_join(design, by = join_by(plotcode))
 
```


## Calculate Profit and Insurance
```{r}
# calculate yearly operating profit (€/ha/a)
dom_bio_a$profit_annual <- dom_bio_a$sum_biomass * (dom_bio_a$price - dom_bio_a$costs_ton)

# calculate the biomass stability as the coefficient of variance of the biomass across years (CV = SD/mean)
dom_bio_a <- dom_bio_a %>% 
  group_by(plotcode) %>% 
  mutate(CV = (sd(sum_biomass)/mean(sum_biomass))) 

# sum up biomass for each plot and calculate the mean biomass and mean profit across years. (calculating "mean of CV" doesn't change it's value on plot level, because the CV was already the same for every year per plot)
dom_bio_final <- dom_bio_a %>% 
  group_by(plotcode) %>% 
  summarise(
    mean_biomass = mean(sum_biomass, na.rm = TRUE),
    profit_mean = mean(profit_annual, na.rm = TRUE),
    CV = mean(CV, na.rm = TRUE),
    .groups = "drop"
  )

# calculate certainty equivalent (CE) and risk premium (RP) with profit, CV and risk aversion coefficient (taken from Rommel & Sagebiel et al., 2023)
dom_bio_final$CE <- dom_bio_final$profit_mean * (1+(dom_bio_final$CV)^2)^(-0.6)
dom_bio_final$RP <- dom_bio_final$profit_mean - dom_bio_final$CE

# merge biomass data with design data
dom_bio_final <- dom_bio_final %>% # merge the biomass and the realized SR data
  left_join(design, by = join_by(plotcode))

# define quadratic relationship between biodiversity and insurance value
  quad <- lm(CV ~ sowndiv + I(sowndiv^2), dom_bio_final)

  # calculate dCV/dSR
  dCVdSR <- function(SR) {
  c <- coef(quad)
  c[2] + 2 * c[3] * SR
}
  
  # insurance value like in Martin's code
  dom_bio_final <- dom_bio_final %>% 
    mutate(dCVdSR = dCVdSR(sowndiv),
         insurance = sowndiv * dCVdSR * (-1))

```


## Descriptive plots
```{r Descriptive plots}
# check the relationship between diversity and productivity
ggplot2::ggplot(dom_bio_final,
                aes(x = sowndiv, y = profit_mean)
                ) +
  geom_point(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Sown Plant Diversity",
    y = "Expected Profit (€/ha/year)") +
  scale_x_continuous(breaks = c(1,2,3, 4,6,9))


# check the relationship between CV and productivity
ggplot2::ggplot(dom_bio_final,
                aes(x = sowndiv, y = insurance)
                ) +
  geom_point(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Species Diversity",
    y = "Stability") +
  scale_x_continuous(breaks = c(1,2,3, 4,6,9))

# check the relationship between plot-aggregated-SOC and sowndiv
ggplot2::ggplot(dom_soc,
                aes(x = sowndiv, y = Corg_diff)
                ) +
  geom_point(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Species Diversity",
    y = "Difference in SOC plotwise")  +
  scale_x_continuous(breaks = c(1,2,3, 4,6,9))

```

## Statistical models
```{r Linear models}
fitProfit <- brm(
  profit_mean ~ log2(sowndiv),
  dom_bio_final) 

    # check model assumptions 
    pp_check(fitProfit, type="scatter_avg")
    pp_check(fitProfit, ndraws=50)
    check_model(fitProfit, check="linearity")
    check_model(fitProfit, check="homogeneity")
    plot(fitProfit)
    
    # check output of model
    summary(fitProfit)
    
fitCorg <- brm(
  Corg_diff ~ log2(sowndiv),
  dom_soc)

# check model assumptions 
    pp_check(fitCorg, type="scatter_avg")
    pp_check(fitCorg, ndraws=50)
    check_model(fitCorg, check="linearity")
    check_model(fitCorg, check="homogeneity")
    plot(fitCorg)
    
    # check output of model
    summary(fitCorg)

```
## Final plots
```{r Plots with effect sizes}
#Profit data with monocultures and sowndiv
c_Profit <- conditional_effects(fitProfit)
plot_Profit_data <- c_Profit[[1]]
  
p_Profit <- ggplot() +
  geom_point(data = dom_bio_final, aes(x = sowndiv, y = profit_mean), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_Profit_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_Profit_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "Operating profit (€/ha/year)") +
  scale_x_continuous(breaks = c(1,2,3, 4,6,9))

p_Profit


#Profit data with monocultures and sowndiv
c_Corg <- conditional_effects(fitCorg)
plot_Corg_data <- c_Corg[[1]]
  
p_Corg <- ggplot() +
  geom_point(data = dom_soc, aes(x = sowndiv, y = Corg_diff), 
             alpha = 0.4, color = "#28a87d") +
   # Model-estimated effect
  geom_line(data = plot_Corg_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_Corg_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "SOC net loss/gain") +
  scale_x_continuous(breaks = c(1,2,3, 4,6,9))

p_Corg

ggsave("Corg_dominance.png",
       plot = p_Corg,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

ggsave("Profit_dominance.png",
       plot = p_Profit,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


```

