---
title: "Ecological Multifunctinality Analysis"
author: "Marius Munschek"
date: "2025-05-30"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the ecological multifunctionality for (mostly) soil functions in the Jena Experiment. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", out.width="100%", out.height="100%")

source("C:/Users/mm68tulo/Documents/R/functions/myfunctions.R")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(ggpubr)
```
## Set up a theme
```{r Theme}
theme_set(theme_custom())

```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
# load synthesis dataset
ecologicalMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality.xlsx") %>% 
  filter(!is.na(sowndiv) & sowndiv < 60) # remove rows with NA in sowndiv and sowndiv greater than 16
ecologicalMF_red <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality_reduced.xlsx") %>% 
  filter(!is.na(sowndiv) & sowndiv < 60 & sowndiv > 1) # remove rows with NA in sowndiv and sowndiv greater than 16
economicMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF.xlsx") %>% 
  filter(sowndiv > 1 & sowndiv < 60) # remove rows with sowndiv greater than 16

ecologicalMF$plotcode <- as.factor(ecologicalMF$plotcode)
ecologicalMF_red$plotcode <- as.factor(ecologicalMF_red$plotcode)
economicMF$plotcode <- as.factor(economicMF$plotcode)

```

## Plot the densities of the data
```{r Densities}
# check distribution of multifunctionality
ggplot2::ggplot(ecologicalMF,
                aes(x = multifunctionality)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Ecological Multifunctionality",
    y = "Density"
  ) +
  expand_limits(x = 0, y = 0)

# check the distribution of realized SR
ggplot2::ggplot(ecologicalMF,
                aes(x = meanSR)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Mean Realized Plant Diversity",
    y = "Density"
  ) 

```

## Statistical Models 
Also single functions or just multifunctionality?
```{r MF ~ sowndiv}
fitMFsown <- brm(
  multifunctionality ~ log2(sowndiv),
 ecologicalMF)

#check model assumptions
    pp_check(fitMFsown, type="scatter_avg") 
    pp_check(fitMFsown, ndraws=50)
    check_model(fitMFsown, check="linearity") 
    check_model(fitMFsown, check="homogeneity")
    plot(fitMFsown)

# check model output
summary(fitMFsown)

```

```{r MF ~ realized SR}
fitMFreal <- brm(
  multifunctionality ~ log2(meanSR),
  ecologicalMF)

# check model assumptions
    pp_check(fitMFreal, type="scatter_avg") 
    pp_check(fitMFreal, ndraws=50)
    check_model(fitMFreal, check="linearity") 
    check_model(fitMFreal, check="homogeneity")
    plot(fitMFreal)
    
#check model output
summary(fitMFreal)

```

## Plotting
### Plot biomass against realized species richness
```{r Plotting}
#plot multifunctionality as a response to sowndiv
p1 <- ggplot(ecologicalMF, aes(x=log2(sowndiv), y=multifunctionality)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
    position = position_jitter(w = 0.03, h = 0)
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
   stat_poly_eq(
    aes(label = paste(..eq.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 5
  ) +
  labs(title = "Ecological Multifunctionality vs. Sown Plant Diversity",
       x = "Log2(Sown Plant Diversity)",
       y = "Ecological Multifunctionality")

p1

# save figure
ggsave("EcologicalMF_sown_noMono.png",
       plot = p1,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

# plot MF as a function of meanSR
p2 <- ggplot(ecologicalMF, aes(x=log2(meanSR), y=multifunctionality)) +
    geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE,
    color = "#287da8"
  ) +
   stat_poly_eq(
    aes(label = paste(..eq.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 5
  ) +
  labs(title = "Ecological Multifunctionality vs. Realized Species Richness",
       x = "Log2(Realized Species Richness)",
       y = "Ecological Multifunctionality")

p2

# save figure
ggsave("EcologicalMF_realizedSR_noMono.png",
       plot = p2,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

```{r correlation economic and ecological MF}
# merge reduced ecological and economic multifunctionality data
joinedMF <- economicMF %>%
  select(c(plotcode, costs_total_mid)) %>% 
  left_join(ecologicalMF_red, by = "plotcode") 

# Correlation between economic and ecological multifunctionality
cor.test(joinedMF$multifunctionality, joinedMF$costs_total_mid)

# plot the correlation
corrPlot <- ggplot(joinedMF, aes(x = multifunctionality, y = costs_total_mid)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4
  ) +
  stat_cor(method = "pearson", label.x = 0.2, label.y = 700, size = 5) +
  geom_smooth(method = "lm", se = F, color = "#287da8") +
  labs(title = "Economic vs. Ecological Multifunctionality",
       x = "Ecological Multifunctionality",
       y = "Economic Multifunctionality (Costs/Benefits)")

corrPlot

# save figure
ggsave("Corr_EconomicMF_EcologicalMF.png",
       plot = corrPlot,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```
# Supplements
## Single Function Models
```{r Model Single Functions}
fitLeachReal

fitBiomassReal

fitMicbioReal

fitMicresReal

fitRootReal

fitSOCReal

fitPolReal

fitAboveReaL

fitBelowReal

fitStabiReal

fitAggReal



```


## Plot the single functions
```{r Plot Single functions}
# plot aggregate stability as function of meanSR
ggplot(ecologicalMF, aes(x=log2(sowndiv), y=aggregates_standardized)) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
    position = position_jitter(w = 0.03, h = 0)
  ) +
   geom_smooth(
    method = "lm",
    se = FALSE
  ) +
  labs(title = "Aggregate Stability vs. Realized Species Richness",
       x = "Log2(Realized Species Richness)",
       y = "Aggregate Stability")


```

