---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output:
  html_document: default
  pdf_document: default
---

# Preface
This R Script is intended to be used to load, explore, and synthesize the datasets for the following metrics measured in the Jena main experiment: Aboveground biomass production, C&N content in the soil, Realized Species Richness, climate variability, leaching N (nitrogen leaching), pollination, flower diversity, total flowering period per plot, above- and belowground herbivory (for belowground measured as energy fluxes in the food web), soil organic carbon (SOC), leaf dry matter content (LDMC), soil diversity, microbial biomass + respiration, 
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE, fig.width = 8, fig.height = 6, dpi = 300, cache = TRUE, cache.lazy = FALSE, cache.path = "cache/", tidy = TRUE, comment = NA, collapse = TRUE)
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
# load data on plot design (needed for 'block code')
design <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Plot_Design/90_20_data.xlsx")

#load data on aboveground biomass production (ID = 624) 
biomassDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/624_9_data.xlsx")
  biomassDat$plotcode <- as.factor(biomassDat$plotcode) # change plotcode to a factor

# load data on SGMs 
SGMS <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/SGMs.xlsx") %>% 
  filter(year > 2010)
  SGMS$year <- as.factor(SGMS$year) # change year to a factor

#load data on realized species richness (ID = 487)
SR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")
  #change data type for realized SR & sowndiv
  SR$realizedSR <- as.numeric(SR$realizedSR)
  SR$sowndiv <- as.numeric(SR$sowndiv)

#load data on soil organic carbon (SOC) (2020: ID = 424 2011: ID = 254): 2011 is to compare the carbon content at the beginning of the experiment with the latest (2020) carbon content. 2020 data is the most recent year that includes bulk density measurements.
carbon20 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/424_7_data.xlsx")
carbon02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/254_3_data.xlsx") 
  # change data types
  carbon02$N_Soil <- as.numeric(carbon02$N_Soil)
  carbon02$Corg_Soil <- as.numeric(carbon02$Corg_Soil)
  carbon02$Cinorg_Soil <- as.numeric(carbon02$Cinorg_Soil)
  carbon02$BulkDensity_Soil <- as.numeric(carbon02$BulkDensity_Soil)

#load data on nitrogen leaching (this should be in kg/ha per year)
leaching <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/Yearly_Leaching_Plot.xlsx") %>% 
  filter(year > 2010) %>%  # filter for years after 2010
  rename_at('plot', ~'plotcode') %>% # rename plot to plotcode
  mutate(plotcode = as.factor(plotcode)) # change plotcode to a factor
```

## Data wrangling
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated. 
```{r Biomass data}
# calculate biomass for each year per plot and average across subsamples
biomassAnnual <- biomassDat %>% 
  filter(year > 2010) %>% 
  group_by(plotcode, year, subsample) %>%
  summarise(sum_biomass = sum(biomass, na.rm = TRUE), .groups = "drop") %>% 
  group_by(plotcode, year) %>%
  summarise(mean_biomass = mean(sum_biomass, na.rm = TRUE), .groups = "drop")

biomassAnnual$year <- as.factor(biomassAnnual$year) # change year to a factor

# merge annual biomass data with the SGMs dataset
biomassAnnual <- biomassAnnual %>% 
  left_join(SGMS, by = join_by(year)) %>% 
  select(plotcode, year, mean_biomass, price, costs)  # select only relevant columns
  
# calculate individual SGM for each year
biomassAnnual$SGM_annual <- biomassAnnual$mean_biomass * 0.1 * biomassAnnual$price - biomassAnnual$costs 

# sum up biomass for each plot and calculate the mean biomass and SGM across years
biomassMeans <- biomassAnnual %>% 
  group_by(plotcode) %>% 
  summarise(
    mean_biomass = mean(mean_biomass, na.rm = TRUE), 
    SGM_mean = mean(SGM_annual, na.rm = TRUE), 
    .groups = "drop"
  )
```

```{r soil carbon}
# rename plot to plotcode and change data types 
carbon02 <- carbon02 %>%
  rename_at('plot', ~'plotcode') %>% 
  mutate(
    plotcode = as.factor(plotcode),
    c_org = as.numeric(Corg_Soil),
    c_inorg = as.numeric(Cinorg_Soil)
  )
  
carbon20 <- carbon20 %>%
  rename_at('plot', ~'plotcode') %>%
  mutate(
    plotcode = as.factor(plotcode),
    c_org = as.numeric(Corg_Soil),
    c_inorg = as.numeric(Cinorg_Soil)
  )


# create a new column with mean C across layers per plot for 2002
carbon02_clean <- carbon02 %>% group_by(plotcode) %>% 
                    mutate(mean_org_02 = mean(c_org), mean_inorg_02 = mean(c_inorg)) %>%
                      distinct(plotcode, mean_org_02, mean_inorg_02) %>% 
                      mutate(mean_C_02 = mean_org_02 + mean_inorg_02)

# create a new column with mean C across layers per plot for 2020
carbon20_clean <- carbon20 %>% group_by(plotcode) %>% 
                      mutate(mean_org_20 = mean(c_org), mean_inorg_20 = mean(c_inorg), mean_bulk = mean(BulkDensity_Soil)) %>%
                      distinct(plotcode, mean_org_20, mean_inorg_20, mean_bulk) %>% 
                      mutate(mean_C_20 = mean_org_20 + mean_inorg_20)

  
```

```{r Species Richness data}

realdiv <- SR %>%                    #calculate the mean realized SR for the years 2003-2019
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)

SR <- distinct(SR, plotcode, sowndiv)

```


```{r leaching}
leachingMean <- leaching %>% 
            summarise(av_leaching = mean(leach_y, na.rm = TRUE), .by = plotcode)
   # calculate the mean leaching per plot

leachingMean$av_leaching <- leachingMean$av_leaching / 100 # convert from g/m2 to kg/ha  
  
# calculate the costs in €/ha/a for each plot

leachingMean$leach_costs_lo <- leachingMean$av_leaching * 6.3 *(-1)
leachingMean$leach_costs_mid <- leachingMean$av_leaching * 11.23 *(-1)
leachingMean$leach_costs_hi <- leachingMean$av_leaching * 20 *(-1)
```

## Synthesize in one Dataset
```{r Data Synthesis}
synthDat <- biomassMeans %>% # merge the biomass and the realized SR data
  left_join(SR, by = join_by(plotcode)) %>%
  left_join(carbon02_clean, by = join_by(plotcode)) %>% 
  left_join(carbon20_clean, by = join_by(plotcode)) %>% 
  left_join(select(design, plotcode, block)) %>% 
  left_join(leachingMean, by = join_by(plotcode))  %>%
  left_join(realdiv, by = join_by(plotcode))

synthDat$block <- as.factor(synthDat$block) # change block to a factor
synthDat$plotcode <- as.factor(synthDat$plotcode)  # change plotcode to a factor
  
# create a column with carbon difference between 2002 and 2020
synthDat$mean_org_diff <- synthDat$mean_org_20 - synthDat$mean_org_02

# create a column with social costs of carbon
synthDat$SOC_costs_mid <- synthDat$mean_org_diff * synthDat$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 195 / 12 # mid price scenario (SOC value multiplied with bulk de
synthDat$SOC_costs_lo <- synthDat$mean_org_diff * synthDat$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 70 / 12 # low price scenario - ETS price
synthDat$SOC_costs_hi <- synthDat$mean_org_diff * synthDat$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 268.64 / 12 # high price scenario
# add up all the costs for total value
synthDat$costs_total_mid <- synthDat$SOC_costs_mid + synthDat$leach_costs_mid + synthDat$SGM_mean

# save the synthesized dataset
 writexl::write_xlsx(synthDat, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF.xlsx")

```
