---
title: "Sub-Analysis of the dominance experiment"
author: "Marius Munschek"
date: "2025-11-06"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the economic multifunctionality of the Jena dominance Experiment. this serves as a sensitivity analysis for the main experiment to prove that my findings hold true, even when only considering the 9 most productive and dominant plant species of the experiment.

```{r setup, include=FALSE}
library(ragg)

knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", out.width="100%", out.height="100%", dev = "ragg_png")

source("C:/Users/mm68tulo/Documents/R/functions/myfunctions.R")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(ggpubr)
```

## Set up a theme
```{r Theme}
theme_set(theme_custom())
```

## Load datasets and filter for relevant observations/measurements
```{r Biomass Data}
#  load biomass data from 2003 to 2017
path <- "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/dom_plantBiomass_2002to2015"
dom_bio_2003 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2003.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2004 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2004.tab"), skip = 40, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2005 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2005.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2006 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2006.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2007 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2007.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2008 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2008.tab"), skip = 38, show_col_types = FALSE) %>% filter(Repl == "mean")
dom_bio_2009 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2009.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2010 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2010.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2011 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2011.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2012 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2012.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2013 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2013.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2014 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2014.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2015 <- read_tsv(file.path(path, "JenExp_biomass_DOM_2015.tab"), skip = 33, show_col_types = FALSE)
dom_bio_2017 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/dom_plantBiomass_2017/376_12_data.xlsx") %>% 
  dplyr::select(c(Plot, bm)) %>% 
  rename(`Experimental plot` = Plot, `Sown plant biom [g/m**2]` = bm)


# merge Biomass files from 2003 to 2017
dom_bio <- bind_rows(
  mutate(dom_bio_2003, year = 2003),
  mutate(dom_bio_2004, year = 2004),
  mutate(dom_bio_2005, year = 2005),
  mutate(dom_bio_2006, year = 2006),
  mutate(dom_bio_2007, year = 2007),
  mutate(dom_bio_2008, year = 2008),
  mutate(dom_bio_2009, year = 2009),
  mutate(dom_bio_2010, year = 2010),
  mutate(dom_bio_2011, year = 2011),
  mutate(dom_bio_2012, year = 2012),
  mutate(dom_bio_2013, year = 2013),
  mutate(dom_bio_2014, year = 2014),
  mutate(dom_bio_2015, year = 2015),
  mutate(dom_bio_2017, year = 2017)
)
  
# load data on hay prices and costs
SGMS <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/SGMs.xlsx") %>% 
  filter(year < 2018)

# get the number of sown species
design <- dom_bio_2005 %>%
  mutate(
    sowndiv = apply(dplyr::select(., 9:17), 1, function(x) sum(!is.na(x)))
  ) %>%
  dplyr::select(`Experimental plot`, sowndiv) %>%
  distinct() %>% 
  rename_at('Experimental plot', ~'plotcode')

# calculate biomass for each year per plot and average across sub samples
dom_bio_a <- dom_bio %>% 
  filter(`Sown plant biom [g/m**2]` >= 0) %>%
  group_by(`Experimental plot`, year) %>%
  summarise(sum_biomass = sum(`Sown plant biom [g/m**2]`, na.rm = TRUE) * 0.01, .groups = "drop")

# merge with profit data
dom_bio_a <- dom_bio_a %>% 
  left_join(SGMS, by = join_by(year)) %>% 
  rename_at('Experimental plot', ~'plotcode')

```

```{r Carbon Data}
# load prepared dataframe for comparing 2002 and 2017 SOC stock (Soc content for 2017 is already in this table)
dom_soc_diff <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/dom_SOC_2017/SOC17_dom.xlsx")  %>% 
  left_join(design, by = join_by(plotcode))

# load SOC and bulk density measures for 2002
dom_soc_02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Dominance_Experiment/carbon02_dom.xlsx")

# join with dom_soc_diff table
dom_soc_diff <- dom_soc_diff %>% 
  left_join(dom_soc_02, by = "main_plot") 

# create a column with annual SOC gain/loss
dom_soc_diff$Corg_diff <- (as.numeric(dom_soc_diff$Corg17) - dom_soc_diff$Corg02) / 15 

# create a column with mid SOC value for beginning and end
dom_soc_diff$Corg02_val <- dom_soc_diff$Corg02 * dom_soc_diff$Bulk02 * 0.3 * 10000 / 1000 * 3.66 * 300
dom_soc_diff$Corg17_val <- as.numeric(dom_soc_diff$Corg17) * dom_soc_diff$Bulk02 * 0.3 * 10000 / 1000 * 3.66 * 300

# create a column with annual SOC-value gain/loss
dom_soc_diff$Corg_val_diff <- (as.numeric(dom_soc_diff$Corg17_val) - dom_soc_diff$Corg02_val) / 15


writexl::write_xlsx(dom_soc_diff, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/soc_dominance_2002+2017.xlsx") 
```


## Calculate Profit and Insurance
```{r}
# calculate yearly operating profit (€/ha/a)
dom_bio_a$profit_annual <- dom_bio_a$sum_biomass * (dom_bio_a$price - dom_bio_a$costs_ton)

# calculate the biomass stability as the coefficient of variance of the biomass across years (CV = SD/mean)
dom_bio_a <- dom_bio_a %>% 
  group_by(plotcode) %>% 
  mutate(CV = (sd(sum_biomass)/mean(sum_biomass))) 

# sum up biomass for each plot and calculate the mean biomass and mean profit across years. (calculating "mean of CV" doesn't change it's value on plot level, because the CV was already the same for every year per plot)
dom_bio_final <- dom_bio_a %>% 
  group_by(plotcode) %>% 
  summarise(
    mean_biomass = mean(sum_biomass, na.rm = TRUE),
    profit_mean = mean(profit_annual, na.rm = TRUE),
    CV = mean(CV, na.rm = TRUE),
    .groups = "drop"
  )

# calculate certainty equivalent (CE) and risk premium (RP) with profit, CV and risk aversion coefficient (taken from Rommel & Sagebiel et al., 2023)
r_RS <- 0.214
gamma <- 1 - r_RS

dom_bio_final$CE <- dom_bio_final$profit_mean * (1+(dom_bio_final$CV)^2)^(-gamma/2)
dom_bio_final$RP <- dom_bio_final$profit_mean - dom_bio_final$CE

# merge biomass data with design data
dom_bio_final <- dom_bio_final %>% # merge the biomass and the realized SR data
  left_join(design, by = join_by(plotcode))

writexl::write_xlsx(dom_bio_final, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/dominance_mean_biomass_profit_2002_to_2017.xlsx") 

```

## Statistical models
```{r Linear models}
fitProfit <- brm(
  profit_mean ~ log2(sowndiv),
  dom_bio_final) 

    # check model assumptions 
    pp_check(fitProfit, type="scatter_avg")
    pp_check(fitProfit, ndraws=50)
    check_model(fitProfit, check="linearity")
    check_model(fitProfit, check="homogeneity")
    plot(fitProfit)
    
    # check output of model
    summary(fitProfit)
    
fitCorg <- brm(
  Corg_val_diff ~ log2(sowndiv),
  dom_soc_diff)

# check model assumptions 
    pp_check(fitCorg, type="scatter_avg")
    pp_check(fitCorg, ndraws=50)
    check_model(fitCorg, check="linearity")
    check_model(fitCorg, check="homogeneity")
    plot(fitCorg)
    
    # check output of model
    summary(fitCorg)

```
## Final plots
```{r Plot profit - sown diversity}
# plot conditional effects together with dominance results
c_Profit <- conditional_effects(fitProfit)
plot_Profit_data <- c_Profit[[1]]

p_Profit <- ggplot() +
  # raw data
  geom_point(
    data = dom_bio_final,
    aes(x = sowndiv, y = profit_mean),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#D55E00",
    colour = "#8C3A00",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data = plot_Profit_data,
    aes(x = sowndiv, y = estimate__),
    colour    = "#D55E00",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_Profit_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#D55E00",
    alpha = 0.15
  ) +
  # x-axis: log2 scale to match your modelling strategy
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 6, 9),
    labels = c(1, 2, 3, 4, 6, 9),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(100, 300, 500, 700)
  ) +
  labs(
    x = "Sown plant species richness",
    y = expression(
  "Profit (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")")
  ) +
  theme_custom()

p_Profit

# Save
ggsave(
  filename = "fig_Profit_dominance.png",
  plot     = p_Profit,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 7,
  height   = 5,
  units    = "in"
)

```


```{r Plot SOC - sown diversity}
# plot conditional effects together with dominance results
c_Corg <- conditional_effects(fitCorg)
plot_Corg_data <- c_Corg[[1]]

p_Corg <- ggplot() +
  # raw data
  geom_point(
    data   = dom_soc_diff,
    aes(x  = sowndiv, y = Corg_val_diff),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#D55E00",
    colour = "#8C3A00",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data     = plot_Corg_data,
    aes(x    = sowndiv, y = estimate__),
    colour   = "#D55E00",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data  = plot_Corg_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#D55E00",
    alpha = 0.15
  ) +
  scale_x_continuous(
    breaks = c(1, 2, 6, 9),
    labels = c(1, 2, 6, 9),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant species richness",
y = expression(
  atop(
    "Social value of",
    "SOC (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")")
    )
  ) +
  theme_custom()

p_Corg

# save plot
ggsave(
  filename = "fig_Corg_dominance.png",
  plot     = p_Corg,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 7,
  height   = 5,
  units    = "in"
)

```

### Create a 2-panel figure
```{r}

fig_dominance <- (p_Profit + p_Corg) +
  plot_annotation(tag_levels = "a") &
  theme(
    # tags inside top-left, slightly nudged right
    plot.tag.position = c(0.06, 0.98),
    plot.tag = element_text(
      face = "bold",
      size = 12,   # match your other figure
      hjust = 0,
      vjust = 1
    ),

    # keep spacing compact and consistent
    axis.title.y = element_text(margin = margin(r = 2)),
    plot.margin = margin(2, 2, 2, 2)
  )

fig_dominance

# Save as PNG
ggsave(
  filename = "Fig_dominance_profit_SOC.png",
  plot     = fig_dominance,
  width    = 6.5, 
  height   = 3.8,
  units    = "in",
  dpi      = 300,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/"
)
```


## Create subset dataframe for supplement figure
```{r}
MainExp17 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MainExp17.xlsx")

MainExp17$Corg_val_diff <- MainExp17$SOC_diff * MainExp17$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 300

dom_bio_final <- dom_bio_final %>% 
  dplyr::select(plotcode, sowndiv, profit_mean)

dom_soc_diff <- dom_soc_diff %>% dplyr::select(plotcode, Corg_val_diff)

dom_bio_final <- dom_bio_final %>% 
  left_join(dom_soc_diff, by = join_by(plotcode)) %>% 
  mutate(experiment = 'dom')

MainExp17 <-  MainExp17 %>% 
  dplyr::select(plotcode, sowndiv, profit_mean, Corg_val_diff) %>% 
  mutate(experiment = 'main')

MainDomDiff <- rbind(MainExp17, dom_bio_final)

writexl::write_xlsx(MainDomDiff, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MainDomDiff.xlsx")
```

```{r Plot difference in productivity: main - dominance}
# create linear model for main experiment profit
fitProfitMain <- brm(
  profit_mean ~ log2(sowndiv),
  MainExp17) 

    # check model assumptions 
    pp_check(fitProfitMain, type="scatter_avg")
    pp_check(fitProfitMain, ndraws=50)
    check_model(fitProfitMain, check="linearity")
    check_model(fitProfitMain, check="homogeneity")
    plot(fitProfit)
    
    # check output of model
    summary(fitProfitMain)
    
# plot conditional effects together with dominance results
c_ProfitMain <- conditional_effects(fitProfitMain)
plot_Profit_Main_data <- c_ProfitMain[[1]]

p_profit_dif <- ggplot() +
  # Raw data points (two experiments)
  geom_point(
    data = MainDomDiff,
    aes(x = sowndiv, y = profit_mean, fill = experiment),
    position = position_jitter(width = 0.15, height = 0),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    colour = "black",
    stroke = 0.3
  ) +
  scale_fill_manual(
  values = c(
    "main" = "#0072B2",
    "dom"  = "#D55E00"
  ),
  breaks = c("main", "dom"),  # <- controls legend order
  name = "Experiment",
  labels = c("Main experiment", "Dominance experiment")
) +

  # Dominance model (orange)
  geom_ribbon(
    data = plot_Profit_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#D55E00",
    alpha = 0.15
  ) +
  geom_line(
    data = plot_Profit_data,
    aes(x = sowndiv, y = estimate__),
    colour    = "#D55E00",
    linewidth = 0.8
  ) +

  # Main experiment model (blue)
  geom_ribbon(
    data = plot_Profit_Main_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  geom_line(
    data = plot_Profit_Main_data,
    aes(x = sowndiv, y = estimate__),
    colour    = "#0072B2",
    linewidth = 0.8
  ) +

  # Axes
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 6, 8, 9, 16),
    labels = c(1, 2, 3, 4, 6, 8, 9, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant species richness",
    y = expression(Profit ~ "(" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")")
  ) +
  theme_custom() +
  theme(
    legend.position  = "bottom",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 11),
    legend.direction = "horizontal",
    legend.box       = "horizontal",
    legend.spacing.x = unit(4, "pt")
  )

p_profit_dif

ggsave(
  filename = "Profit_difference.png",
  plot     = p_profit_dif,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 5.9,
  height   = 4.5,
  units    = "in"
)

```


```{r Plot difference in SOC: main - dominance}
#Soil organic carbon gain/loss and sowndiv - for main experiment until 2017
fitCorgMain <- brm(
  Corg_val_diff ~ log2(sowndiv),
  MainExp17)

# check model assumptions 
    pp_check(fitCorgMain, type="scatter_avg")
    pp_check(fitCorgMain, ndraws=50)
    check_model(fitCorgMain, check="linearity")
    check_model(fitCorgMain, check="homogeneity")
    plot(fitCorgMain)
    
    # check output of model
    summary(fitCorgMain)

c_Corg_Main <- conditional_effects(fitCorgMain)
plot_Corg_data_Main <- c_Corg_Main[[1]]

# make a nice plot  
# SOC benefit ~ sown diversity (main vs dominance comparison)

p_Corg_diff <- ggplot() +
  # raw data points (two experiments)
  geom_point(
    data = MainDomDiff,
    aes(x = sowndiv, y = Corg_val_diff, fill = experiment),
    position = position_jitter(width = 0.15, height = 0),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    colour = "black",
    stroke = 0.3
  ) +
 scale_fill_manual(
  values = c(
    "main" = "#0072B2",
    "dom"  = "#D55E00"
  ),
  breaks = c("main", "dom"),  # <- controls legend order
  name = "Experiment",
  labels = c("Main experiment", "Dominance experiment")
) +

  # Main experiment model (blue)
  geom_ribbon(
    data = plot_Corg_data_Main,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  geom_line(
    data     = plot_Corg_data_Main,
    aes(x    = sowndiv, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +

  # Dominance experiment model (orange)
  geom_ribbon(
    data = plot_Corg_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#D55E00",
    alpha = 0.15
  ) +
  geom_line(
    data     = plot_Corg_data,
    aes(x    = sowndiv, y = estimate__),
    colour   = "#D55E00",
    linewidth = 0.8
  ) +

  # axes
  scale_x_continuous(
    breaks = c(1, 2, 3, 4, 6, 8, 9, 16),
    labels = c(1, 2, 3, 4, 6, 8, 9, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant species richness",
    y = expression(
    "Social value of SOC (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")"
  )
  ) +
  theme_custom() +
  theme(
    legend.position  = "bottom",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 11),
    legend.direction = "horizontal",
    legend.box       = "horizontal",
    legend.spacing.x = unit(4, "pt")
  )

p_Corg_diff

ggsave(
  filename = "SOC_difference.png",
  plot     = p_Corg_diff,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 5.9,
  height   = 4.5,
  units    = "in"
)

```


