---
title: "Economic Multifunctionality Analysis"
author: "Marius Munschek"
date: "2025-05-28"
output: html_document
---

#Preface
This R Script is written by Marius Munschek to calculate the economic multifunctionality for ecosystem services in the Jena Experiment. The performance of the following ecosystem sevices per plot is averaged over the years 2011 - 2020, then translated into monetary values which are aggregated to obtain economic multifunctionality values per plot:
- Biomass production
- Nitrogen leaching
- Carbon sequestration

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, fig.width=10, fig.height=6, dpi=300, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", fig.path="figures/", out.width="100%", out.height="100%")

source("C:/Users/mm68tulo/Documents/R/functions/myfunctions.R")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
```

## Theme Custom
```{r Theme}
theme_set(theme_custom())
```

## Load datasets - Filter for relevant observations/measurements
```{r Read Data}
#load synthesis dataset
synthDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF_final.xlsx") %>% 
  filter(sowndiv < 60) %>%  #remove 60 species plots
  mutate(block = as.factor(block))



```

## Data Wrangling
Change data types and filter for relevant observations.
```{r Synthesis data}
synthDat$plotcode <- as.factor(synthDat$plotcode) #change the plotcode to a factor

# add insurance to mid costs
synthDat$costs_total_mid <- synthDat$costs_total_mid + synthDat$insurance_RP

# add up all the low costs for lower total value
synthDat$costs_total_lo <- synthDat$SOC_costs_lo + synthDat$leach_costs_lo + synthDat$profit_mean + synthDat$insurance_RP

# add up all the high costs for higher total value
synthDat$costs_total_hi <- synthDat$SOC_costs_hi + synthDat$leach_costs_hi + synthDat$profit_mean + synthDat$insurance_RP

#noMono <- synthDat %>% 
 # filter(sowndiv > 1) #filter out monocultures

```

## Statistical Models
### Linear models with monocultures
#### &Sown diversity
```{r Sown Diversity WITH Monocultures, include=FALSE}
# Profit against sown plant diversity
fitProfit <- brm(
  profit_mean ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat) 

    # check model assumptions 
    pp_check(fitProfit, type="scatter_avg")
    pp_check(fitProfit, ndraws=50)
    check_model(fitProfit, check="linearity")
    check_model(fitProfit, check="homogeneity")
    plot(fitProfit)
    
    # check output of model
    summary(fitProfit)

# SOC against sown plant diversity
fitSOC <- brm(
  SOC_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitSOC, type="scatter_avg") 
    pp_check(fitSOC, ndraws=50)
    check_model(fitSOC, check="linearity") 
    check_model(fitSOC, check="homogeneity")
    plot(fitSOC)
    
    # check output of model
    summary(fitSOC)

# Leaching costs against sown plant diversity
fitLeach <- brm(
  leach_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitLeach, type="scatter_avg") 
    pp_check(fitLeach, ndraws=50)
    check_model(fitLeach, check="linearity") 
    check_model(fitLeach, check="homogeneity")
    plot(fitLeach)
    
    # check output of model
    summary(fitLeach)

# Insurance value against sown plant diversity
fitInsurance <- brm(
  insurance_RP ~log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat
)    
    
    # check model assumptions
    pp_check(fitInsurance, type="scatter_avg") 
    pp_check(fitInsurance, ndraws=50)
    check_model(fitInsurance, check="linearity") 
    check_model(fitInsurance, check="homogeneity")
    plot(fitInsurance)
    
    #check model output
    summary(fitInsurance)
    
    
# Total Costs (medium) against sown plant diversity
fitTotal <- brm(
  costs_total_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitTotal, type="scatter_avg") 
    pp_check(fitTotal, ndraws=50)
    check_model(fitTotal, check="linearity") 
    check_model(fitTotal, check="homogeneity")
    plot(fitTotal)
    
    # check output of model
    summary(fitTotal)
    
```

#### &Realized diversity
```{r Realized Diversity WITH Monocultures, include=FALSE}
# Profit against species richness - including monocultures
fitProfitReal <- brm(
  profit_mean ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat) 

    # check model assumptions 
    pp_check(fitProfitReal, type="scatter_avg")
    pp_check(fitProfitReal, ndraws=50)
    check_model(fitProfitReal, check="linearity")
    check_model(fitProfitReal, check="homogeneity")
    plot(fitProfitReal)
    
    # check output of model
    summary(fitProfitReal)

# SOC against species richness - including monocultures
fitSOCreal <- brm(
  SOC_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitSOCreal, type="scatter_avg") 
    pp_check(fitSOCreal, ndraws=50)
    check_model(fitSOCreal, check="linearity") 
    check_model(fitSOCreal, check="homogeneity")
    plot(fitSOCreal)
    
    # check output of model
    summary(fitSOCreal)

# Leaching costs against species richness - including monocultures
fitLeachReal <- brm(
  leach_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitLeachReal, type="scatter_avg") 
    pp_check(fitLeachReal, ndraws=50)
    check_model(fitLeachReal, check="linearity") 
    check_model(fitLeachReal, check="homogeneity")
    plot(fitLeachReal)
    
    # check output of model
    summary(fitLeachReal)

# Insurance against species richness - including monocultures
fitInsuranceReal <- brm(
  insurance_RP ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitInsuranceReal, type="scatter_avg") 
    pp_check(fitInsuranceReal, ndraws=50)
    check_model(fitInsuranceReal, check="linearity") 
    check_model(fitInsuranceReal, check="homogeneity")
    plot(fitInsuranceReal)
    
    # check output of model
    summary(fitInsuranceReal)        

# Total Costs against mean realized species richness - including monocultures
fitTotalReal <- brm(
  costs_total_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitTotalReal, type="scatter_avg") 
    pp_check(fitTotalReal, ndraws=50)
    check_model(fitTotalReal, check="linearity") 
    check_model(fitTotalReal, check="homogeneity")
    plot(fitTotalReal)
    
    # check output of model
    summary(fitTotalReal)


```

## Plotting
```{r Plot sown diversity  - profit}
#Profit data with monocultures and sowndiv
c_sown_Profit <- conditional_effects(fitProfit)
plot_sown_Profit_data <- c_sown_Profit[[1]]
  
p_sown_Profit <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = profit_mean),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data = plot_sown_Profit_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval
  geom_ribbon(
    data = plot_sown_Profit_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant diversity (species)",
    y = "Profit (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6   # width:height ~ 1 : 0.6 (tweak to taste)
  )

p_sown_Profit

ggsave(
  filename = "Fig_sown_profit.png",
  plot     = p_sown_Profit,
  width    = 8,     # in inches; adjust to Nature’s figure width
  height   = 5,
  units    = "in",
  device   = "png",
  dpi      = 300,
  path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/" 
)
```

```{r Plot sown diversity - SOC}
# SOC costs with monocultures and mean sown diversity
c_sown_soc <- conditional_effects(fitSOC)
plot_sown_soc_data <- c_sown_soc[[1]]

p_sown_soc <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = SOC_costs_mid),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data   = plot_sown_soc_data,
    aes(x  = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_sown_soc_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1500, -1000, -500, 0, 500, 1000)
  ) +
  labs(
    x = "Sown plant diversity (species)",
    y = "SOC cost (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6   # width:height ~ 1 : 0.6, matching Fig_sown_profit
  )

p_sown_soc

ggsave(
  filename = "SOC_sown.png",
  plot     = p_sown_soc,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```

```{r Plot sown diversity - Leaching}
#Leaching costs and mean sown diversity
c_sown_leach <- conditional_effects(fitLeach)
plot_sown_leach_data <- c_sown_leach[[1]]
  
# Leaching costs: sown diversity
p_sown_leach <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = leach_costs_mid),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data = plot_sown_leach_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval
  geom_ribbon(
    data = plot_sown_leach_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-125, -100, -75, -50, -25)
  ) +
  labs(
    x = "Sown plant diversity (species)",
    y = "Leaching cost (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_sown_leach

ggsave(
  filename = "Fig_Leach_sown.png",
  plot     = p_sown_leach,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```

```{r Plot sown diversity - Insurance}
#Insurance value and sown diversity
c_sown_insurance <- conditional_effects(fitInsurance)
plot_sown_insurance_data <- c_sown_insurance[[1]]
  
# Insurance value ~ sown diversity
p_sown_insurance <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = insurance_RP),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data = plot_sown_insurance_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_sown_insurance_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Sown plant diversity (species)",
    y = "Insurance"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_sown_insurance

ggsave(
  filename = "fig_Insurance_sown.png",
  plot     = p_sown_insurance,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)


#Risk premium and sown diversity
c_sown_rp <- conditional_effects(fitRP)
plot_sown_rp_data <- c_sown_rp[[1]]
  
p_sown_rp <- ggplot() +
  geom_point(data = synthDat, aes(x = sowndiv, y = RP), 
             alpha = 0.4, color = "#0072B2") +
   # Model-estimated effect
  geom_line(data = plot_sown_rp_data, aes(x = sowndiv, y = estimate__), color = "#0072B2", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_sown_rp_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#0072B2", alpha = 0.2) +
  labs(x = "Sown plant diversity", y = "Risk premium (€/ha/a)") +
  scale_x_continuous(breaks = c(1,2,4, 8, 16))

p_sown_rp

ggsave("RP_sown.png",
       plot = p_sown_rp,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)
```

```{r Plot sown diversity - Economic MF}
#economic MF and sown diversity
c_sown_total <- conditional_effects(fitTotal)
plot_sown_total_data <- c_sown_total[[1]]
  
# Economic multifunctionality ~ sown diversity
p_sown_total <- ggplot() +
  # raw data
  geom_point(
    data = synthDat,
    aes(x = sowndiv, y = costs_total_mid),
    alpha = 0.5,
    size  = 2,
    shape = 21,
    fill  = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model line
  geom_line(
    data = plot_sown_total_data,
    aes(x = sowndiv, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_sown_total_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1500, -1000, -500, 0, 500, 1000)
  ) +
  labs(
    x = "Sown plant diversity (species)",
    y = "Economic multifunctionality (€/ha/year)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_sown_total

ggsave(
  filename = "fig_total_sown.png",
  plot     = p_sown_total,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)


```

```{r Create multi-panel plot}

fig_sown_all <- (
  p_sown_Profit +
  p_sown_soc
) / (
  p_sown_leach +
  p_sown_insurance
) +
  plot_annotation(
    tag_levels = "a"
  ) &
  theme(
    plot.tag = element_text(
      face = "bold",
      size = 14,
      hjust = -0.2,
      vjust = 1.5
    )
  )

fig_sown_all

ggsave(
  filename = "Fig_sown_multifunctionality.png",
  plot     = fig_sown_all,
  width    = 8,
  height   = 8,
  units    = "in",
  dpi      = 300,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/"
)


```


### Sensitivity Analysis with different price scenarios
```{r Sensitivity Analysis}
# Low Total Costs against species richness
fitTotalLo <- brm(
  costs_total_lo ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitTotalLo, type="scatter_avg") 
    pp_check(fitTotalLo, ndraws=50)
    check_model(fitTotalLo, check="linearity") 
    check_model(fitTotalLo, check="homogeneity")
    plot(fitTotalLo)
    
    # check output of model
    summary(fitTotalLo)

# High Total Costs against species richness
fitTotalHi <- brm(
  costs_total_hi ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  synthDat)

    # check model assumptions
    pp_check(fitTotalHi, type="scatter_avg") 
    pp_check(fitTotalHi, ndraws=50)
    check_model(fitTotalHi, check="linearity") 
    check_model(fitTotalHi, check="homogeneity")
    plot(fitTotal4)
    
    # check output of model
    summary(fitTotalHi)
    
```

### Plot sensitivity analysis
```{r Plot Sown Diversity - Economic MF - Alternative Pricing}
c_sown_total <- conditional_effects(fitTotal)
plot_sown_total_data <- c_sown_total[[1]]

c_sown_total_lo <- conditional_effects(fitTotalLo)
plot_sown_total_data_lo <- c_sown_total_lo[[1]]

c_sown_total_hi <- conditional_effects(fitTotalHi)
plot_sown_total_data_hi <- c_sown_total_hi[[1]]


# Combine the three conditional_effects data frames
plot_sown_total_all <- bind_rows(
  mutate(plot_sown_total_data,    scenario = "Mid price (default)"),
  mutate(plot_sown_total_data_lo, scenario = "Low price"),
  mutate(plot_sown_total_data_hi, scenario = "High price")
)

# prepare data for combined plot
library(tidyr)

raw_sown_total_all <- synthDat %>%
  select(sowndiv, costs_total_mid, costs_total_lo, costs_total_hi) %>%
  pivot_longer(
    cols      = c(costs_total_mid, costs_total_lo, costs_total_hi),
    names_to  = "scenario_raw",
    values_to = "economic_mf"
  ) %>%
  mutate(
    scenario = dplyr::recode(
      scenario_raw,
      "costs_total_mid" = "Mid price (default)",
      "costs_total_lo"  = "Low price",
      "costs_total_hi"  = "High price"
    )
  )

# Plot without raw data points
p_sown_total <- ggplot() +
  
  # horizontal reference line at y = 0
  geom_hline(
    yintercept = 0,
    colour     = "grey80",
    linewidth  = 0.6
  ) +
  
  # CI ribbon for mid/default scenario only
  geom_ribbon(
    data = filter(plot_sown_total_all, scenario == "Mid price (default)"),
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  
  # model lines for all three price scenarios
  geom_line(
    data = plot_sown_total_all,
    aes(x = sowndiv, y = estimate__, colour = scenario, linetype = scenario),
    linewidth = 0.9
  ) +
  
  # colours and line types for scenarios
  scale_colour_manual(
    name   = "Price scenario",
    values = c(
      "Mid price (default)" = "#0072B2",  # blue
      "Low price"           = "#009E73",  # green
      "High price"          = "#E69F00"   # orange
    )
  ) +
  scale_linetype_manual(
    name   = "Price scenario",
    values = c(
      "Mid price (default)" = "solid",
      "Low price"           = "dashed",
      "High price"          = "dotted"
    )
  ) +
  
  # log2 x-axis
  scale_x_continuous(
    trans  = "log2",
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  
  # y-axis
  scale_y_continuous(
    breaks = c(-1500, -1000, -500, 0, 500, 1000, 1500)
  ) +
  
  # labels
  labs(
    x = "Sown plant diversity (species)",
    y = "Economic multifunctionality (€/ha/yr)"
  ) +
  
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio     = 0.6,
    legend.position  = "bottom",
    legend.direction = "horizontal",
    legend.box       = "horizontal",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 11),
    legend.spacing.x = unit(4, "pt")
  )

p_sown_total

ggsave(
  filename = "total_sown_different_pricing.png",
  plot     = p_sown_total,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```

## Supplementary Analysis
#### Plot realized species diversity effects including monocultures
```{r Plot Realized SR - Profit}
# Profit data with monocultures and realized diversity
c_real_profit_all <- conditional_effects(fitProfitReal)
plot_real_profit_all_data <- c_real_profit_all[[1]]

p_real_profit_all <- ggplot() +
  # raw data
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = profit_mean),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data     = plot_real_profit_all_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data  = plot_real_profit_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  
  # --- LOG2 AXIS ---
  scale_x_continuous(
  trans  = "log2",
  expand = expansion(mult = c(0.02, 0.02))
  ) + 

  # labels
  labs(
    x = "Mean realized plant species (n)",
    y = "Profit (€/ha/yr)"
  ) +
  
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_real_profit_all

ggsave(
  filename = "Profit_real_all.png",
  plot     = p_real_profit_all,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```


```{r Plot Realized SR - SOC}
# SOC benefits without monocultures and realised diversity
c_real_soc_all <- conditional_effects(fitSOCreal)
plot_real_soc_all_data <- c_real_soc_all[[1]]

p_real_soc_all <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = SOC_costs_mid),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  geom_line(
    data     = plot_real_soc_all_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  geom_ribbon(
    data  = plot_real_soc_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1500,-1000, -500, 0, 500, 1000)
  ) +
  labs(
    x = "Mean realized plant species (n)",
    y = "SOC benefit (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(aspect.ratio = 0.6)


p_real_soc_all

ggsave(
  filename = "SOC_real_all.png",
  plot     = p_real_soc_all,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)
```


```{r Plot Realized SR - Leaching}
# Leaching costs without monocultures and realised diversity
c_real_leach_all <- conditional_effects(fitLeachReal)
plot_real_leach_all_data <- c_real_leach_all[[1]]

p_real_leach_all <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = leach_costs_mid),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  geom_line(
    data     = plot_real_leach_all_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  geom_ribbon(
    data  = plot_real_leach_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
   scale_y_continuous(
    breaks = c(-125,-100,-75,-50,-25, 0)
  ) +  
  labs(
    x = "Mean realized plant species (n)",
    y = "Leaching cost (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(aspect.ratio = 0.6)


p_real_leach_all

ggsave(
  filename = "Leach_real_all.png",
  plot     = p_real_leach_all,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)
```


```{r Plot Realized SR - Insurance}
# Insurance value without monocultures and realised diversity
c_real_insurance_all <- conditional_effects(fitInsuranceReal)
plot_real_insurance_all_data <- c_real_insurance_all[[1]]

p_real_insurance_all <- ggplot() +
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = insurance_RP),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data = plot_real_insurance_all_data,
    aes(x = meanSR, y = estimate__),
    colour = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data = plot_real_insurance_all_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill = "#0072B2",
    alpha = 0.15
  ) +
  scale_x_continuous(
    trans  = "log2",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Mean realized plant species (n)",
    y = "Insurance (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(aspect.ratio = 0.6)


p_real_insurance_all

ggsave(
  filename = "Insurance_real_all.png",
  plot     = p_real_insurance_all,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)

```

```{r Plot Realized SR - Multi-panel-PLot}
p_services_real <- (
  p_real_profit_all + p_real_soc_all
) /
(
  p_real_leach_all + p_real_insurance_all
) +
  plot_annotation(tag_levels = "a")

ggsave(
  filename = "Single_Services_real.png",
  plot     = p_services_real,
  bg       = "white",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 8,
  units    = "in"
)

```

```{r Plot Realized SR - Economic MF}
# Economic MF and realised diversity
c_real_total <- conditional_effects(fitTotalReal)
plot_real_total_data <- c_real_total[[1]]

# Economic MF (total costs) vs mean realised diversity

p_real_total <- ggplot() +
  # raw data
  geom_point(
    data   = synthDat,
    aes(x  = meanSR, y = costs_total_mid),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  # model-estimated effect
  geom_line(
    data     = plot_real_total_data,
    aes(x    = meanSR, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +
  # credible interval ribbon
  geom_ribbon(
    data  = plot_real_total_data,
    aes(x = meanSR, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +
  # log2 x-axis, no manual breaks/limits
  scale_x_continuous(
    trans  = "log2",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    x = "Mean realised plant species (n)",
    y = "Economic multifunctionality (€/ha/yr)"
  ) +
  coord_cartesian(clip = "off") +
  theme_custom() +
  theme(
    aspect.ratio = 0.6
  )

p_real_total

ggsave(
  filename = "total_real_all.png",
  plot     = p_real_total,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 8,
  height   = 5,
  units    = "in"
)


```


### sowndiv without Monocultures
```{r Sown Diversity WITHOUT Monocultures}
# Profit against species richness
fitProfitSown <- brm(
  profit_mean ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono) 

    # check model assumptions 
    pp_check(fitProfitSown, type="scatter_avg")
    pp_check(fitProfitSown, ndraws=50)
    check_model(fitProfitSown, check="linearity")
    check_model(fitProfitSown, check="homogeneity")
    plot(fitProfitSown)
    
    # check output of model
    summary(fitProfitSown)

# SOC against species richness
fitSOCsown <- brm(
  SOC_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono)

    # check model assumptions
    pp_check(fitSOCsown, type="scatter_avg") 
    pp_check(fitSOCsown, ndraws=50)
    check_model(fitSOCsown, check="linearity") 
    check_model(fitSOCsown, check="homogeneity")
    plot(fitSOCsown)
    
    # check output of model
    summary(fitSOCsown)

# Leaching costs against species richness
fitLeachSown <- brm(
  leach_costs_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono)

    # check model assumptions
    pp_check(fitLeachSown, type="scatter_avg") 
    pp_check(fitLeachSown, ndraws=50)
    check_model(fitLeachSown, check="linearity") 
    check_model(fitLeachSown, check="homogeneity")
    plot(fitLeachSown)
    
    # check output of model
    summary(fitLeachSown)

# Insurance value against sowndiv
    fitInsuranceSown <- brm(
      insurance ~ log2(sowndiv) + (1|block),
      control = list(adapt_delta = 0.95),
      iter = 3000,
      noMono
    )
    
    # check model assumptions
    pp_check(fitInsuranceSown, type="scatter_avg") 
    pp_check(fitInsuranceSown, ndraws=50)
    check_model(fitInsuranceSown, check="linearity") 
    check_model(fitInsuranceSown, check="homogeneity")
    plot(fitInsuranceSown)
    
    # check output of model
    summary(fitInsuranceSown)
    
# Total Costs against sowndiv
fitTotalSown <- brm(
  costs_total_mid ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  noMono)

    # check model assumptions
    pp_check(fitTotalSown, type="scatter_avg") 
    pp_check(fitTotalSown, ndraws=50)
    check_model(fitTotalSown, check="linearity") 
    check_model(fitTotalSown, check="homogeneity")
    plot(fitTotalSown)
    
    # check output of model
    summary(fitTotalSown)

```

### Realized Diversity without Monocultures
```{r Realized Diversity without Monocultures}
# Profit against species richness
fitProfitReal <- brm(
  profit_mean ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono) 

    # check model assumptions 
    pp_check(fitProfitReal, type="scatter_avg")
    pp_check(fitProfitReal, ndraws=50)
    check_model(fitProfitReal, check="linearity")
    check_model(fitProfitReal, check="homogeneity")
    plot(fitProfitReal)
    
    # check output of model
    summary(fitProfitReal)

# SOC against species richness
fitSOCreal <- brm(
  SOC_costs_mid ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
  noMono)

    # check model assumptions
    pp_check(fitSOCreal, type="scatter_avg") 
    pp_check(fitSOCreal, ndraws=50)
    check_model(fitSOCreal, check="linearity") 
    check_model(fitSOCreal, check="homogeneity")
    plot(fitSOCreal)
    
    # check output of model
    summary(fitSOCreal)

# Leaching costs against species richness
    fitLeachReal <- brm(
      leach_costs_mid ~ log2(meanSR) + (1|block),
      control = list(adapt_delta = 0.95),
      iter = 3000,
      noMono)

    # check model assumptions
    pp_check(fitLeachReal, type="scatter_avg") 
    pp_check(fitLeachReal, ndraws=50)
    check_model(fitLeachReal, check="linearity") 
    check_model(fitLeachReal, check="homogeneity")
    plot(fitLeachReal)
    
    # check output of model
    summary(fitLeachReal)
    
# Insurance value against realized species richness
    fitInsuranceReal <- brm(
      insurance ~ log2(meanSR) + (1|block),
      control = list(adapt_delta = 0.95),
      iter = 3000,
      noMono
    )
    
    # check model assumptions
    pp_check(fitInsuranceReal, type="scatter_avg") 
    pp_check(fitInsuranceReal, ndraws=50)
    check_model(fitInsuranceReal, check="linearity") 
    check_model(fitInsuranceReal, check="homogeneity")
    plot(fitInsuranceReal)
    
    # check output of model
    summary(fitInsuranceReal)    

# Total Costs against realized species richness
fitTotalReal <- brm(
  costs_total_mid ~ log2(meanSR) + (1|block),
  noMono)

    # check model assumptions
    pp_check(fitTotalReal, type="scatter_avg") 
    pp_check(fitTotalReal, ndraws=50)
    check_model(fitTotalReal, check="linearity") 
    check_model(fitTotalReal, check="homogeneity")
    plot(fitTotalReal)
    
    # check output of model
    summary(fitTotalReal)

    
```





### Plot realized species diversity effects - no Monocultures
```{r Plotting}
#Profit data without monocultures and realdiv
c_real_Profit <- conditional_effects(fitProfitReal)
plot_real_Profit_data <- c_real_Profit[[1]]
  
p_real_Profit <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = profit_mean), 
             alpha = 0.4, color = "#0072B2") +
   # Model-estimated effect
  geom_line(data = plot_real_Profit_data, aes(x = meanSR, y = estimate__), color = "#0072B2", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_Profit_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#0072B2", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Mean operating profit (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(50, 100, 150, 200, 250))
  

p_real_Profit

ggsave("Profit_real_noMono.png",
       plot = p_real_Profit,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)


#SOC costs without monocultures and mean realized diversity
c_real_soc <- conditional_effects(fitSOCreal)
plot_real_soc_data <- c_real_soc[[1]]
  
p_real_soc <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = SOC_costs_mid), 
             alpha = 0.4, color = "#0072B2") +
   # Model-estimated effect
  geom_line(data = plot_real_soc_data, aes(x = meanSR, y = estimate__), color = "#0072B2", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_soc_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#0072B2", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Social benefits of SOC sequestration (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_real_soc

ggsave("SOC_real_noMono.png",
       plot = p_real_soc,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Leaching costs without monocultures and mean realized diversity
c_real_leach <- conditional_effects(fitLeachReal)
plot_real_leach_data <- c_real_leach[[1]]
  
p_real_leach <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = leach_costs_mid), 
             alpha = 0.4, color = "#0072B2") +
   # Model-estimated effect
  geom_line(data = plot_real_leach_data, aes(x = meanSR, y = estimate__), color = "#0072B2", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_leach_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#0072B2", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Social costs of nitrogen leaching (€/ha/year)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12))

p_real_leach

ggsave("Leach_real_noMono.png",
       plot = p_real_leach,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#Insurance value without monocultures and mean realized diversity
c_real_insurance <- conditional_effects(fitInsuranceReal)
plot_real_insurance_data <- c_real_insurance[[1]]
  
p_real_insurance <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = insurance), 
             alpha = 0.4, color = "#0072B2") +
   # Model-estimated effect
  geom_line(data = plot_real_insurance_data, aes(x = meanSR, y = estimate__), color = "#0072B2", size = .8, alpha = .6) +

  # Credible interval ribbon
  geom_ribbon(data = plot_real_insurance_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#0072B2", alpha = 0.2) +
  labs(x = "Mean realized plant diversity", y = "Insurance value (€)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12))

p_real_insurance

ggsave("Insurance_real_noMono.png",
       plot = p_real_insurance,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

# combine the plots for realized SR in one grid
p_services_real <- p_real_Profit + p_real_soc + p_real_leach + p_real_insurance +
  plot_annotation(title = "Profit, SOC benefit, Leaching Costs, and Insurance value plotted against realized plant diversity") &
  theme(plot.title = element_text(size=13,hjust=0.5))

ggsave("Single_Services_real.png",
       plot = last_plot(),
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#economic MF without monocultures and mean realized diversity
c_real_total <- conditional_effects(fitTotalReal)
plot_real_total_data <- c_real_total[[1]]

c_real_total_lo <- conditional_effects(fitTotalLo)
plot_sown_total_data_lo <- c_real_total_lo[[1]]

c_real_total_hi <- conditional_effects(fitTotalHi)
plot_sown_total_data_hi <- c_real_total_hi[[1]]
  
p_real_total <- ggplot() +
  geom_point(data = noMono, aes(x = meanSR, y = costs_total_mid), 
             alpha = 0.4, color = "#0072B2") +
   # Model-estimated effect
  geom_line(data = plot_real_total_data, aes(x = meanSR, y = estimate__), color = "blue", size = 1, alpha = .6) +
  geom_line(data = plot_sown_total_data_lo, aes(x = meanSR, y = estimate__), color = "green", linewidth = .8, alpha = .6) +
  geom_line(data = plot_sown_total_data_hi, aes(x = meanSR, y = estimate__), color = "yellow", linewidth = .8, alpha = .6) +
  
  # add line at y = 0
  geom_hline(yintercept = 0, color = "grey90", linewidth = 0.8, linetype = 1, alpha = .5) +

  # Credible interval ribbon

#    geom_ribbon(data = plot_sown_total_data_lo, aes(x = meanSR, ymin = lower__, ymax = upper__), 
#              fill = "green", alpha = 0.2) +
#      geom_ribbon(data = plot_sown_total_data_hi, aes(x = meanSR, ymin = lower__, ymax = upper__), 
#              fill = "yellow", alpha = 0.2) +
    geom_ribbon(data = plot_real_total_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.4) +
  labs(x = "Mean realized plant diversity", y = "Economic multifunctionality (€/ha/a)") +
  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12)) +
  scale_y_continuous(breaks = c(-500, -250, 0, 250, 500, 750))

p_real_total

ggsave("total_real_noMono.png",
       plot = p_real_total,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```
