---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output: html_document
---

# Preface
This R Script is intended to be used to load, explore, and synthesize the datasets for the following metrics measured in the Jena main experiment: Aboveground biomass production, C&N content in the soil, Realized Species Richness, climate variability, leaching N (nitrogen leaching), pollination, flower diversity, total flowering period per plot, above- and belowground herbivory (for belowground measured as energy fluxes in the food web), soil organic carbon (SOC), leaf dry matter content (LDMC), soil diversity, microbial biomass + respiration, 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation
## Loading packages
```{r Packages, include=FALSE}
library(dplyr)
library(readr)
library(readxl)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data, include=FALSE}
# load data on plot design (needed for 'block code')
design <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Plot_Design/90_20_data.xlsx")

#load data on aboveground biomass production (ID = 624) 
biomassDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/624_9_data.xlsx")

#load data on root biomass production (ID = 714) 

#load data on realized species richness (ID = 487)
SR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")
  #change data type for realized SR & sowndiv
  SR$realizedSR <- as.numeric(SR$realizedSR)
  SR$sowndiv <- as.numeric(SR$sowndiv)

#load data on soil CN (2022: ID = 655; 2020: ID = 424 2002: ID = 278): 2002 is to compare the carbon content at the beginning of the experiment with the current (2022) carbon content. 2020 data includes bulk density measurements.

#soilCN22 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/655_12_data.xlsx")  
  
soilCN20 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/424_7_data.xlsx")
soilCN02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/278_2_data.xlsx") %>% 
  filter(year == 2002) 

#load data on biomass CN (ID = 557)
biomassCN <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Biomass_Above/557_3_data.xlsx")

# load data on microbial biomass and respiration from 2002 to 2020 (ID = 78) 2022 (ID = 398) and 2021 (ID = 240)
microbial <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration_03-20/78_5_data.xlsx")

#load data on enzyme activity?

#load climate data
climate <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Climate_Data/484_7_data.xlsx")

#load LDMC data
ldmc <- read_csv("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/LDMC/Leaf_Root_traits_dbef_2021.csv")

#load energy flux data (for belowground herbivory calculation)

#load aboveground herbivory data (only treatment D3 -> main experiment)
herbivoryAbove <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Above/480_4_data.xlsx") %>%  filter(treatment == "D3")

#load data on soil diversity
  # Nematodes 2017 (ID = 323)
  # Bacteria
  # AMF
  # Oomycota
  # Cercozoa
  # Arthropodes


#load data on nitrogen leaching (this should be in kg/ha per year)
leaching <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/Overall_Leaching_Plot.xlsx")

#soil infiltration rates


#load data on pollinators
pollinators <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

#load data on flower cover
flowers <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

```

## Data wrangling
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated. 
```{r Biomass data}
biomassDat$plotcode <- as.factor(biomassDat$plotcode) #change the plotcode to a factor
# samplingSeasons <- unique(biomassDat[c("season","year")]) # used to check whether every year had the same number of sampling seasons

# calculate total biomass per plot for 2022
biomassDat <- biomassDat %>% 
  filter(year == 2022) %>% 
  group_by(plotcode, year) %>% 
  summarise(sum_biomass = sum(biomass, na.rm = TRUE)/2, .groups = "drop") # divide by 2 to get the mean biomass per plot (2 subsamples per plot)

# introduce new column with SGM
biomassDat$SGM <- (biomassDat$sum_biomass*0.1)*(6.15)-96

# exclude the years before 2018
# sum up biomass for each plot and calculate the mean across subsamples and years
biomassMeans <- biomassDat %>% 
  filter(year > 2004) %>% 
  group_by(plotcode, year, season) %>% 
  summarise(sum_biomass = sum(biomass, na.rm = TRUE), .groups = "drop") %>%
  group_by(plotcode) %>%
  summarise(mean_biomass = mean(sum_biomass, na.rm = TRUE), .groups = "drop")

biomassMeans$SGM <- (biomassMeans$mean_biomass*0.1)*(6.83)-180

```

```{r Yield stability}
#here I need to look at the harvested biomass across several years and how it fluctuates (calculate the mean difference/variance?)


```

```{r Biomass CN data}
biomassCN <- biomassCN %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
biomassCN$`C:N ratio` <- as.numeric(biomassCN$`C:N ratio`) # change CN ratio values to numeric

biomassCN <- biomassCN %>% group_by(plotcode) %>%                      # calculate mean CN value between the two measurements
  mutate(CNratio_av = mean(`C:N ratio`, na.rm = TRUE))
```

```{r Microbial data}
microbial <- microbial %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
```

```{r Soil CN}
# rename plot to plotcode and change data types 
soilCN02 <- soilCN02 %>%
  rename_at('plot', ~'plotcode') %>% 
  mutate(
    plotcode = as.factor(plotcode),
    Corg_Soil = as.numeric(Corg_Soil),
    Cinorg_Soil = as.numeric(Cinorg_Soil)
  )
  
soilCN20 <- soilCN20 %>%
  rename_at('plot', ~'plotcode') %>%
  mutate(
    plotcode = as.factor(plotcode),
    Corg_Soil = as.numeric(Corg_Soil),
    Cinorg_Soil = as.numeric(Cinorg_Soil)
  )
 
# soilCN22 <- soilCN22 %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
#  soilCN22$plotcode <- as.factor(soilCN22$plotcode)

# create a new column with mean C across layers per plot for 2002
soilCN02_clean <- soilCN02 %>% group_by(plotcode) %>% 
                    mutate(mean_org_02 = mean(Corg_Soil), mean_inorg_02 = mean(Cinorg_Soil)) %>%
                      distinct(plotcode, mean_org_02, mean_inorg_02) %>% 
                      mutate(mean_C_02 = mean_org_02 + mean_inorg_02)

# create a new column with mean C across layers per plot for 2020
soilCN20_clean <- soilCN20 %>% group_by(plotcode) %>% 
                      mutate(mean_org_20 = mean(Corg_Soil), mean_inorg_20 = mean(Cinorg_Soil), mean_bulk = mean(BulkDensity_Soil)) %>%
                      distinct(plotcode, mean_org_20, mean_inorg_20, mean_bulk) %>% 
                      mutate(mean_C_20 = mean_org_20 + mean_inorg_20)


# create a new column with mean C across layers per plot for 2022
# soilCN22_clean <- soilCN22 %>% group_by(plotcode) %>% 
#                      mutate(mean_C_22 = mean(TC), mean_N_22 = mean(TN)) %>%
#                      distinct(plotcode, mean_C_22, mean_N_22)  
                      
# convert carbon unit from percentage to g/kg 
# soilCN22_clean$mean_C_mass_22 <- soilCN22_clean$mean_C_22 * 10
  
```

```{r Species Richness data}
SR <- distinct(SR, plotcode, sowndiv)

realdiv <- filter(SR, year > 2018 & season == "Total" ) #keep only total SR from year 2018 onwards

realdiv <- SR %>%                    #calculate the mean realized SR for the years 2003-2019
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)
```


```{r leaching}
leaching <- leaching %>% 
            rename_at('plot', ~'plotcode') # rename plot to plotcode

# calculate the costs in €/ha/a for each plot

leaching$leach_costs <- leaching$TNleaching_av * 11.23 *(-1)
```

```{r Herbivory}
herbivoryAbove <- herbivoryAbove %>% 
            rename_at('plot', ~'plotcode') # rename plot to plotcode

```

## Synthesize in one Dataset
```{r Data Synthesis}
synthDat <- biomassMeans %>% # merge the biomass and the realized SR data
  left_join(SR, by = join_by(plotcode)) %>%
  left_join(soilCN02_clean, by = join_by(plotcode)) %>% 
  left_join(soilCN20_clean, by = join_by(plotcode)) %>% 
  left_join(select(design, plotcode, block)) %>% 
  left_join(leaching, by = join_by(plotcode))  %>%
  left_join(realdiv, by = join_by(plotcode))

  left_join(herbivoryAbove, by )

synthDat$block <- as.factor(synthDat$block) # change block to a factor
synthDat$plotcode <- as.factor(synthDat$plotcode)  # change plotcode to a factor
  
# create a column with carbon difference between 2002 and 2020
synthDat$mean_org_diff <- synthDat$mean_org_20 - synthDat$mean_org_02

# create a column with social costs of carbon
synthDat$SOC_costs <- synthDat$mean_org_diff * synthDat$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 195 / 20 # in €/ha/a and converted to CO2 and a carbon CO2 price of 195€/t

# add up all the costs for total value
synthDat$costs_total <- synthDat$SOC_costs + synthDat$leach_costs + synthDat$SGM

# save the synthesized dataset
write.csv(synthDat, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/synthesized_means_data.csv", row.names = FALSE)

```
