---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output: html_document
---

# Preface
This R Script is intended to be used to load, explore, and synthesize the datasets for the following metrics measured in the Jena main experiment: Aboveground biomass production, C&N content in the soil, Realized Species Richness, climate variability, water N, pollination, flower diversity, total flowering period per plot, above- and belowground herbivory, soil orgnic carbon (SOC), leaf dry matter content (LDMC), soil aggregate stability, soil diversity, microbial biomass + respiration, 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(lme4)
```


## Load datasets and filter for relevant observations/measurements
```{r Read Data}
#load data on aboveground biomass production (ID = 624) 
biomassDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/624_9_data.xlsx")
head(biomass)

#load data on realized species richness (ID = 487)
realSR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")

realSR$realizedSR <- as.numeric(realSR$realizedSR)
head(realSR)

# BE AWARE THAT YOU NEED TO CHECK SR INFLUENCE FOR DIFFERENT YEARS, ACCORDING TO MEASURED FUNCTION

#load data on soil CN (2022: ID = 655; 2002: ID = 278)
soilCN22 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/655_12_data.xlsx")
soilCN02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/278_2_data.xlsx") %>% 
  filter(depthU < 0.1) #filter for first two soil layers (0-5 mm & 5-10mm) 

#load data on biomass CN (ID = 557)
biomassCN <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Biomass_Above/557_3_data.xlsx")

# load data on microbial biomass and respiration from 2022 (ID = 398)
microbial <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration/398_3_data.xlsx")

#load data on soil aggregates

#load data on enzyme activity?

#load climate data
climate <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Climate_Data/484_7_data.xlsx")

#load LDMC data
ldmc <- read_csv("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/LDMC/Leaf_Root_traits_dbef_2021.csv")

#load energy flux data (for belowground herbivory calculation)

#load aboveground herbivory data (only treatment D3 -> main experiment)
herbivoryAbove <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Above/480_4_data.xlsx") %>%  filter(herbivoryAbove$treatment == "D3")

#load data on soil diversity

#load data on water quality
water <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/Overall_Leaching_Plot.xlsx")

#load data on pollinators
pollinators <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

#load data on flower cover
flowers <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

```


## Data wrangling
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated. 
```{r Biomass data}
biomassDat$plotcode <- as.factor(biomassDat$plotcode) #change the plotcode to a factor

#create new dataset that contains only measurements for the year 2022
biomassPlot <- biomassDat %>% 
  filter(year == 2022) %>% 
  group_by(plotcode) %>% 
  mutate(sum_biomass = sum(biomass, na.rm = TRUE)/2) %>% #sum up biomass for each plot and calculate mean between subsamples 
  ungroup() %>%
  distinct(plotcode, sum_biomass)
```

```{r Yield stability}
#here I need to look at the harvested biomass across several years and how it fluctuates (calculate the mean difference/variance?)


```


```{r Biomass CN data}
biomassCN <- biomassCN %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
biomassCN$`C:N ratio` <- as.numeric(biomassCN$`C:N ratio`) # change CN ratio values to numeric

biomassCN <- biomassCN %>% group_by(plotcode) %>%                      # calculate mean CN value between the two measurements
  mutate(CNratio_av = mean(`C:N ratio`, na.rm = TRUE))
```

```{r Microbial data}
microbial <- microbial %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
```

```{r Soil CN}
soilCN22 <- soilCN22 %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
soilCN02 <- soilCN02 %>% rename_at('plot', ~'plotcode') # rename plot to plotcode

soilCN02 %>% group_by(plotcode) %>% 
  mutate(C_total = sum(Corg_Soil) + sum(Cinorg_Soil)) #need to aggregate for 0-10mm and sum up organic and inorganic
  

```

```{r Species Richness data}
realSR22 <- filter(realSR, year == 2022 & season == "Total" ) #keep only total SR in 2022

realSRmean <- filter(realSR, year < 2019 & season == "Total" ) %>% #calculate the mean realized SR for the years 2003-2019
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)
```

--> should I calculate the mean realized species richness for the nitrogen leaching analysis? I mean, the yearly average of leaking is calculated from the years 2003 - 2019, therefore I should take the corresponding realized SRs, right?




```{r Water}
water <- water %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
```

```{r Herbivory}
herbivoryAbove <- herbivoryAbove %>% rename_at('plot', ~'plotcode') # rename plot to plotcode

```



```{r Data Synthesis}
synthesis22 <- biomassPlot %>% # merge the biomass and the realized SR data
  left_join(realSRmean, by = join_by(plotcode)) %>% 
  left_join(water, by = join_by(plotcode)) # %>%
  left_join(herbivoryAbove, by )
```



Note to myself: Should I use the corresponding realized species richness from the respective years for each soil function?

## Run some basic linear models
```{r}
#check distribution of data
hist(synthesis22$sum_biomass)
plot(synthesis22$realizedSR, synthesis22$sum_biomass)

lmBio <- lm(sum_biomass ~ as.numeric(realizedSR), synthesis22)
summary(lmBio)

plot(lmBio)


#check the distribution of the data
hist(synthesis22$TNleaching_av)
plot(synthesis22$meanSR, synthesis22$TNleaching_av)

#create linear model
lmN <- lm(TNleaching_av ~ meanSR, synthesis22)
plot(lmN)
summary(lmN)
```



## Plot biomass against realized species richness

```{r Plotting}
plot(synthesis$realizedSR, synthesis$sum_biomass) #how does the relationship look like?
```

