---
title: "Ecological Multifunctinality Analysis"
author: "Marius Munschek"
date: "2025-05-30"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the ecological multifunctionality for (mostly) soil functions in the Jena Experiment. 

```{r setup, include=FALSE}
library(ragg)

knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE, cache=TRUE, cache.lazy=FALSE, cache.path="cache/", out.width="100%", out.height="100%", dev = "ragg_png")

source("C:/Users/mm68tulo/Documents/R/functions/myfunctions.R")
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
library(ggpubr)
library(BayesFactor)
```
## Set up a theme
```{r Theme}
theme_set(theme_custom())
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data}
# load synthesis dataset
ecologicalMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality_reduced.xlsx") %>% 
  filter(!is.na(sowndiv) & sowndiv < 60) # remove rows with NA in sowndiv and sowndiv greater than 16
economicMF <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/economic_MF_final.xlsx") %>% 
  filter(sowndiv < 60) # remove rows with sowndiv greater than 16

ecologicalMF$plotcode <- as.factor(ecologicalMF$plotcode)
ecologicalMF$plotcode <- as.factor(ecologicalMF$plotcode)
ecologicalMF$block <- as.factor(ecologicalMF$block)
ecologicalMF$block <- as.factor(ecologicalMF$block)
economicMF$plotcode <- as.factor(economicMF$plotcode)

```

## Statistical Models 
Also single functions or just multifunctionality?
```{r MF ~ sowndiv}
fitMFsown <- brm(
  multifunctionality ~ log2(sowndiv) + (1|block),
  control = list(adapt_delta = 0.95),
  iter = 3000,
 ecologicalMF)

#check model assumptions
    pp_check(fitMFsown, type="scatter_avg") 
    pp_check(fitMFsown, ndraws=50)
    check_model(fitMFsown, check="linearity") 
    check_model(fitMFsown, check="homogeneity")
    plot(fitMFsown)

# check model output
summary(fitMFsown)

```

```{r MF ~ realized SR}
fitMFreal <- brm(
  multifunctionality ~ log2(meanSR) + (1|block),
  control = list(adapt_delta = 0.95),
  ecologicalMF)

# check model assumptions
    pp_check(fitMFreal, type="scatter_avg") 
    pp_check(fitMFreal, ndraws=50)
    check_model(fitMFreal, check="linearity") 
    check_model(fitMFreal, check="homogeneity")
    plot(fitMFreal)
    
#check model output
summary(fitMFreal)

```

## Plotting
### Plot multifunctionality against species richness
```{r Plotting}
#plot multifunctionality as a response to sowndiv
c_sown_mf <- conditional_effects(fitMFsown)
plot_sown_mf_data <- c_sown_mf[[1]]

p_sown_mf <- ggplot() +
  # raw data
  geom_point(
    data   = ecologicalMF,
    aes(x  = sowndiv, y = multifunctionality),
    alpha  = 0.5,
    size   = 2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4,
    position = position_jitter(width = 0.12, height = 0)
  ) +
  
  # model-estimated effect
  geom_line(
    data     = plot_sown_mf_data,
    aes(x    = sowndiv, y = estimate__),
    colour   = "#0072B2",
    linewidth = 0.8
  ) +

  # credible interval ribbon
  geom_ribbon(
    data  = plot_sown_mf_data,
    aes(x = sowndiv, ymin = lower__, ymax = upper__),
    fill  = "#0072B2",
    alpha = 0.15
  ) +

  # axis scaling
  scale_x_continuous(
    breaks = c(1, 2, 4, 8, 16),
    labels = c(1, 2, 4, 8, 16),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  
  labs(
    x = "Sown plant species diversity",
    y = "Ecological multifunctionality"
  ) +
  theme_custom() 

p_sown_mf

# save graph
ggsave(
  filename = "EcologicalMF_sown.png",
  plot     = p_sown_mf,
  width    = 5.9,
  height   = 4.5,
  units    = "in",
  device   = "png",
  dpi      = 300,
  path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/" 
)
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#

# plot MF as a function of meanSR
c_real_mf <- conditional_effects(fitMFreal)
plot_real_mf_data <- c_real_mf[[1]]

p_real_mf <- ggplot() +
  geom_point(data = ecologicalMF,
             aes(x = meanSR, y=multifunctionality),
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
    position = position_jitter(w = 0.03, h = 0)
  ) +
  # Model-estimated effect
  geom_line(data = plot_real_mf_data, aes(x = meanSR, y = estimate__), color = "blue", size = .8, alpha = .6) +
  # Credible interval ribbon
  geom_ribbon(data = plot_real_mf_data, aes(x = meanSR, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(title = "",
       x = "Mean realized plant diversity",
       y = "Ecological multifunctionality") +
  scale_x_continuous(breaks = c(2,4,6,8,10,12))

p_real_mf
# save figure
ggsave("EcologicalMF_realizedSR_noMono.png",
       plot = p_real_mf,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

```{r correlation economic and ecological MF}
# merge reduced ecological and economic multifunctionality data
joinedMF <- economicMF %>%
  select(c(plotcode, costs_total_mid)) %>% 
  left_join(ecologicalMF, by = "plotcode") 

# Correlation between economic and ecological multifunctionality
bf <- correlationBF(joinedMF$multifunctionality, joinedMF$costs_total_mid)

post <- posterior(bf, iterations = 20000)
rho_draws <- post[, "rho"]

quantile(rho_draws, c(.025, .5, .975))
mean(rho_draws)

# plot the correlation
corrPlot <- ggplot(joinedMF, aes(x = multifunctionality, y = costs_total_mid)) +
  # raw data
  geom_point(
    alpha  = 0.5,
    size   = 2.2,
    shape  = 21,
    fill   = "#0072B2",
    colour = "#004C70",
    stroke = 0.4
  ) +
  
  # linear regression line
  geom_smooth(
    method = "lm",
    se     = FALSE,
    linewidth = 0.8,
    colour = "#0072B2"
  ) +
  
  scale_x_continuous(
    breaks = seq(0.1, 0.7, by = 0.1),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    breaks = c(-1000, -500, 0, 500, 1000)
  ) +
  
  labs(
    x = "Ecological multifunctionality",
    y = expression(
  "EMF (" * "\u20AC ha"^{-1} ~ "yr"^{-1} * ")")
  ) +
  theme_custom()

corrPlot

# save figure
ggsave(
  filename = "Corr_EconomicMF_EcologicalMF.png",
  plot     = corrPlot,
  device   = "png",
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
  dpi      = 300,
  width    = 6.27,
  height   = 5,
  units    = "in"
)


# create two panel plot with economic MF
fig_MF_and_corr <- (
  p_sown_total /
  plot_spacer() /
  corrPlot
) +
  plot_layout(heights = c(1, 0.09, 1)) +
  plot_annotation(tag_levels = "a") &
  theme(
    plot.tag.position = c(0.08, 0.98),
    plot.tag = element_text(face = "bold", hjust = 0, vjust = 1),
    axis.title.y = element_text(margin = margin(r = 2)),
    plot.margin = margin(10, 15, 10, 15)
  )

fig_MF_and_corr

ggsave(
  filename = "Fig_MF_and_corr.png",
  plot     = fig_MF_and_corr,
  width    = 5.9,
  height   = 6.6,
  units    = "in",
  dpi      = 300,
  device   = ragg::agg_png,
  path     = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/"
)

```
# Supplements
## Single Function Models
```{r Model Single Functions}
fitMicbio <- brm(
  mean_micbiomass ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitMicbio)

fitMicres <- brm(
  mean_micrespiration ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitMicres)

fitRoot <- brm(
  mean_root_biomass ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitRoot)

fitPol <- brm(
  mean_diversity ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitPol)

fitAbove <- brm(
  mean_above_control ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitAbove)

fitBelow <- brm(
  mean_below_control ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitBelow)

fitAgg <- brm(
  mean_agg ~ log2(sowndiv) + (1|block),
 ecologicalMF)

summary(fitAgg)

```


## Plot the single functions
```{r Plot Single functions}
#plot biomass as a response to sowndiv
c_sown_biomass <- conditional_effects(fitMFsown)
plot_sown_mf_data <- c_sown_mf[[1]]

p_sown_mf <- ggplot() +
  geom_point(data = ecologicalMF,
             aes(x = sowndiv, y=multifunctionality),
    color = "#28a87d",
    alpha = .5,
    shape = 20,
    stroke = 1,
    size = 4,
    position = position_jitter(w = 0.03, h = 0)
  ) +
  # Model-estimated effect
  geom_line(data = plot_sown_mf_data, aes(x = sowndiv, y = estimate__), color = "blue", size = .8, alpha = .6) +
  # Credible interval ribbon
  geom_ribbon(data = plot_sown_mf_data, aes(x = sowndiv, ymin = lower__, ymax = upper__), 
              fill = "#287da8", alpha = 0.2) +
  labs(title = "",
       x = "Sown plant diversity",
       y = "Ecological multifunctionality") +
  scale_x_continuous(breaks = c(1,2,4,8,16))

p_sown_mf

# save figure
ggsave("EcologicalMF_sown_noMono.png",
       plot = p_sown_mf,
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

