---
title: "Data Exploration + Synthesis"
author: "Marius Munschek"
date: "2025-02-20"
output: html_document
---

# Preface
This R Script is to be used to load, explore, and synthesize the data for assessing the (soil) ecological multifunctionality in the Jena experiment and it's response to different levels of plant diversity. The following ecological functions were are incuded: Aboveground biomass production, C&N content in the soil, Realized Species Richness, nitrogen leaching (into the groundwater), pollination, above- and belowground herbivory (for belowground measured as energy fluxes in the food web), soil organic carbon (SOC), leaf dry matter content (LDMC), soil diversity, microbial biomass + respiration. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(warn = TRUE, 
                      message = FALSE, 
                      error = FALSE, 
                      echo = TRUE, 
                      cache = FALSE, 
                      fig.width = 8, 
                      fig.height = 6, 
                      dpi = 300,
                      out.width = "100%",
                      fig.align = "center",
                      comment = NA,
                      tidy = TRUE,
                      collapse = TRUE)
```

# Preparation
## Loading packages
```{r Packages}
library(dplyr)
library(readr)
library(readxl)
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data, include=FALSE}
# load data on plot design (needed for 'block code')
design <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Plot_Design/90_20_data.xlsx")

#load data on realized species richness (ID = 487)
SR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")
  #change data type for realized SR & sowndiv
  SR$realizedSR <- as.numeric(SR$realizedSR)
  SR$sowndiv <- as.numeric(SR$sowndiv)

#load data on root biomass production (ID = 714) 
roots <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Root Biomass_03-24/714_4_data.xlsx") %>% 
  rename_at('plot', ~'plotcode') # rename plot to plotcode

roots$plotcode <- as.factor(roots$plotcode) # change plotcode to a factor
roots$year <- as.factor(roots$year) # change year to a factor  
roots$root.bm <- as.numeric(roots$root.bm) # change root biomass values to numeric

# load data on microbial biomass and respiration from 2002 to 2020 (ID = 78), 2021 (ID = 240) (and 2022 (ID = 398) they measured for two different soil layers, so I rather not include this)
microbial1 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration_03-20/78_5_data.xlsx")
mirobial2 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration_21/240_6_data.xlsx") %>%
  filter(grepl("D3", plot)) %>%  # filter for treatment D3 (main experiment)
  mutate(plot = gsub("D3", "", plot))
  
# change data types
microbial1$basal_respiration <- as.numeric(microbial1$basal_respiration)
microbial1$soil_microbial_biomass_C <- as.numeric(microbial1$soil_microbial_biomass_C)
microbial1$respiratory_quotient <- as.numeric(microbial1$respiratory_quotient)
microbial1$soil_water_content <- as.numeric(microbial1$soil_water_content)
  
# merge the datasets
microbial <- bind_rows(mirobial2, microbial1) %>%
    rename_at('plot', ~'plotcode') %>% # rename plot to plotcode
    mutate(plotcode = as.factor(plotcode)) # change plotcode to a factor


#load pathogen control data (Aboveground by Andrew Barnes (2020), Belowground by Angelos Amyntas (2025))
pathogensAbove <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Fluxes/Barnes_2020_Jena Experiment.xlsx") %>%
  rename_at('plot', ~'plotcode') %>% # rename plot to plotcode
  mutate(plotcode = as.factor(plotcode))  # change plotcode to a factor
  
pathogensBelow <- read_csv("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Fluxes/fluxes1000_dir100.csv") %>% 
  filter(Treatment == "Treatment3") %>%  # filter for treatment D3 (main experiment)
  rename_at('Plot', ~'plotcode') # rename Plot to plotcode

#load data on pollinators
pollinators <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/423_3_data.xlsx") %>% 
  mutate(species_richness = as.numeric(species_richness)) # change species richness to numeric

#load data on aggregate stability
aggregates <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Aggregates/FINAL_WSA_JE_0-10_May2025_ZJ.xlsx", sheet = 2) %>% 
  rename_at('Plotcode', ~'plotcode') %>%  # rename Plotcode to plotcode
  mutate(WSA = as.numeric(`WSA (%)`))
```

## Data wrangling
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated. 
```{r Biomass data}
# calculate mean biomass for root dataset
rootsMean <- roots %>% 
  group_by(plotcode) %>% 
  mutate(mean_root_biomass = mean(root.bm, na.rm = TRUE)) %>% 
  distinct(plotcode, mean_root_biomass) %>%
  ungroup() %>% 
  na.omit() %>%
  mutate(root_biomass_standardized = (mean_root_biomass - min(mean_root_biomass)) / (max(mean_root_biomass)- min(mean_root_biomass))) # standardize the root biomass values

```

```{r Microbial data}
microbialMean <- microbial %>%
  filter(basal_respiration >= 0) %>%  # filter out negative values and years before 2010
  group_by(plotcode) %>% 
  mutate(mean_micbiomass = mean(soil_microbial_biomass_C, na.rm = T), mean_micrespiration = mean(basal_respiration, na.rm = T)
         ) %>% 
  distinct(plotcode, mean_micbiomass, mean_micrespiration) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(micbiomass_standardized = (mean_micbiomass - min(mean_micbiomass)) / (max(mean_micbiomass) - min(mean_micbiomass)), # standardize the microbial biomass values
         micrespiration_standardized = (mean_micrespiration - min(mean_micrespiration)) / (max(mean_micrespiration) - min(mean_micrespiration))
         ) # standardize the microbial respiration values

```

```{r Species Richness data}
realdiv <- SR %>% #calculate the mean realized SR for the years 2003-2019
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)

SR <- distinct(SR, plotcode, sowndiv)

```

```{r Pathogen control}
# calculate the mean aboveground pathogen control per plot
pathogensAboveMean <- pathogensAbove %>% 
  group_by(plotcode) %>% 
  summarise(mean_above_control = mean(top.bottom.ratio, na.rm = TRUE), .groups = "drop") %>%
  mutate(above_pathogen_control_standardized = (mean_above_control - min(mean_above_control)) / (max(mean_above_control) - min(mean_above_control))) # standardize the values

# calculate the mean belowground pathogen control per plot
pathogensBelowMean <- pathogensBelow %>% 
  group_by(plotcode) %>% 
  summarise(mean_below_control = mean(control, na.rm = TRUE), .groups = "drop") %>%
  mutate(below_pathogen_control_standardized = (mean_below_control - min(mean_below_control)) / (max(mean_below_control) - min(mean_below_control))) # standardize the values

```

```{r Pollinators}
# calculate the mean pollinator diversity per plot
pollinatorsMean <- pollinators %>% 
  group_by(plotcode) %>% 
  mutate(mean_diversity = mean(species_richness, na.rm = TRUE)) %>%
  distinct(plotcode, mean_diversity) %>%
  ungroup() %>%
  mutate(mean_div_standardized = (mean_diversity - min(mean_diversity)) / (max(mean_diversity) - min(mean_diversity))) # standardize the pollinator richness values

```  

```{r Aggregates}
aggregatesMean <- aggregates %>% 
  group_by(plotcode) %>% 
  mutate(mean_WSA = mean(WSA, na.rm =T)) %>% 
  distinct(plotcode, mean_WSA) %>% 
  ungroup %>% 
  mutate(stand_WSA = (mean_WSA - min(mean_WSA)) / (max(mean_WSA) - min(mean_WSA))) # standardize the values

```

## Synthesize in one Dataset
```{r Data Synthesis}
ecologicalMF <- microbialMean %>% 
  left_join(rootsMean, by = join_by(plotcode)) %>% 
  left_join(realdiv, by = join_by(plotcode)) %>%
  left_join(select(design, plotcode, block)) %>%
  left_join(SR, by = join_by(plotcode)) %>% 
  left_join(pollinatorsMean, by = join_by(plotcode)) %>%
  left_join(pathogensAboveMean, by = join_by(plotcode)) %>%
  left_join(pathogensBelowMean, by = join_by(plotcode)) %>% 
  left_join(aggregatesMean, by = join_by(plotcode)) %>% 
  filter(sowndiv < 60)

# calculate the ecological multifunctionality as the mean of the standardized values  
ecologicalMF$multifunctionality <- rowMeans(
  ecologicalMF[, c(
    "micbiomass_standardized",
    "root_biomass_standardized",
    "micrespiration_standardized",
    "mean_div_standardized",
    "above_pathogen_control_standardized",
    "below_pathogen_control_standardized",
    "stand_WSA"
  )],
  na.rm = TRUE
)


#save the ecological multifunctionality dataset
writexl::write_xlsx(ecologicalMF, "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/ecological_multifunctionality_reduced.xlsx")

```

