---
title: "Ecological Multifunctinality Analysis"
author: "Marius Munschek"
date: "2025-05-30"
output: html_document
---

# Preface
This R Script is intended to be used to calculate the ecological multifunctionality for (mostly) soil functions in the Jena experiment. I will focus on the following functions:
- Biomass production
- Nitrogen leaching
- Carbon sequestration 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation
## Loading packages
```{r Packages, include=FALSE}
library(dplyr)
library(readr)
library(readxl)
library(lme4)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
```

## Set up a theme
```{r include=FALSE}
theme_set(theme_minimal(base_size = 14)+
            theme(plot.title = element_text(size=13,hjust=0.5),
                  axis.text = element_text(size = 7),
                  axis.text.y = element_text(size = 13),
                  axis.text.x = element_text(size = 13)))
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data, include=FALSE}
# load data on plot design (needed for 'block code')
design <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Plot_Design/90_20_data.xlsx")

#load data on aboveground biomass production (ID = 624) 
biomassDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Biomass_Above/624_9_data.xlsx")

#load data on realized species richness (ID = 487)
SR <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Realized_Species_Richness/487_3_data.xlsx")
  #change data type for realized SR & sowndiv
  SR$realizedSR <- as.numeric(SR$realizedSR)
  SR$sowndiv <- as.numeric(SR$sowndiv)

#load data on soil CN (2022: ID = 655; 2020: ID = 424 2002: ID = 278): 2002 is to compare the carbon content at the beginning of the experiment with the current (2022) carbon content. 2020 data includes bulk density measurements.
  # Dataset on deep soil measurement (ID = 433) from 0-100cm is only one Block (D2) and most recent measurement year is 2016.

soilCN20 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/424_7_data.xlsx")
# soilCN22 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/655_12_data.xlsx")
soilCN02 <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Soil/278_2_data.xlsx") %>% 
  filter(year == 2002) 

#load data on biomass CN (ID = 557)
biomassCN <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/C&N_Biomass_Above/557_3_data.xlsx")

# load data on microbial biomass and respiration from 2022 (ID = 398) and 2021 (ID = 240)
microbial <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/MicBiomass_MicRespiration/398_3_data.xlsx")

#load data on soil aggregates

#load data on enzyme activity?

#load climate data
climate <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Climate_Data/484_7_data.xlsx")

#load LDMC data
ldmc <- read_csv("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/LDMC/Leaf_Root_traits_dbef_2021.csv")

#load energy flux data (for belowground herbivory calculation)

#load aboveground herbivory data (only treatment D3 -> main experiment)
herbivoryAbove <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Herbivory_Above/480_4_data.xlsx") %>%  filter(treatment == "D3")

#load data on soil diversity

#load data on nitrogen leaching (this should be in kg/ha per year)
leaching <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/WaterN/Overall_Leaching_Plot.xlsx")

#load data on pollinators
pollinators <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

#load data on flower cover
flowers <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/Pollinators/PollinatorObservation_Main_2011_id140.xlsx")

```

## Data wrangling
Before the individual datasets can be merged, they have to be harmonized. Additionally for some datasets, aggregated values have to be calculated. 
```{r Biomass data}
biomassDat$plotcode <- as.factor(biomassDat$plotcode) #change the plotcode to a factor
# samplingSeasons <- unique(biomassDat[c("season","year")]) # used to check whether every year had the same number of sampling seasons

# calculate total biomass per plot for 2022
biomassDat <- biomassDat %>% 
  filter(year == 2022) %>% 
  group_by(plotcode, year) %>% 
  summarise(sum_biomass = sum(biomass, na.rm = TRUE)/2, .groups = "drop") # divide by 2 to get the mean biomass per plot (2 subsamples per plot)

# introduce new column with SGM
biomassDat$SGM <- (biomassDat$sum_biomass*0.1)*(6.15)-96

# exclude the years before 2018
# sum up biomass for each plot and calculate the mean across subsamples and years
biomassMeans <- biomassDat %>% 
  filter(year > 2018) %>% 
  group_by(plotcode, year, season) %>% 
  summarise(sum_biomass = sum(biomass, na.rm = TRUE), .groups = "drop") %>%
  group_by(plotcode) %>%
  summarise(mean_biomass = mean(sum_biomass, na.rm = TRUE), .groups = "drop")

biomassMeans$SGM <- (biomassMeans$mean_biomass*0.1)*(6.83)-180

```

```{r Yield stability}
#here I need to look at the harvested biomass across several years and how it fluctuates (calculate the mean difference/variance?)


```

```{r Biomass CN data}
biomassCN <- biomassCN %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
biomassCN$`C:N ratio` <- as.numeric(biomassCN$`C:N ratio`) # change CN ratio values to numeric

biomassCN <- biomassCN %>% group_by(plotcode) %>%                      # calculate mean CN value between the two measurements
  mutate(CNratio_av = mean(`C:N ratio`, na.rm = TRUE))
```

```{r Microbial data}
microbial <- microbial %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
```

```{r Soil CN}
# rename plot to plotcode and change data types 
soilCN02 <- soilCN02 %>%
  rename_at('plot', ~'plotcode') %>% 
  mutate(
    plotcode = as.factor(plotcode),
    Corg_Soil = as.numeric(Corg_Soil),
    Cinorg_Soil = as.numeric(Cinorg_Soil)
  )
  
soilCN20 <- soilCN20 %>%
  rename_at('plot', ~'plotcode') %>%
  mutate(
    plotcode = as.factor(plotcode),
    Corg_Soil = as.numeric(Corg_Soil),
    Cinorg_Soil = as.numeric(Cinorg_Soil)
  )
 
# soilCN22 <- soilCN22 %>% rename_at('plot', ~'plotcode') # rename plot to plotcode
#  soilCN22$plotcode <- as.factor(soilCN22$plotcode)

# create a new column with mean C across layers per plot for 2002
soilCN02_clean <- soilCN02 %>% group_by(plotcode) %>% 
                    mutate(mean_org_02 = mean(Corg_Soil), mean_inorg_02 = mean(Cinorg_Soil)) %>%
                      distinct(plotcode, mean_org_02, mean_inorg_02) %>% 
                      mutate(mean_C_02 = mean_org_02 + mean_inorg_02)

# create a new column with mean C across layers per plot for 2020
soilCN20_clean <- soilCN20 %>% group_by(plotcode) %>% 
                      mutate(mean_org_20 = mean(Corg_Soil), mean_inorg_20 = mean(Cinorg_Soil), mean_bulk = mean(BulkDensity_Soil)) %>%
                      distinct(plotcode, mean_org_20, mean_inorg_20, mean_bulk) %>% 
                      mutate(mean_C_20 = mean_org_20 + mean_inorg_20)


# create a new column with mean C across layers per plot for 2022
# soilCN22_clean <- soilCN22 %>% group_by(plotcode) %>% 
#                      mutate(mean_C_22 = mean(TC), mean_N_22 = mean(TN)) %>%
#                      distinct(plotcode, mean_C_22, mean_N_22)  
                      
# convert carbon unit from percentage to g/kg 
# soilCN22_clean$mean_C_mass_22 <- soilCN22_clean$mean_C_22 * 10
  
```

```{r Species Richness data}
SR <- distinct(SR, plotcode, sowndiv)

realdiv <- filter(SR, year > 2018 & season == "Total" ) #keep only total SR from year 2018 onwards

realdiv <- SR %>%                    #calculate the mean realized SR for the years 2003-2019
  group_by(plotcode) %>% 
  mutate(meanSR = mean(realizedSR, na.rm = TRUE)) %>% 
  ungroup() %>%
  distinct(plotcode, meanSR)
```


```{r leaching}
leaching <- leaching %>% 
            rename_at('plot', ~'plotcode') # rename plot to plotcode

# calculate the costs in €/ha/a for each plot

leaching$leach_costs <- leaching$TNleaching_av * 11.23 *(-1)
```

```{r Herbivory}
herbivoryAbove <- herbivoryAbove %>% 
            rename_at('plot', ~'plotcode') # rename plot to plotcode

```

## Synthesize in one Dataset
```{r Data Synthesis}
synthDat <- biomassMeans %>% # merge the biomass and the realized SR data
  left_join(SR, by = join_by(plotcode)) %>%
  left_join(soilCN02_clean, by = join_by(plotcode)) %>% 
  left_join(soilCN20_clean, by = join_by(plotcode)) %>% 
  left_join(select(design, plotcode, block)) %>% 
  left_join(leaching, by = join_by(plotcode)) # %>%
  left_join(herbivoryAbove, by )

synthDat$block <- as.factor(synthDat$block) # change block to a factor
synthDat$plotcode <- as.factor(synthDat$plotcode)  # change plotcode to a factor
  
# create a column with carbon difference between 2002 and 2022
synthDat$mean_org_diff <- synthDat$mean_org_20 - synthDat$mean_org_02

# create a column with social costs of carbon
synthDat$SOC_costs <- synthDat$mean_org_diff * synthDat$mean_bulk * 0.3 * 10000 / 1000 * 3.66 * 195 / 20 # in €/ha/a and converted to CO2 and a carbon CO2 price of 195€/t

# add up all the costs for total value
synthDat$costs_total <- synthDat$SOC_costs + synthDat$leach_costs + synthDat$SGM



```

## Calculations
### Ecological Multifunctionality (averaging approach)
```{r Ecological MF}
# I need to standardize all the values and then average across functions/services

```

### Multidiversity Index
```{r Multidiversity}

```

### Soil Health Index
```{r Soil health}

```


## Statistical Models 
```{r Biomass ~ SR}

# with sown species number
lmBioSown <- lm(
  mean_biomass ~ log2(as.numeric(sowndiv)),
  synthDat)

plot(lmBioSown)
summary(lmBioSown)

# with realized species diversity
lmSGM <- lm(
  SGM ~ log2(as.numeric(sowndiv)),
  synthDat) 

plot(lmSGM) # check model assumptions
summary(lmSGM)


# check if species richness increases biomass

# check if species richness increases SOC
lmSOC <- lmer(
  mean_C_20 ~ log2(sowndiv) + (1|block),
  synthDat)

summary(lmSOC)

# check if block has effect on SOC differences between 2002 and 2020

lmSOCdiff <- lmer(
  mean_org_diff ~ log2(sowndiv) + (1|block),
  synthDat)

summary(lmSOCdiff)

# check if species richness increases total value
lmCostsSR <- lm(
  costs_total ~ log2(as.numeric(sowndiv)),
  synthDat)

plot(lmCostsSR)
summary(lmCostsSR)

```


```{r SOC ~ realized SR}

```

```{r N leaching ~ realized SR}
#check the distribution of the data
hist(synthDat$TNleaching_av)
plot(synthDat$sowndiv, synthDat$TNleaching_av)

lmLeach <- lm(TNleaching_av ~ log2(sowndiv), synthDat)
plot(lmLeach)
summary(lmLeach)
```

```{r HerbivoryAbove ~ Realized SR}

```


## Plotting
### Plot biomass against realized species richness
```{r Plotting}
#check distribution of biomass data
p1 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = mean_biomass)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "Summed Biomass per Plot in 2022 (g)")

p1

ggsave("Biomass.png",
       plot = last_plot(),
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of SGM data
p2 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = SGM)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "standard gross margin (€/ha)")

ggsave("SGM.png",
       plot = last_plot(),
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of climate regulation cost data
p3 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = SOC_costs)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "climate regulation costs/gains (€/ha)")

ggsave("SOC.png", plot = last_plot(), device = "png", path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/", width = 10, height = 15, units = "cm", dpi = 300)

# check distribution of leaching costs data
p4 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = leach_costs)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "leaching costs (€/ha/a)")

ggsave("Leaching.png", plot = last_plot(), device = "png", path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/", width = 10, height = 15, units = "cm", dpi = 300)

# combine the above 4 plots in one grid
p1 + p2 + p3 + p4 +
  plot_layout(ncol = 2) +
  plot_annotation(title = "") &
  theme(plot.title = element_text(size=20,hjust=0.5))

ggsave("Single_Services.png",
       plot = last_plot(),
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of total costs across sowndiv
p5 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = costs_total)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    stroke = 1,
    size = 4
  ) +
  labs(
    x = "",
    y = ""
    ) +
   geom_smooth(
    method = "lm"
  )

p5

ggsave("TotalCosts2.png",
       plot = p5,
       device = "png",
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

# create boxplots for each factorial sowndiv and SGM
#p1 <- ggplot2::ggplot(synthDat,
#                aes(x = as.factor(sowndiv), y = SGM)
#                ) +
#  geom_boxplot(
#  ) +
#  labs(
#    x = "Sown Plant Diversity",
#    y = "SGM (€/ha)")


# create boxplots for each factorial sowndiv and leaching costs
#p2 <- ggplot2::ggplot(synthDat,
#                aes(x = as.factor(sowndiv), y = leach_costs)
#                ) +
#  geom_boxplot(
#  ) +
#  labs(
#    x = "Sown Plant Diversity",
#    y = "Social costs of nitrogen leaching (€/ha/a)")



# create boxplots for each factorial sowndiv and SOC costs
#p3 <- ggplot2::ggplot(synthDat,
#                aes(x = as.factor(sowndiv), y = SOC_costs)
#                ) +
#  geom_boxplot(
#  ) +
#  labs(
#    x = "Sown Plant Diversity",
#    y = "Social costs/benefits of climate regulation (€/ha/a)")



# combine the three plots above in a grid
#library(patchwork)
#p1 + p2 + p3 +
#  plot_layout(ncol = 3) +
#  plot_annotation(title = "Costs and Benefits of Sown Plant Diversity") &
#  theme(plot.title = element_text(size=20,hjust=0.5))

#ggsave("Costs_Benefits.png", plot = last_plot(), device = "png", path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/", width = 10, height = 15, units = "cm", dpi = 300)


```

