---
title: "Economic Multifunctionality Analysis"
author: "Marius Munschek"
date: "2025-05-28"
output: html_document
---

#Preface
This R Script is intended to be used to calculate the economic multifunctionality for ecosystem services in the Jena experiment. I will focus on the following ecosystem sevices:
- Biomass production
- Nitrogen leaching
- Carbon sequestration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation
## Loading packages
```{r Packages, include=FALSE}
library(dplyr)
library(readr)
library(readxl)
library(lme4)
library(ggplot2)
library(patchwork)
library(brms)
library(performance)
library(ggpmisc)
```

## Set up a theme
```{r include=FALSE}
theme_set(theme_classic(base_size = 14)+
            theme(plot.title = element_text(size=13,hjust=0.5),
                  axis.text = element_text(size = 7),
                  axis.text.y = element_text(size = 13),
                  axis.text.x = element_text(size = 13)))
```

## Load datasets and filter for relevant observations/measurements
```{r Read Data, include=FALSE}
#load synthesis dataset
synthDat <- read_xlsx("C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/1_Data/synthesized_means_data.xlsx")

```

## Data wrangling
Change data types and filter for relevant observations.
```{r Synthesis data}
synthDat$plotcode <- as.factor(synthDat$plotcode) #change the plotcode to a factor

# add up all the low costs for lower total value
synthDat$costs_total_lo <- synthDat$SOC_costs_lo + synthDat$leach_costs_lo + synthDat$SGM_mean

# add up all the high costs for higher total value
synthDat$costs_total_hi <- synthDat$SOC_costs_hi + synthDat$leach_costs_hi + synthDat$SGM_mean

noMono <- synthDat %>% 
  filter(sowndiv > 1 & sowndiv < 60) #filter out monocultures and all-species-mixtures

```


## Plot the densities of the data
```{r Densities, include=FALSE}

# check distribution of sowndiv
p1 <- ggplot2::ggplot(synthDat,
                aes(x = log2(sowndiv))
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "Density"
  )

p1

# check distribution of realSR
p2 <- ggplot2::ggplot(noMono,
                aes(x = meanSR)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Realized Species Richness",
    y = "Density"
  )

p2

p3 <- ggplot2::ggplot(synthDat,
                aes(x = mean_biomass)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Biomass (g)",
    y = "Density"
  )

p4 <- ggplot2::ggplot(synthDat,
                aes(x = SGM_mean)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Standard Gross Margin (€/ha)",
    y = "Density"
  )

p5 <- ggplot2::ggplot(synthDat,
                aes(x = SOC_costs_mid)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "SOC Costs (€/ha)",
    y = "Density"
  )

p6 <- ggplot2::ggplot(synthDat,
                aes(x = leach_costs_mid)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Leaching Costs (€/ha/a)",
    y = "Density"
  )

p7 <- ggplot2::ggplot(synthDat,
                aes(x = costs_total_mid)
                ) +
  geom_density(
    fill = "#28a87d",
    alpha = .5
  ) +
  labs(
    x = "Total Costs (€/ha/a)",
    y = "Density"
  )


# show all in one grid
p1 + p2 + p3 + p4 + p5 + p6 + p7 +
  plot_layout(ncol = 3) +
  plot_annotation(title = "Densities of Sown Plant Diversity, Realized Species Richness, Biomass, SGM, SOC Costs, Leaching Costs and Total Costs") &
  theme(plot.title = element_text(size=20,hjust=0.5))

ggsave("Densities.png",
       plot = last_plot(),
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

```

## Statistical Models 

```{r Sown Diversity WITHOUT Monocultures, include=FALSE}

# Biomass against species richness
fitBiomassSown <- brm(
  mean_biomass ~ log2(sowndiv),
  noMono)

    # check model assumptions
    pp_check(fitBiomassSR2, type="scatter_avg") 
    pp_check(fitBiomassSR2, ndraws=50)
    check_model(fitBiomassSR2, check="linearity") 
    check_model(fitBiomassSR2, check="homogeneity")
    plot(fitBiomassSR2)
    
    # check output of model
    summary(fitBiomassSR2)

# SGM against species richness
fitSGMsown <- brm(
  SGM_mean ~ log2(sowndiv),
  noMono) 

    # check model assumptions 
    pp_check(fitSGM2, type="scatter_avg")
    pp_check(fitSGM2, ndraws=50)
    check_model(fitSGM2, check="linearity")
    check_model(fitSGM2, check="homogeneity")
    plot(fitSGM2)
    
    # check output of model
    summary(fitSGM2)

# SOC against species richness
fitSOCsown <- brm(
  mean_org_diff ~ log2(sowndiv),
  noMono)

    # check model assumptions
    pp_check(fitSOC2, type="scatter_avg") 
    pp_check(fitSOC2, ndraws=50)
    check_model(fitSOC2, check="linearity") 
    check_model(fitSOC2, check="homogeneity")
    plot(fitSOC2)
    
    # check output of model
    summary(fitSOC2)

# Leaching costs against species richness
fitLeachSown <- brm(
  leach_costs_mid ~ log2(sowndiv),
  noMono)

    # check model assumptions
    pp_check(fitLeach2, type="scatter_avg") 
    pp_check(fitLeach2, ndraws=50)
    check_model(fitLeach2, check="linearity") 
    check_model(fitLeach2, check="homogeneity")
    plot(fitLeach2)
    
    # check output of model
    summary(fitLeach2)

# Total Costs against species richness
fitTotalSown <- brm(
  costs_total_mid ~ log2(sowndiv),
  noMono)

    # check model assumptions
    pp_check(fitTotal2, type="scatter_avg") 
    pp_check(fitTotal2, ndraws=50)
    check_model(fitTotal2, check="linearity") 
    check_model(fitTotal2, check="homogeneity")
    plot(fitTotal2)
    
    # check output of model
    summary(fitTotal2)

fitTotal2.1 <- brm(
  bf(costs_total_mid ~ log2(sowndiv),
     sigma ~ log2(sowndiv)),
  noMono)

    # check model assumptions
    pp_check(fitTotal2.1, type="scatter_avg") 
    pp_check(fitTotal2.1, ndraws=50)
    check_model(fitTotal2.1, check="linearity") 
    check_model(fitTotal2.1, check="homogeneity")
    plot(fitTotal2.1)        

    loo(fitTotal2, fitTotal4) # compare the two models

```


```{r Realized Diversity without Monocultures, include=FALSE}

# Biomass against species richness
fitBiomassReal <- brm(
  mean_biomass ~ meanSR,
  noMono)

    # check model assumptions
    pp_check(fitBiomassSR4, type="scatter_avg") 
    pp_check(fitBiomassSR4, ndraws=50)
    check_model(fitBiomassSR4, check="linearity") 
    check_model(fitBiomassSR4, check="homogeneity")
    plot(fitBiomassSR4)
    
    # check output of model
    summary(fitBiomassSR4)

# SGM against species richness
fitSGMreal <- brm(
  SGM_mean ~ meanSR,
  noMono) 

    # check model assumptions 
    pp_check(fitSGM4, type="scatter_avg")
    pp_check(fitSGM4, ndraws=50)
    check_model(fitSGM4, check="linearity")
    check_model(fitSGM4, check="homogeneity")
    plot(fitSGM4)
    
    # check output of model
    summary(fitSGM4)

# SOC against species richness
fitSOCreal <- brm(
  mean_org_diff ~ meanSR,
  noMono)

    # check model assumptions
    pp_check(fitSOC4, type="scatter_avg") 
    pp_check(fitSOC4, ndraws=50)
    check_model(fitSOC4, check="linearity") 
    check_model(fitSOC4, check="homogeneity")
    plot(fitSOC4)
    
    # check output of model
    summary(fitSOC4)

# Leaching costs against species richness
fitLeachReal <- brm(
  leach_costs_mid ~ meanSR,
  noMono)

    # check model assumptions
    pp_check(fitLeach4, type="scatter_avg") 
    pp_check(fitLeach4, ndraws=50)
    check_model(fitLeach4, check="linearity") 
    check_model(fitLeach4, check="homogeneity")
    plot(fitLeach4)
    
    # check output of model
    summary(fitLeach4)

# Total Costs against species richness
fitTotalReal <- brm(
  costs_total_mid ~ meanSR,
  noMono)

    # check model assumptions
    pp_check(fitTotalReal, type="scatter_avg") 
    pp_check(fitTotalReal, ndraws=50)
    check_model(fitTotalReal, check="linearity") 
    check_model(fitTotalReal, check="homogeneity")
    plot(fitTotalReal)
    
    # check output of model
    summary(fitTotalReal)

shapiro.test(residuals(fitTotalReal)) # check normality of residuals

    
```

### Sensitivity Analysis with different price scenarios

```{r Realized Diversity without Monocultures, include=FALSE}

# Low Total Costs against species richness
fitTotalLo <- brm(
  costs_total_mid ~ meanSR,
  noMono)

    # check model assumptions
    pp_check(fitTotal4, type="scatter_avg") 
    pp_check(fitTotal4, ndraws=50)
    check_model(fitTotal4, check="linearity") 
    check_model(fitTotal4, check="homogeneity")
    plot(fitTotal4)
    
    # check output of model
    summary(fitTotal4)

# High Total Costs against species richness
fitTotalHi <- brm(
  costs_total_mid ~ meanSR,
  noMono)

    # check model assumptions
    pp_check(fitTotal4, type="scatter_avg") 
    pp_check(fitTotal4, ndraws=50)
    check_model(fitTotal4, check="linearity") 
    check_model(fitTotal4, check="homogeneity")
    plot(fitTotal4)
    
    # check output of model
    summary(fitTotal4)

    
```


## Plotting
### Plot biomass against realized species richness
```{r Plotting}
#check distribution of biomass data with monocultures
p1 <- ggplot2::ggplot(synthDat,
                aes(x = log2(sowndiv), y = mean_biomass)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "Summed Biomass per Plot in 2022 (g)")

p1

ggsave("Biomass.png",
       plot = last_plot(),
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of biomass data without monocultures
p1_noMono <- ggplot2::ggplot(noMono,
                aes(x = log2(sowndiv), y = mean_biomass)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "Summed Biomass per Plot in 2022 (g)")

p1_noMono

#check distribution of SGM data with monocultures
p2 <- ggplot2::ggplot(synthDat,
                aes(x = as.numeric(meanSR), y = SGM)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "standard gross margin (€/ha)")

p2

ggsave("SGM.png",
       plot = last_plot(),
       device = "png",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of climate regulation cost data
p3 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = SOC_costs)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "climate regulation costs/gains (€/ha)")

ggsave("SOC.png", plot = last_plot(), device = "png", path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/", width = 10, height = 15, units = "cm", dpi = 300)

# check distribution of leaching costs data
p4 <- ggplot2::ggplot(synthDat,
                aes(x = log2(as.numeric(sowndiv)), y = leach_costs)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 4
  ) +
   geom_smooth(
    method = "lm"
  ) +
  labs(
    x = "log2(Sown Plant Diversity)",
    y = "leaching costs (€/ha/a)")

ggsave("Leaching.png", plot = last_plot(), device = "png", path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/", width = 10, height = 15, units = "cm", dpi = 300)

# combine the above 4 plots in one grid
p1 + p2 + p3 + p4 +
  plot_layout(ncol = 2) +
  plot_annotation(title = "") &
  theme(plot.title = element_text(size=20,hjust=0.5))

ggsave("Single_Services.png",
       plot = last_plot(),
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of mid total costs across sown diversity
pCostsSown <- ggplot2::ggplot(noMono,
                aes(x = log2(sowndiv), y = costs_total_mid)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    stroke = 1,
    size = 4
  ) +
  labs(
    x = "",
    y = ""
    ) +
   geom_smooth(
    method = "lm",
    se = FALSE,
    color = "blue"
  ) +
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 5
  )

pCostsSown

#check distribution of mid total costs across realized diversity
pCostsReal <- ggplot2::ggplot(noMono,
                aes(x = meanSR, y = costs_total_mid)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    stroke = 1,
    size = 4
  ) +
  labs(
    x = "",
    y = ""
    ) +
   geom_smooth(
    method = "lm",
    se = FALSE,
    color = "blue"
  ) +
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE,
    size = 5
  )

pCostsReal

ggsave("TotalCosts2.png",
       plot = p5,
       device = "png",
       bg = "white",
       path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/",
       dpi = 300)

#check distribution of low total costs across sowndiv
p6 <- ggplot2::ggplot(noMono,
                aes(x = log2(sowndiv), y = costs_total_lo)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    stroke = 1,
    size = 4
  ) +
  labs(
    x = "",
    y = ""
    ) +
   geom_smooth(
    method = "lm"
  )

p6

#check distribution of high total costs across sowndiv
p7 <- ggplot2::ggplot(noMono,
                aes(x = log2(sowndiv), y = costs_total_hi)
                ) +
  geom_point(
    color = "#28a87d",
    alpha = .5,
    stroke = 1,
    size = 4
  ) +
  labs(
    x = "",
    y = ""
    ) +
   geom_smooth(
    method = "lm"
  )

p7


# create boxplots for each factorial sowndiv and SGM
#p1 <- ggplot2::ggplot(synthDat,
#                aes(x = as.factor(sowndiv), y = SGM)
#                ) +
#  geom_boxplot(
#  ) +
#  labs(
#    x = "Sown Plant Diversity",
#    y = "SGM (€/ha)")


# create boxplots for each factorial sowndiv and leaching costs
#p2 <- ggplot2::ggplot(synthDat,
#                aes(x = as.factor(sowndiv), y = leach_costs)
#                ) +
#  geom_boxplot(
#  ) +
#  labs(
#    x = "Sown Plant Diversity",
#    y = "Social costs of nitrogen leaching (€/ha/a)")



# create boxplots for each factorial sowndiv and SOC costs
#p3 <- ggplot2::ggplot(synthDat,
#                aes(x = as.factor(sowndiv), y = SOC_costs)
#                ) +
#  geom_boxplot(
#  ) +
#  labs(
#    x = "Sown Plant Diversity",
#    y = "Social costs/benefits of climate regulation (€/ha/a)")



# combine the three plots above in a grid
#library(patchwork)
#p1 + p2 + p3 +
#  plot_layout(ncol = 3) +
#  plot_annotation(title = "Costs and Benefits of Sown Plant Diversity") &
#  theme(plot.title = element_text(size=20,hjust=0.5))

#ggsave("Costs_Benefits.png", plot = last_plot(), device = "png", path = "C:/Users/mm68tulo/Desktop/Jena Chapter/Data Analysis/3_Graphs/", width = 10, height = 15, units = "cm", dpi = 300)


```

## Supplementary Analysis
### With Monocultures

```{r Sown Diversity WITH Monocultures, include=FALSE}

# Biomass against species richness
fitBiomassSR <- brm(
  mean_biomass ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitBiomassSR, type="scatter_avg") 
    pp_check(fitBiomassSR, ndraws=50)
    check_model(fitBiomassSR, check="linearity") 
    check_model(fitBiomassSR, check="homogeneity")
    plot(fitBiomassSR)
    
    # check output of model
    summary(fitBiomassSR)

# SGM against species richness
fitSGM <- brm(
  SGM_mean ~ log2(sowndiv),
  synthDat) 

    # check model assumptions 
    pp_check(fitSGM, type="scatter_avg")
    pp_check(fitSGM, ndraws=50)
    check_model(fitSGM, check="linearity")
    check_model(fitSGM, check="homogeneity")
    plot(fitSGM)
    
    # check output of model
    summary(fitSGM)

# SOC against species richness
fitSOC <- brm(
  mean_org_diff ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitSOC, type="scatter_avg") 
    pp_check(fitSOC, ndraws=50)
    check_model(fitSOC, check="linearity") 
    check_model(fitSOC, check="homogeneity")
    plot(fitSOC)
    
    # check output of model
    summary(fitSOC)

# Leaching costs against species richness
fitLeach <- brm(
  leach_costs_mid ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitLeach, type="scatter_avg") 
    pp_check(fitLeach, ndraws=50)
    check_model(fitLeach, check="linearity") 
    check_model(fitLeach, check="homogeneity")
    plot(fitLeach)
    
    # check output of model
    summary(fitLeach)

# Total Costs against species richness
fitTotal <- brm(
  costs_total_mid ~ log2(sowndiv),
  synthDat)

    # check model assumptions
    pp_check(fitTotal, type="scatter_avg") 
    pp_check(fitTotal, ndraws=50)
    check_model(fitTotal, check="linearity") 
    check_model(fitTotal, check="homogeneity")
    plot(fitTotal)
    
    # check output of model
    summary(fitTotal)
    
```

```{r Realized Diversity With Monocultures, include=FALSE}

# Biomass against species richness
fitBiomassSR3 <- brm(
  mean_biomass ~ meanSR,
  synthDat)

    # check model assumptions
    pp_check(fitBiomassSR3, type="scatter_avg") 
    pp_check(fitBiomassSR3, ndraws=50)
    check_model(fitBiomassSR3, check="linearity") 
    check_model(fitBiomassSR3, check="homogeneity")
    plot(fitBiomassSR3)
    
    # check output of model
    summary(fitBiomassSR3)

# SGM against species richness
fitSGM3 <- brm(
  SGM_mean ~ meanSR,
  synthDat) 

    # check model assumptions 
    pp_check(fitSGM3, type="scatter_avg")
    pp_check(fitSGM3, ndraws=50)
    check_model(fitSGM3, check="linearity")
    check_model(fitSGM3, check="homogeneity")
    plot(fitSGM3)
    
    # check output of model
    summary(fitSGM3)

# SOC against species richness
fitSOC3 <- brm(
  mean_org_diff ~ meanSR,
  synthDat)

    # check model assumptions
    pp_check(fitSOC3, type="scatter_avg") 
    pp_check(fitSOC3, ndraws=50)
    check_model(fitSOC3, check="linearity") 
    check_model(fitSOC3, check="homogeneity")
    plot(fitSOC3)
    
    # check output of model
    summary(fitSOC3)

# Leaching costs against species richness
fitLeach3 <- brm(
  leach_costs_mid ~ meanSR,
  synthDat)

    # check model assumptions
    pp_check(fitLeach3, type="scatter_avg") 
    pp_check(fitLeach3, ndraws=50)
    check_model(fitLeach3, check="linearity") 
    check_model(fitLeach3, check="homogeneity")
    plot(fitLeach3)
    
    # check output of model
    summary(fitLeach3)

# Total Costs against species richness
fitTotal3 <- brm(
  costs_total_mid ~ meanSR,
  synthDat)

    # check model assumptions
    pp_check(fitTotal3, type="scatter_avg") 
    pp_check(fitTotal3, ndraws=50)
    check_model(fitTotal3, check="linearity") 
    check_model(fitTotal3, check="homogeneity")
    plot(fitTotal3)
    
    # check output of model
    summary(fitTotal3)


```
